### 2023-11-13 Updates/2023-10-01.Rmd
updates

---
title: "Preliminary aquatic challenge synthesis"
output: html_document
date: "2023-10-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(arrow)
library(lubridate)
library(tidytext)

cols_modeltype <- c('empirical' = "#482576FF",
          'multi-model ensemble' =  "#35608DFF",
          'process' = "#21908CFF",
          'ML' = '#BBDF27FF',
          'null' = '#43BF71FF')
```

## Inventory of submissions

To start we can look at what the submissions look like, number of teams, general characteristics, which variables and sites were forecasted etc. Some of this might go in the methods and/or results.

```{r message=FALSE}
# read the metadata from the google sheet
googlesheets4::gs4_deauth() 
model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing", sheet = 'descriptions') |> 
  mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0)) |> 
  na.omit(variable)
```

What and where was forecasted?

```{r}
scores <- arrow::open_dataset(file.path(here::here(),"scores")) |> 
  filter(model_id %in% model_meta$model_id)

distinct_submissions <- scores |> 
  distinct(site_id, variable, model_id) |> 
  collect()

head(distinct_submissions)
```

How many sites did each model forecast?

```{r}
distinct_submissions |> 
  group_by(model_id) |> 
  distinct(site_id) |> 
  summarise(site_n = n()) |> 
  group_by(site_n) |> 
  summarise(n_models = n())
```

How many variables did each model forecast?

```{r}
distinct_submissions |> 
  group_by(model_id) |> 
  distinct(variable) |> 
  summarise(variable_n = n()) |> 
  group_by(variable_n) |> 
  summarise(n_models = n())
```

When assessing model performance we can look at raw scores (CRPS) or skill - score relative to a null model. We can use the climatology null model to calculate the skill scores for each model

```{r}
lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')

# extract just the climatology scores
climatology_lake <- scores |>
  filter(model_id == 'climatology', 
         site_id %in% lake_sites,
         variable == 'temperature') |>
  collect() |> 
  select(reference_datetime, datetime, crps, logs, site_id, model_id, variable) |> 
  arrange(reference_datetime, datetime, site_id) |> 
  pivot_wider(values_from = c(crps, logs),
              names_from = model_id)

# get the model scores
lake_scores <- scores |>
  filter(site_id %in% lake_sites,
         model_id %in% model_meta$model_id,
         variable == 'temperature') |>
  collect()

# Calculate the skill scores
lake_skill <- lake_scores |>
  filter(model_id != 'climatology') |> 
  select(reference_datetime, datetime, crps, logs, site_id, model_id, variable) |> 
  full_join(climatology_lake, by = c('reference_datetime', 'datetime', 'site_id', 'variable')) |>

  # calcaulte the skill relative to climatology (- is bad, + is good)
  mutate(skill_crps = crps_climatology - crps,
         skill_logs = logs_climatology - logs,
         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  # consistent forecast period
  filter(horizon <= 30,
         horizon >= 1) |> 
  arrange(reference_datetime, site_id, datetime, model_id)

```

### Model structure

Looking at the structure of the forecast models. Are they dynamic, include initial conditions, update parameters or use data assimilation?

```{r}
# Aspects of DA
# do the forecasts contain initial conditions?
model_meta |> 
  summarise(initial_cond = sum(initial_cond, na.rm = T),
            # do the forecast models update parameters?
            updates_parameters = sum(updates_parameters, na.rm = T),
            # is the forecast 'dynamic'?
            dynamic = sum(dynamic, na.rm = T))
```

How does model structure relate to performance?

```{r}
lake_skill |> 
  group_by(model_id) |> 
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  slice_max(mean_skill, n= 10, na_rm = T) |> 
  left_join(model_meta) |> 
  mutate(model_id = fct_reorder(model_id, mean_skill)) |>
  ggplot(aes(x=mean_skill, 
             y = model_id,
             fill = model_type)) +
  geom_bar(stat = 'identity') +
  tidytext::scale_y_reordered() +
  labs(y='model_id', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  scale_x_continuous(limits = c(-0.2,0.2),
                     breaks = c(-0.2,-0.1,0,0.1,0.2))+
  theme_bw()

```

But the best models (and model types) are not consistent across sites. Looking at the individual sites:

```{r}
lake_skill |> 
  group_by(model_id, site_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  group_by(site_id) |> 
  slice_max(mean_crps, n = 5, na_rm = T) |> 
  mutate(model_id = fct_reorder(model_id, desc(mean_crps))) |>
  left_join(model_meta) |> 
  ggplot(aes(x=mean_crps, 
             y = reorder_within(x = model_id, within = site_id, by = mean_crps),
             fill = model_type)) +
  geom_bar(stat = 'identity') +
  scale_y_reordered() +
  scale_fill_manual(values = cols_modeltype) +
  facet_wrap(~site_id, scales = 'free_y', ncol = 2) +
  labs(y='model_id', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0)
```

We can also look at how this varies across the forecast horizon. For one site it might look like this:

```{r}
# first find the top models
SUGG_mods <- 
  lake_skill |> 
  filter(site_id == 'SUGG') |> 
  mutate(horizon = datetime - as_date(reference_datetime)) |> 
  group_by(model_id) |> 
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |> 
  slice_max(mean_skill, n = 10, na_rm = T)

lake_skill |> 
  filter(site_id == 'SUGG',
         model_id %in% SUGG_mods$model_id) |> 
  mutate(horizon = datetime - as_date(reference_datetime)) |> 
  group_by(model_id, horizon, site_id) |> 
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>
  left_join(model_meta)  |>

  ggplot(aes(x=horizon, y=mean_skill, colour = model_id)) +
  geom_line() +
  facet_wrap(~site_id) +
  geom_hline(yintercept = 0) +
  theme_bw()
```

Generally forecast skill reduces across the horizon, with more models beating climatology null at shorter horizons and only a few at longer horizons. For SUGG example, cb_prophet and tg_randfor are the only ones to be skillful across the full horizon.

Another way to think about this would be to look at generally how do model types do across all site? Or how do all our models do at forecasting particular site? Are some sites more "forecastable" than others?

We can look at the top 10 models on average to see how they do overall and at each site.

```{r}
# top 10 average models
top_mods <- lake_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  slice_max(mean_crps, n = 10, na_rm = T) |> 
  distinct(model_id)

lake_scores |> 
  left_join(model_meta, by = 'model_id') |>
  filter(model_id %in% top_mods$model_id) |>
  ggplot(aes(x=crps, y = model_type)) +
  geom_violin(aes(colour = model_type))

lake_scores |> 
  left_join(model_meta, by = 'model_id') |> 
  filter(model_id %in% top_mods$model_id) |>
  # group_by(model_type, site_id) |> 
  # summarise(mean_crps = mean(crps, na.rm = T)) |> 
  # mutate(model_type = fct_reorder(model_type, desc(mean_crps))) |>
  ggplot(aes(x=crps, y = model_type)) +
  geom_violin(aes(colour = model_type)) +  
  facet_wrap(~site_id)

```

### Uncertainty representation

The uncertainty each model represents might be an interesting characteristic to investigate how skill varies?

```{r}
# types of uncertainty represented
# Number of models that contain each type
model_meta |> 
  summarise(across(contains('unc'), ~ sum(.x, na.rm = T)))

```

Driver uncertainty is the most common type included and initial condition uncertainty the least common. Most forecasts only include one type of uncertainty.

```{r}
# How many sources are represented in each model?
model_meta |> 
  select(model_id, contains('unc_')) |> 
  mutate(unc_n = rowSums(pick(where(is.numeric)))) |> 
  group_by(unc_n) |> 
  summarise(n = n())
```

Most forecasts only represent one source of uncertainty (usually driver), but a good number also include multiple sources of uncertainty.

### NOAA weather covariates

I've also looked at what weather covariates are used in each model from those available from NOAA. No model uses covaraite drivers other than NOAA weather forecasts.

```{r}
NOAA_vars <- c('air_temperature',
               'air_pressure',
               'relative_humidity',
               'surface_downwelling_longwave_flux_in_air',
               'surface_downwelling_shortwave_flux_in_air', 
               'precipitation_flux',
               'eastward_wind',
               'northward_wind')

NOAA_summary <- model_meta |> 
  select(model_id)

for (i in 1:length(NOAA_vars)) {
  detect_var <- NOAA_vars[i]

  NOAA_summary <- 
    NOAA_summary |> 
    mutate(detect_var = str_detect(model_meta$NOAA_var, detect_var))

  colnames(NOAA_summary)[1 + i] <- detect_var
}

```

Summarising the driver variables to see how many models use each of the variables. Air temperature being the most common driver variable followed by short wave radiation.

```{r}
NOAA_summary |> 
  select(-model_id) |> 
  summarise_all(~sum(.x, na.rm = T)) |> 
  pivot_longer(cols = everything(),
               names_to = 'NOAA_vars', 
               values_to = 'n_models') |> 
  arrange(desc(n_models))

```

We also see that many models use just this single driver variable, with the next common to use all 8 (in many of the complex process models).

```{r}
NOAA_summary |> 
  mutate(vars_n = rowSums(pick(where(is.logical)))) |> 
  group_by(vars_n) |> 
  summarise(n_models = n()) |> 
  arrange(desc(n_models))

```

In addition to looking at how each of these models performs and looking at relationships with the driver variable and number of drivers used, I also started to think about how this relates to the performance of the NOAA forecasts themselves. Does the skill of the aquatic forecasts mirror the driver variables that are included in the model?

### NOAA forecast performance

```{r eval = F}
NOAA_scores <- arrow::open_dataset(file.path(here::here(),"scores")) |> 
  filter(model_id == 'NOAA_daily',
         site_id %in% c('ARIK', 'BARC', 'BLDE'),
         variable %in% c('air_temperature', 'eastward_wind')) |> 
  collect()

NOAA_scores |> 
  mutate(horizon = as.numeric(datetime - as_date(reference_datetime))) |> 
  filter(horizon <= 30) |> 
  group_by(site_id, variable, horizon) |> 
  summarise(crps = mean(crps, na.rm = T)) |>
  ggplot(aes(x=horizon, y=crps, colour = site_id)) +
  geom_line() +
  facet_wrap(~variable)
```

## Things to think about

1.  Are all models equally represented? To make them comparable.
2.  Same with sites? Do we have the same number of days/models etc.?
3.  What to do about models that don't forecast all sites?

### 2023-11-21 01_get_model_list.R
add rescored

@@ -29,6 +29,6 @@ scores_s3 |>
   dplyr::filter(reference_datetime %in% forecast_dates,
                 variable %in% variables,
                 model_id %in% model_meta$model_id) |>
-  arrow::write_dataset("scores", partitioning=c("model_id", "site_id"))
+  arrow::write_dataset("scores_rescored", partitioning=c("model_id", "site_id"))
 
 #====================================================================#

### 2023-11-21 01a_inventory.R
noaa_var by variable

@@ -2,7 +2,8 @@ library(tidyverse)
 # inventory
 googlesheets4::gs4_deauth()
 model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing") |> 
-  mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0))
+  mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0)) |> 
+  filter(!is.na(model_type))
 
 # types of uncertainty represented
 # Number of models that contain each type
@@ -27,7 +28,7 @@ NOAA_vars <- c('air_temperature',
                'northward_wind')
 
 NOAA_summary <- model_meta |> 
-  select(model_id) 
+  select(model_id, variable) 
 
 
 for (i in 1:length(NOAA_vars)) {
@@ -37,17 +38,18 @@ for (i in 1:length(NOAA_vars)) {
     NOAA_summary |> 
     mutate(detect_var = str_detect(model_meta$NOAA_var, detect_var)) 
   
-  colnames(NOAA_summary)[1 + i] <- detect_var
+  colnames(NOAA_summary)[2 + i] <- detect_var
 }
 
 # Number of models that contain each variable
 NOAA_summary |> 
-  select(-model_id) |> 
+  select(-model_id) |>
+  group_by(variable) |> 
   summarise_all(~sum(.x, na.rm = T))
 
 NOAA_summary |> 
   mutate(vars_n = rowSums(pick(where(is.logical)))) |> 
-  group_by(vars_n) |> 
+  group_by(variable, vars_n) |> 
   summarise(n = n())
 
 # Aspects of DA
@@ -62,6 +64,10 @@ model_meta |>
 # is the forecast 'dynamic'?
 model_meta |> 
   summarise(dynamic = sum(dynamic, na.rm = T))
+
+# looking at drivers by variable:
+model_meta |> 
+  mutate(n_noaa = length(str_split_1(string = NOAA_var, pattern = ',')))
 
 # summary of submissions --------------------
 scores <- arrow::open_dataset('scores')

### 2023-11-21 03_score_NOAA.R
score vs obs

@@ -19,12 +19,10 @@ scored_forecasts <-
   mutate(reference_datetime = as_date(reference_datetime))
 
 # 3. Generate list of forecasts to be scored
-aquatic_sites <- arrow::open_dataset('scores') |>
-  distinct(site_id) |>
-  pull()
+lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
 
 site_dates <- expand.grid(reference_datetime = dates,
-                          site_id = aquatic_sites)
+                          site_id = lake_sites)
 
 to_score <- site_dates |> 
   anti_join(scored_forecasts) 
@@ -33,6 +31,7 @@ to_score <- site_dates |>
 
 to_score <- to_score |> 
   filter(!reference_datetime %in% as_date(c('2023-01-06', '2023-01-19', '2023-05-22')))
+
 # Loop through all site-date combos to score
 for (i in 1:nrow(to_score)) {
   # get the forecast
@@ -40,7 +39,8 @@ for (i in 1:nrow(to_score)) {
   site <- to_score$site_id[i]
 
   stage_2 <- neon4cast::noaa_stage2(start_date = forecast_date) |>
-    filter(site_id == site) |>
+    filter(site_id == site,
+           variable == 'air_temperature') |>
     collect()|> 
     mutate(datetime = as_date(datetime)) |> 
     group_by(parameter, reference_datetime, datetime, site_id, variable, family) |> 
@@ -49,7 +49,8 @@ for (i in 1:nrow(to_score)) {
   # get the observations
   stage_3 <- neon4cast::noaa_stage3() |>
     filter(site_id  == site,
-           datetime %in% unique(stage_2$datetime)) |>
+           datetime %in% unique(stage_2$datetime),
+           variable == 'air_temperature') |>
     collect() |>
 
     # compare with the mean prediction (as if it were the observation)
@@ -121,34 +122,48 @@ for (i in 1:nrow(to_score)) {
   obs_start <- format(forecast_date, "%Y-%m")
   obs_end <- format(forecast_date + days(35), "%Y-%m")
   
-  neon_obs <- loadByProduct(dpID="DP1.20046.001", 'BARC',
-                            startdate=obs_start, enddate=obs_end, 
-                            package="expanded", 
-                            token = Sys.getenv("NEON_TOKEN"),
-                            check.size = F,
-                            release = "current")
[…]

### 2023-11-21 02_chla_skills_scores.R
organise DO, chl scores

@@ -1,0 +1,85 @@
+
+library(tidyverse)
+library(arrow)
+library(lubridate)
+library(tidytext)
+
+lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
+cols_modeltype <- c('empirical' = "#482576FF",
+                    'multi-model ensemble' =  "#35608DFF",
+                    'process' = "#21908CFF",
+                    'ML' = '#BBDF27FF',
+                    'null' = '#43BF71FF')
+# read in the metadata from the google sheet
+googlesheets4::gs4_deauth()
+model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")
+
+# Open dataset-------------
+# calculate chla skill scores
+chla_scores <- arrow::open_dataset("scores") |>
+  filter(site_id %in% lake_sites,
+         # model_id != 'climatology',
+         model_id %in% model_meta$model_id,
+         variable == 'chla') |>
+  collect()
+
+# Calculate skill scores ------------
+chla_persistence <- arrow::open_dataset("scores") |>
+  filter(model_id == 'persistenceRW', 
+         site_id %in% lake_sites,
+         variable == 'chla') |>
+  collect() |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  arrange(reference_datetime, datetime, site_id) |> 
+  mutate(bias = observation - mean) |> 
+  pivot_wider(values_from = c(crps, logs),
+              names_from = model_id)
+
+chla_skill <- chla_scores |>
+  filter(model_id != 'persistenceRW') |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  full_join(chla_persistence) |> 
+  
+  # calcaulte the skill relative to climatology (- is bad, + is good)
+  mutate(skill_crps = crps_persistenceRW - crps,
+         skill_logs = logs_persistenceRW - logs,
+         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+  # consistent forecast period
+  filter(horizon <= 30,
+         horizon >= 1,
+         model_id != 'fARIMA') |> 
+  arrange(reference_datetime, site_id, datetime, model_id) 
[…]

### 2023-11-21 02_oxygen_skill_scores.R

@@ -1,0 +1,67 @@
+
+library(tidyverse)
+library(arrow)
+library(lubridate)
+library(tidytext)
+
+lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
+cols_modeltype <- c('empirical' = "#482576FF",
+                    'multi-model ensemble' =  "#35608DFF",
+                    'process' = "#21908CFF",
+                    'ML' = '#BBDF27FF',
+                    'null' = '#43BF71FF')
+# read in the metadata from the google sheet
+googlesheets4::gs4_deauth()
+model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")
+
+# Open dataset-------------
+# calculate DO skill scores
+DO_scores <- arrow::open_dataset("scores") |>
+  filter(site_id %in% lake_sites,
+         # model_id != 'climatology',
+         model_id %in% model_meta$model_id,
+         variable == 'oxygen') |>
+  collect()
+
+# Calculate skill scores ------------
+DO_climatology <- arrow::open_dataset("scores") |>
+  filter(model_id == 'climatology', 
+         site_id %in% lake_sites,
+         variable == 'oxygen') |>
+  collect() |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  arrange(reference_datetime, datetime, site_id) |> 
+  pivot_wider(values_from = c(crps, logs),
+              names_from = model_id)
+
+DO_skill <- DO_scores |>
+  filter(model_id != 'climatology') |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  full_join(DO_climatology) |> 
+  
+  # calcaulte the skill relative to climatology (- is bad, + is good)
+  mutate(skill_crps = crps_climatology - crps,
+         skill_logs = logs_climatology - logs,
+         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+  # consistent forecast period
+  filter(horizon <= 30,
+         horizon >= 1,
+         model_id != 'fARIMA') |> 
+  arrange(reference_datetime, site_id, datetime, model_id) 
+#------------------------------------#
+
+DO_skill |> 
[…]

### 2023-11-21 02_temperature_skill_scores.R

@@ -16,22 +16,16 @@ model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/
 
 # Open dataset-------------
 # start with lake temperatures
-lake_scores <- arrow::open_dataset("scores") |>
+temperature_scores <- arrow::open_dataset("scores") |>
   filter(site_id %in% lake_sites,
          # model_id != 'climatology',
          model_id %in% model_meta$model_id,
          variable == 'temperature') |>
   collect()
 
-stream_scores <- arrow::open_dataset("scores") |>
-  filter(!site_id %in% lake_sites,
-         # model_id != 'climatology',
-         model_id %in% model_meta$model_id,
-         variable == 'temperature') |>
-  collect()
 
 # Calculate skill scores ------------
-climatology_lake <- arrow::open_dataset("scores") |>
+temperature_climatology <- arrow::open_dataset("scores") |>
   filter(model_id == 'climatology', 
          site_id %in% lake_sites,
          variable == 'temperature') |>
@@ -41,10 +35,10 @@ climatology_lake <- arrow::open_dataset("scores") |>
   pivot_wider(values_from = c(crps, logs),
               names_from = model_id)
 
-lake_skill_scores <- lake_scores |>
+temperature_skill <- temperature_scores |>
   filter(model_id != 'climatology') |> 
   select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  full_join(climatology_lake) |> 
+  full_join(temperature_climatology) |> 
   
   # calcaulte the skill relative to climatology (- is bad, + is good)
   mutate(skill_crps = crps_climatology - crps,
@@ -56,39 +50,10 @@ lake_skill_scores <- lake_scores |>
          model_id != 'fARIMA') |> 
   arrange(reference_datetime, site_id, datetime, model_id) 
 
-
-# stream skill scores
-climatology_stream <- arrow::open_dataset("scores") |>
-  filter(model_id == 'climatology', 
-         !site_id %in% lake_sites,
-         variable == 'temperature') |>
-  collect() |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
[…]

### 2023-11-21 03_plot_NOAA.R
simple noaa plots

@@ -1,0 +1,23 @@
+library(tidyverse)
+library(arrow)
+library(lubridate)
+
+lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
+
+# Open dataset-------------
+# start with lake temperatures
+NOAA_scores <- arrow::open_dataset("scores") |>
+  filter(site_id %in% lake_sites,
+         model_id %in% c('NOAA_daily', 'NOAA_daily_obs')) |>
+  collect()
+
+NOAA_scores |> 
+  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
+  group_by(model_id, site_id, horizon) |> 
+  summarise(mean_crps = mean(crps, na.rm = T)) |> 
+  filter(horizon <= 30) |> 
+  ggplot(aes(x=horizon, y=mean_crps, colour = model_id)) +
+  geom_line() +
+  facet_wrap(~site_id) +
+  theme_bw() +
+  labs(y = 'CRPS score (oC)', title = 'NOAA air temperature forecast')

### 2023-11-21 Updates/2023-11-17 figures_draft.Rmd
rename

---
title: "figures?"
output: html_document
date: "2023-11-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(arrow)
library(lubridate)
library(tidytext)
```

## Read in the data and model metadata

```{r read-in-data}

lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
cols_modeltype <- c('empirical' = "#482576FF",
          'multi-model ensemble' =  "#35608DFF",
          'process' = "#21908CFF",
          'ML' = '#BBDF27FF',
          'null' = '#43BF71FF')
# read in the metadata from the google sheet
googlesheets4::gs4_deauth()
model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")

# Open dataset-------------
# start with lake temperatures
temperature_scores <- arrow::open_dataset("../scores_rescored") |>
  filter(site_id %in% lake_sites,
         # model_id != 'climatology',
         model_id %in% model_meta$model_id,
         variable == 'temperature') |>
  collect()

# Calculate skill scores ------------
temperature_climatology <- arrow::open_dataset("../scores_rescored") |>
  filter(model_id == 'climatology', 
         site_id %in% lake_sites,
         variable == 'temperature') |>
  collect() |> 
  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
  arrange(reference_datetime, datetime, site_id) |> 
  pivot_wider(values_from = c(crps, logs),
              names_from = model_id)

temperature_skill <- temperature_scores |>
  filter(model_id != 'climatology') |> 
  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
  full_join(temperature_climatology) |>

  # calcaulte the skill relative to climatology (- is bad, + is good)
  mutate(skill_crps = crps_climatology - crps,
         skill_logs = logs_climatology - logs,
         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  # consistent forecast period
  filter(horizon <= 30,
         horizon >= 1,
         model_id != 'fARIMA') |> 
  arrange(reference_datetime, site_id, datetime, model_id) 
```

## Figure 1

Model type impacts of water temperature forecast skill (relative to climatology). Individual model mean CRPS skill, coloured by model_type (as defined by submitters). 
Overall and then by site. 
```{r model-type, fig.width=10}
top5_temp_site <- temperature_skill |> 
  group_by(model_id, site_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  group_by(site_id) |> 
  slice_max(mean_crps, n= 5, na_rm = T) |> 
  mutate(model_id = fct_reorder(model_id, desc(mean_crps))) |>
  left_join(model_meta) |> 
  ggplot(aes(x=mean_crps, 
             y = reorder_within(x = model_id, within = site_id, by = mean_crps),
             fill = model_type)) +
  geom_bar(stat = 'identity') +
  scale_y_reordered() +
  scale_fill_manual(values = cols_modeltype) +
  facet_wrap(~site_id, scales = 'free_y',nrow = 4) +
  labs(y='model_id', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  theme_bw()

top10_temp <- temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  slice_max(mean_crps, n= 10, na_rm = T) |> 
  left_join(model_meta) |> 
  mutate(model_id = fct_reorder(model_id, mean_crps)) |>
  ggplot(aes(x=mean_crps, 
             y = model_id,
             fill = model_type)) +
  geom_bar(stat = 'identity') +
  scale_y_reordered() +
  labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  scale_fill_manual(values = cols_modeltype) +
  theme_bw()

ggpubr::ggarrange(top10_temp, top5_temp_site, nrow = 1, 
                  # widths = c(1,0.5), 
                  common.legend = T,
                  legend = 'top')

```

## Figure 2
How does model type performance, on avergae, differ by site. Either comparing sites by model type or vise versa. 
This uses all model_ids (maybe get rid of the 'worst') performers.

```{r model_type-site}
ggpubr::ggarrange(temperature_scores |> 
  left_join(model_meta, by = 'model_id') |> 
  group_by(model_type, model_id, site_id) |> 
  summarise(mean_crps = mean(crps, na.rm = T)) |>  
  ggplot(aes(x = mean_crps, 
             y = site_id)) +
  geom_boxplot() +
  facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
  labs(y='model_id', x = 'mean CRPS score') +
  theme_bw(),

  temperature_scores |> 
  left_join(model_meta, by = 'model_id') |> 
  group_by(model_type, model_id, site_id) |> 
  summarise(mean_crps = mean(crps, na.rm = T)) |>  
  ggplot(aes(x = mean_crps, 
             y = model_type)) +
  geom_boxplot() +
  facet_wrap(~site_id, scales = 'free_y',nrow = 4) +
  labs(y='model_id', x = 'mean CRPS score') +
  theme_bw(),
  nrow = 1, widths = c(1, 1.5))

```

## Figure 3
How does the performance vary across horizon? This is the top 10 models at each site (so this doesn't really look at model type...)

```{r horizon-score}
# top 10 average models
top_mods <- temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  slice_max(mean_crps, n = 10, na_rm = T) |> 
  distinct(model_id)

temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(site_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = site_id)) +
  geom_hline(yintercept = 0) + 
  geom_line()  +  
  labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
  theme_bw()

```

So maybe something more like this...
This is all sites together, so more like figure 1.

```{r NOAA-scores}
temperature_skill |> 
  filter(model_id %in% top_mods$model_id) |> 
  group_by(model_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = model_id)) +
  geom_hline(yintercept = 0) + 
  geom_line()  +  
  labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
  theme_bw()
```

## Figure 4 - inclusion of covariates

```{r}

temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  slice_max(mean_crps, n= 10, na_rm = T) |> 
  left_join(model_meta) |> 
  mutate(model_id = fct_reorder(model_id, mean_crps)) |>
  ggplot(aes(x=mean_crps, 
             y = model_id,
             fill = as.factor(uses_NOAA))) +
  geom_bar(stat = 'identity') +
  scale_y_reordered() +
  labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  scale_fill_viridis_d(option = 'inferno') +
  theme_bw()

temperature_skill |> 
  group_by(model_id, site_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  group_by(site_id) |> 
  slice_max(mean_crps, n= 5, na_rm = T) |> 
  mutate(model_id = fct_reorder(model_id, desc(mean_crps))) |>
  left_join(model_meta) |> 
  ggplot(aes(x=mean_crps, 
             y = reorder_within(x = model_id, within = site_id, by = mean_crps),
             fill = as.factor(uses_NOAA))) +
  geom_bar(stat = 'identity') +
  scale_y_reordered() +
  scale_fill_viridis_d(option = 'inferno') +
  facet_wrap(~site_id, scales = 'free_y',nrow = 4) +
  labs(y='model_id', x = 'mean CRPS skill (vs climatology)') +
  geom_vline(xintercept = 0) +
  theme_bw()
```

## Figure 5 - NOAA scores

can this be related to the weather forecasts?

```{r}
NOAA_scores <- arrow::open_dataset("../scores") |>
  filter(site_id %in% lake_sites,
         model_id %in% c('NOAA_daily', 'NOAA_daily_obs')) |>
  collect()

NOAA_scores |> 
  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
  group_by(model_id, site_id, horizon) |> 
  summarise(mean_crps = mean(crps, na.rm = T)) |> 
  filter(horizon <= 30) |> 
  ggplot(aes(x=horizon, y=mean_crps, colour = model_id)) +
  geom_line() +
  facet_wrap(~site_id) +
  theme_bw() +
  labs(y = 'CRPS score (oC)', title = 'NOAA air temperature forecast')

```

## site differences in observations?

```{r obs-cv}
temperature_scores |> 
  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  filter(between(horizon, 0,30)) |> 
  group_by(site_id, variable, horizon) |> 
  summarise(cv = sd(observation, na.rm = T)/ mean(observation, na.rm = T)) |> 
  ggplot(aes(x=horizon, y= cv, colour = site_id)) +
  geom_point() +
  scale_colour_viridis_d(option = 'turbo', begin = 0.1, end = 0.9) +
  theme_bw()
```

Model rankings
```{r eval = F}
temperature_scores |> 
  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
  filter(horizon < 30, horizon > 0) |> 
  select(reference_datetime, horizon, model_id, site_id, crps) |>
  arrange(reference_datetime, horizon, site_id, crps) |>
  group_by(reference_datetime, horizon, site_id) |>
  mutate(rank = row_number()) |>
  group_by(horizon, site_id, rank, model_id) |>
  summarise(proportion = (n()/length(unique(temperature_scores$reference_datetime))))

```

### 2023-11-22 R/create_all_mme.R
function to generate mme of all submitted models

@@ -1,0 +1,115 @@
+create_all_mme <- function(forecast_models = 'all', # vector of list of model names
+                           n = 200, # what size ensemble do you want?
+                           ensemble_name, # what is the name of the ensemble output
+                           forecast_date, # when is the forecast for (what forecast to grab)
+                           var, # which variable do you want a forecast for
+                           h = 30, # what is the required forecast horizon
+                           theme, # challenge theme 
+                           omit = NA,  # do you want to omit any?
+                           out_dir, # where to save
+                           out_format = 'csv') # file type
+  {
+  s3_theme <- s3_bucket(paste0('neon4cast-forecasts/parquet/', theme),
+                        endpoint_override= "data.ecoforecast.org")
+  
+  mods <- data.frame(model_id = s3_theme$ls(),
+                     present = NA) |> 
+    filter(!model_id %in% omit)
+  
+  message('Checking which models exist for this forecast date')
+  
+  for (i in 1:nrow(mods)) {
+    
+    s3_model <- s3_bucket(paste0('neon4cast-forecasts/parquet/', theme, '/',
+                                 mods$model_id[i]),
+                          endpoint_override= "data.ecoforecast.org")
+    
+    output <- ifelse(length(which(s3_model$ls() == paste0('reference_datetime=',forecast_date))) == 1, T, F)
+    
+    mods$present[i] <- output
+    
+  }
+  
+  # which models are present for this forecast date?
+  mods_present <- mods |> 
+    filter(present == T)
+  
+  # get all the forecasts that are present for this date and variable.
+  # function will generate an ensemble if it's a normal distribution
+  forecasts <- map_dfr(
+    .x = gsub(mods_present$model_id, pattern=  'model_id=', replacement = ''), 
+    .f = ~ get_forecast(model_id = .x, theme = theme, forecast_date = forecast_date))
+  
+  # How many from each forecast should be sampled
+  n_models <- distinct(forecasts, model_id, site_id) |> 
+    group_by(site_id) |> 
+    summarise(n_models = n()) |> 
+    mutate(sample = round(n / n_models, digits = 0))
+
+  # split into a list of dataframes
+  sitewise_mme <- split(forecasts, f = forecasts$site_id)
+
+  mme_forecast <- NULL
+  
[…]

### 2023-11-22 R/get_forecast.R
get an individual forecast for a model_id/reference_datetime

@@ -1,0 +1,53 @@
+
+get_forecast  <- function(theme, 
+                          model_id, 
+                          forecast_date) {
+  s3_model <- s3_bucket(file.path('neon4cast-forecasts/parquet', theme, 
+                                  paste0('model_id=', model_id), 
+                                  paste0('reference_datetime=', forecast_date)),
+                        endpoint_override= "data.ecoforecast.org")
+  
+  forecast <- arrow::open_dataset(s3_model) |>
+    collect() |> 
+    filter(variable == var, 
+           datetime > forecast_date) |>
+    group_by(site_id) |> 
+    # remove sites that contain NAs
+    filter(!any(is.na(prediction))) |> 
+    ungroup() |> 
+    mutate(model_id = model_id,
+           horizon = as_date(datetime) - as_date(forecast_date)) |> 
+    filter(horizon <= h) |> 
+    select(-horizon)
+  
+  if (nrow(forecast) == 0) {
+    message('No forecast for ', var, ' using ', model_id)
+    return(forecast)
+  } else {
+    # do some more data wrangling
+    message(model_id, ' read in')
+    
+    if (forecast$family[1] != 'sample') {
+      message('generate ensemble of 100')
+      
+      forecast <- forecast |> 
+        select(datetime, site_id, variable, family, parameter, prediction, model_id) |> 
+        pivot_wider(names_from = parameter,
+                    values_from = prediction, 
+                    id_cols = c(datetime, site_id, model_id)) |> 
+        
+        group_by(site_id, datetime, model_id) |> 
+        # sample from the distribution based on the mean and sd
+        summarise(prediction = rnorm(100, mean = mu, sd = sigma)) |> 
+        group_by(site_id, datetime) |> 
+        # parameter value needs to be character
+        mutate(parameter = as.character(row_number()),
+               # model_id = ensemble_name, 
+               reference_datetime = forecast_date,
+               variable = var,
+               family = 'ensemble')
+    }
+  }     
+  return(forecast)
+
+}

### 2023-11-22 AGU_analysis/model_means.R
script for model mean

@@ -1,0 +1,38 @@
+# AGU analysis - what does the mean look like?
+# take all submitted models (the spaghetti) and return the ensemble
+
+# Submit all the ensemble forecast to challenge!
+setwd(here::here())
+library(tidyverse)
+library(arrow)
+library(lubridate)
+source('./R/create_all_mme.R')
+source('./R/get_forecast.R')
+
+Sys.unsetenv("AWS_ACCESS_KEY_ID")
+Sys.unsetenv("AWS_SECRET_ACCESS_KEY")
+Sys.unsetenv("AWS_DEFAULT_REGION")
+Sys.unsetenv("AWS_S3_ENDPOINT")
+Sys.setenv(AWS_EC2_METADATA_DISABLED="TRUE")
+
+# where are the raw forecasts
+s3 <- "neon4cast-forecasts/parquet/aquatics/"
+
+forecast_dates <- seq(as_date('2023-01-01'), Sys.Date(), 1)
+
+
+for (i in 1:length(forecast_dates)) {
+  date <- forecast_dates[1]
+  
+  mme_file <- create_all_mme(forecast_models = 'all',
+                             ensemble_name = 'all_submissions',
+                             forecast_date = date,
+                             var = 'temperature', 
+                             h = 30, 
+                             theme = 'aquatics',
+                             n = 200, 
+                             omit = 'tg_auto_adam', 
+                             out_dir = 'Forecasts/parquet',
+                             output = 'parquet')
+  
+}

### 2023-11-22 Updates/2023-11-17 figures_draft.Rmd

@@
 library(tidytext)
+library(ggpattern)
 ```
@@
 
-Model type impacts of water temperature forecast skill (relative to climatology). Individual model mean CRPS skill, coloured by model_type (as defined by submitters). 
-Overall and then by site. 
+Top 10 models and all models coloured by type and showing if they include covariates or not. 
 ```{r model-type, fig.width=10}
-top5_temp_site <- temperature_skill |> 
-  group_by(model_id, site_id) |> 
-  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
-  group_by(site_id) |> 
-  slice_max(mean_crps, n= 5, na_rm = T) |> 
-  mutate(model_id = fct_reorder(model_id, desc(mean_crps))) |>
-  left_join(model_meta) |> 
-  ggplot(aes(x=mean_crps, 
-             y = reorder_within(x = model_id, within = site_id, by = mean_crps),
-             fill = model_type)) +
-  geom_bar(stat = 'identity') +
-  scale_y_reordered() +
-  scale_fill_manual(values = cols_modeltype) +
-  facet_wrap(~site_id, scales = 'free_y',nrow = 4) +
-  labs(y='model_id', x = 'mean CRPS skill (vs climatology)') +
-  geom_vline(xintercept = 0) +
-  theme_bw()
-
-top10_temp <- temperature_skill |> 
+top10_temp <- 
+  temperature_skill |> 
   group_by(model_id) |> 
@@
   mutate(model_id = fct_reorder(model_id, mean_crps)) |> 
-  ggplot(aes(x=mean_crps, 
-             y = model_id,
-             fill = model_type)) +
-  geom_bar(stat = 'identity') +
+  ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
+  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
+                   stat = "identity",
+                   alpha = 0.8, 
+                   pattern_fill = "black",
+                   colour = "black", 
+                   pattern_spacing = 0.02,
+                   pattern_alpha = 0.5) + 
   scale_y_reordered() +
@@
   scale_fill_manual(values = cols_modeltype) +
+  scale_pattern_manual(values=c('none', 'stripe')) +
+  scale_pattern_type_manual(values=c(NA, NA)) +
   theme_bw()
 
-ggpubr::ggarrange(top10_temp, top5_temp_site, nrow = 1, 
-                  # widths = c(1,0.5), 
+all_temperature <- temperature_skill |> 
+  group_by(model_id) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  left_join(model_meta) |> 
+  mutate(model_id = fct_reorder(model_id, mean_crps)) |> 
+  ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
+  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
+                   stat = "identity",
+                   alpha = 0.8, 
+                   pattern_fill = "black",
+                   colour = "black", 
+                   pattern_spacing = 0.02,
+                   pattern_alpha = 0.8) + 
+  scale_y_reordered() +
+  labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
+  geom_vline(xintercept = 0) +
+  scale_fill_manual(values = cols_modeltype) +
+  scale_pattern_manual(values=c('none', 'stripe')) +
+  scale_pattern_type_manual(values=c(NA, NA)) +
+  theme_bw()
+
+ggpubr::ggarrange(top10_temp, all_temperature, nrow = 1, 
                   common.legend = T,
-                  legend = 'top')
+                  legend = 'top', labels = 'auto')
 
@@
 
+9 of 10 top 10 models use covariates, and process models and ML are the most common best performers. 
+There are a lot of really bad empirical models.
 ## Figure 2
-How does model type performance, on avergae, differ by site. Either comparing sites by model type or vise versa. 
+How does model type performance, on avergae, differ by site.
+
 This uses all model_ids (maybe get rid of the 'worst') performers.
@@
 ```{r model_type-site}
-ggpubr::ggarrange(temperature_scores |> 
+temperature_skill |> 
   left_join(model_meta, by = 'model_id') |>
+  filter(model_type != 'null') |> 
   group_by(model_type, model_id, site_id) |> 
-  summarise(mean_crps = mean(crps, na.rm = T)) |>  
-  ggplot(aes(x = mean_crps, 
-             y = site_id)) +
+  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
+  ggplot(aes(x = mean_skill, 
+             y = site_id, 
+             colour = model_type)) +
+  geom_vline(xintercept = 0, linetype = 'dashed') +
   geom_boxplot() +
@@
   labs(y='model_id', x = 'mean CRPS score') +
-  theme_bw(),
+  scale_colour_manual(values = cols_modeltype) +
+  theme_bw()
 
-  temperature_scores |> 
-  left_join(model_meta, by = 'model_id') |> 
-  group_by(model_type, model_id, site_id) |> 
-  summarise(mean_crps = mean(crps, na.rm = T)) |>  
-  ggplot(aes(x = mean_crps, 
-             y = model_type)) +
-  geom_boxplot() +
-  facet_wrap(~site_id, scales = 'free_y',nrow = 4) +
-  labs(y='model_id', x = 'mean CRPS score') +
-  theme_bw(),
-  nrow = 1, widths = c(1, 1.5))
-
+temperature_climatology |> 
+  group_by(site_id) |> 
+  summarise(mean_score = mean(crps_climatology, na.rm = T))  |> 
+  arrange(mean_score)
 ```
 
+Gains at PRLA and PRPO relative to climatology (these have the worst climatology - see table), especially using process-models and ML.
+
+The best climatology sites (BARC, SUGG) are difficult to beat even though perform well overall.
+
 ## Figure 3
-How does the performance vary across horizon? This is the top 10 models at each site (so this doesn't really look at model type...)
+How does the performance vary across horizon? 
+Do the models perform better consistently across the horizon at each site?
+How do the best performing individual models perform?
 
@@
 
-temperature_skill |> 
+A <- temperature_skill |> 
   filter(model_id %in% top_mods$model_id) |> 
@@
   labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
+  scale_color_viridis_d(option = 'H') +
   theme_bw()
 
-```
-
-So maybe something more like this...
-This is all sites together, so more like figure 1.
-
-```{r NOAA-scores}
-temperature_skill |> 
+B <- temperature_skill |> 
   filter(model_id %in% top_mods$model_id) |> 
@@
   labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
-  theme_bw()
-```
-
-## Figure 4 - inclusion of covariates
-
-```{r}
-
-temperature_skill |> 
-  group_by(model_id) |> 
-  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
-  ungroup() |> 
-  slice_max(mean_crps, n= 10, na_rm = T) |> 
-  left_join(model_meta) |> 
-  mutate(model_id = fct_reorder(model_id, mean_crps)) |>
-  ggplot(aes(x=mean_crps, 
-             y = model_id,
-             fill = as.factor(uses_NOAA))) +
-  geom_bar(stat = 'identity') +
-  scale_y_reordered() +
-  labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
-  geom_vline(xintercept = 0) +
-  scale_fill_viridis_d(option = 'inferno') +
+  scale_color_viridis_d(option = 'A',) +
   theme_bw()
 
-temperature_skill |> 
-  group_by(model_id, site_id) |> 
-  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
-  group_by(site_id) |> 
-  slice_max(mean_crps, n= 5, na_rm = T) |> 
-  mutate(model_id = fct_reorder(model_id, desc(mean_crps))) |>
-  left_join(model_meta) |> 
-  ggplot(aes(x=mean_crps, 
-             y = reorder_within(x = model_id, within = site_id, by = mean_crps),
-             fill = as.factor(uses_NOAA))) +
-  geom_bar(stat = 'identity') +
-  scale_y_reordered() +
-  scale_fill_viridis_d(option = 'inferno') +
-  facet_wrap(~site_id, scales = 'free_y',nrow = 4) +
-  labs(y='model_id', x = 'mean CRPS skill (vs climatology)') +
-  geom_vline(xintercept = 0) +
-  theme_bw()
+ggpubr::ggarrange(A, B, labels = 'auto', align = 'v', nrow = 2)
 ```
+Models beeating climatology at the shorter horizons (as expected) PRLA and PRPO always better than climatology.
 
-## Figure 5 - NOAA scores
+Individual model performance decreases relative to climatology mostly except for random forest and lasso regression which improve after short horizons (may be related to the precision rather than accuracy).
+Less degradation in performance of ML models across the horizon (xgboost and random forest) than process models.
+
+## Figure 4 - NOAA scores
 
 can this be related to the weather forecasts?
+
+There is only one month when the observations overlap for the observational data...
+But still the performance is better vs the observations than the stage 3
 
@@
   mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
+  select(site_id, crps, horizon, reference_datetime, model_id) |> 
+  pivot_wider(names_from = site_id, values_from = crps, id_cols = horizon:model_id) |> 
+  filter(reference_datetime > as_date('2023-05-25'),
+         reference_datetime < as_date('2023-07-31')) |> 
+  # only a 1 month overlap....
+  pivot_longer(cols = BARC:TOOK, names_to = 'site_id', values_to = 'crps') |> 
   group_by(model_id, site_id, horizon) |> 
@@
   geom_point() +
-  scale_colour_viridis_d(option = 'turbo', begin = 0.1, end = 0.9) +
+  scale_colour_viridis_d(option = 'turbo') +
   theme_bw()
@@
 
-Model rankings
-```{r eval = F}
-temperature_scores |> 
-  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
-  filter(horizon < 30, horizon > 0) |> 
-  select(reference_datetime, horizon, model_id, site_id, crps) |>
-  arrange(reference_datetime, horizon, site_id, crps) |>
-  group_by(reference_datetime, horizon, site_id) |>
-  mutate(rank = row_number()) |>
-  group_by(horizon, site_id, rank, model_id) |>
-  summarise(proportion = (n()/length(unique(temperature_scores$reference_datetime))))
+# Supplementary figures
 
+```{r echo = F}
+temperature_skill |> 
+  left_join(model_meta, by = 'model_id') |>
+  filter(model_type != 'null') |> 
+  group_by(model_type, model_id, site_id) |> 
+  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
+  ggplot(aes(x = mean_skill, 
+             y = site_id, 
+             colour = model_type)) +
+  geom_vline(xintercept = 0, linetype = 'dashed') +
+  geom_boxplot() +
+  facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
+  labs(y='model_id', x = 'mean CRPS score', title = 'all dates') +
+  scale_colour_manual(values = cols_modeltype) +
+  theme_bw()
+
+temperature_skill |> 
+  left_join(model_meta, by = 'model_id') |>
+  filter(model_type != 'null',
+         reference_datetime > as_date('2023-06-28'),
+         reference_datetime < as_date('2023-09-18')) |> 
+  group_by(model_type, model_id, site_id) |> 
+  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
+  ggplot(aes(x = mean_skill, 
+             y = site_id, 
+             colour = model_type)) +
+  geom_vline(xintercept = 0, linetype = 'dashed') +
+  geom_boxplot() +
+  facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
+  labs(y='model_id', x = 'mean CRPS score', title = 'overlapping only') +
+  scale_colour_manual(values = cols_modeltype) +
+  theme_bw() 
 ```

### 2023-11-22 AGU_analysis/model_means.R
only do missing forecasts

@@ -20,10 +20,31 @@ s3 <- "neon4cast-forecasts/parquet/aquatics/"
 
 forecast_dates <- seq(as_date('2023-01-01'), Sys.Date(), 1)
 
+# look to see what forecasts have already been generated
+out_format <- 'csv'
+out_dir = 'Forecasts'
+
+if (out_format == 'parquet') {
+  done <- open_dataset(out_dir) |> 
+    distinct(reference_datetime) |> 
+    collect()
+} else {
+  done_files <- list.files(out_dir, pattern = out_format)
+  dates <- map_chr(.x = done_files, ~paste(str_split_1(string = .x, pattern = '-')[2],
+                                           str_split_1(string = .x, pattern = '-')[3], 
+                                           str_split_1(string = .x, pattern = '-')[4], sep = '-')
+  )
+  done <- data.frame(reference_datetime = dates)               
+  
+}
+
+# skips ones already done
+forecast_dates <- forecast_dates[-which(as.character(forecast_dates) %in% done$reference_datetime)]
 
 for (i in 1:length(forecast_dates)) {
-  date <- forecast_dates[1]
+  date <- forecast_dates[i]
   
+  # generate an MME using all 'available' forecasts
   mme_file <- create_all_mme(forecast_models = 'all',
                              ensemble_name = 'all_submissions',
                              forecast_date = date,
@@ -32,7 +53,116 @@ for (i in 1:length(forecast_dates)) {
                              theme = 'aquatics',
                              n = 200, 
                              omit = 'tg_auto_adam', 
-                             out_dir = 'Forecasts/parquet',
-                             output = 'parquet')
-  
+                             out_dir = out_dir,
+                             out_format = out_format)
+  if (length(mme_file) == 0) {
+    message(date, ' failed')
+  }
 }
+#-------------------------------#
+
+#--------- Scoring --------------
+# source('R/scoring_function_arrow.R')
+# 
+# local_path <- 'scores'
+# 
+# if (!dir.exists(local_path)) {
+#   dir.create(local_path, recursive = T)
+# }
+# 
+# 
+# #### Read in forecasts ####
+# # Open the dataset of forecast parquets to be scored
+# forecast_parquets <- './Forecasts/parquet'
+# open_parquets <- arrow::open_dataset(forecast_parquets)
+# 
+# 
+# # forecasts to scores
[…]

### 2023-11-22 R/create_all_mme.R
edits

@@ -14,10 +14,9 @@ create_all_mme <- function(forecast_models = 'all', # vector of list of model na
   
   mods <- data.frame(model_id = s3_theme$ls(),
                      present = NA) |> 
-    filter(!model_id %in% omit)
+    filter(!model_id %in% paste0('model_id=', omit))
   
-  message('Checking which models exist for this forecast date')
-  
+  message(paste0('---- Checking which models exist for ', forecast_date,'  -----'))
   for (i in 1:nrow(mods)) {
     
     s3_model <- s3_bucket(paste0('neon4cast-forecasts/parquet/', theme, '/',
@@ -38,7 +37,7 @@ create_all_mme <- function(forecast_models = 'all', # vector of list of model na
   # function will generate an ensemble if it's a normal distribution
   forecasts <- map_dfr(
     .x = gsub(mods_present$model_id, pattern=  'model_id=', replacement = ''), 
-    .f = ~ get_forecast(model_id = .x, theme = theme, forecast_date = forecast_date))
+    .f = ~ get_forecast(model_id = .x, theme = theme, forecast_date = forecast_date, var = var, h = h))
   
   # How many from each forecast should be sampled
   n_models <- distinct(forecasts, model_id, site_id) |> 
@@ -97,7 +96,9 @@ create_all_mme <- function(forecast_models = 'all', # vector of list of model na
       file.remove(file.path('./Forecasts/ensembles', filename))
       message('forecast not valid')
     } else {
+      message('forecast saved')
       return(filename)
+      
     }
   }
   
@@ -106,7 +107,11 @@ create_all_mme <- function(forecast_models = 'all', # vector of list of model na
       select(-any_of(c('pubDate'))) |> 
       mutate(date = as_date(datetime)) |> 
       group_by(model_id, reference_datetime, date) |> 
-      arrow::write_parquet(out_dir)
+      arrow::write_dataset(path = out_dir, format = 'parquet')
+    message('forecast saved')
+    return(T)
+    
+    
   }
   
   

### 2023-11-22 R/get_forecast.R
add extra args

@@ -1,7 +1,9 @@
 
 get_forecast  <- function(theme, 
                           model_id, 
-                          forecast_date) {
+                          forecast_date, 
+                          var, 
+                          h = 30) {
   s3_model <- s3_bucket(file.path('neon4cast-forecasts/parquet', theme, 
                                   paste0('model_id=', model_id), 
                                   paste0('reference_datetime=', forecast_date)),
@@ -38,7 +40,7 @@ get_forecast  <- function(theme,
         
         group_by(site_id, datetime, model_id) |> 
         # sample from the distribution based on the mean and sd
-        summarise(prediction = rnorm(100, mean = mu, sd = sigma)) |> 
+        reframe(prediction = rnorm(100, mean = mu, sd = sigma)) |> 
         group_by(site_id, datetime) |> 
         # parameter value needs to be character
         mutate(parameter = as.character(row_number()),

### 2023-11-24 AGU_analysis/model_means.R
score parquets

@@ -21,8 +21,8 @@ s3 <- "neon4cast-forecasts/parquet/aquatics/"
 forecast_dates <- seq(as_date('2023-01-01'), Sys.Date(), 1)
 
 # look to see what forecasts have already been generated
-out_format <- 'csv'
-out_dir = 'Forecasts'
+out_format <- 'parquet'
+out_dir = 'Forecasts/parquet'
 
 if (out_format == 'parquet') {
   done <- open_dataset(out_dir) |> 
@@ -62,107 +62,56 @@ for (i in 1:length(forecast_dates)) {
 #-------------------------------#
 
 #--------- Scoring --------------
-# source('R/scoring_function_arrow.R')
-# 
-# local_path <- 'scores'
-# 
-# if (!dir.exists(local_path)) {
-#   dir.create(local_path, recursive = T)
-# }
-# 
-# 
-# #### Read in forecasts ####
-# # Open the dataset of forecast parquets to be scored
-# forecast_parquets <- './Forecasts/parquet'
-# open_parquets <- arrow::open_dataset(forecast_parquets)
-# 
-# 
-# # forecasts to scores
-# models <- 'all_submissions'
-# 
-# # vector of unique reference_datetimes
-# unique_reftime <- open_parquets |>
-#   distinct(reference_datetime) |>
-#   collect() |>
-#   arrange(reference_datetime) |>
-#   pull(reference_datetime)
-# 
-# # get all combinations of model and reference_datetime
-# to_score <- expand.grid(model_id = models, reference_datetime = unique_reftime)
-# 
-# #==================================#
-# 
-# #### Regular scoring function ####
-# for (i in 1:nrow(to_score)) {
-#   # subset the model and reference datetime
-#   forecast_df <- open_parquets|>
-#     dplyr::filter(model_id == to_score$model_id[i],
-#                   reference_datetime == to_score$reference_datetime[i]) |>
-#     collect()
-#   
-#   if (nrow(forecast_df) != 0) {
-#     # Score the forecast
-#     generate_forecast_score_arrow(targets_file = 'https://data.ecoforecast.org/neon4cast-targets/aquatics/aquatics-targets.csv.gz',
-#                                   forecast_df = forecast_df,
-#                                   local_directory = local_path)
-#     message(i, "/", nrow(to_score), " forecasts scored")
-#     
-#     
-#   } else {
[…]

### 2023-11-24 Updates/2023-11-17 figures_draft.Rmd
update patter

@@
                    colour = "black", 
-                   pattern_spacing = 0.02,
-                   pattern_alpha = 0.5) + 
+                   pattern_spacing = 0.04,
+                   pattern_alpha = 1) + 
   scale_y_reordered() +
@@
   scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw()
+  theme_bw()  + 
+    guides(pattern = guide_legend(override.aes = list(fill = "white")),
+           fill = guide_legend(override.aes = list(pattern = "none")))
 
-all_temperature <- temperature_skill |> 
+all_temp <- temperature_skill |> 
     group_by(model_id) |> 
@@
                      colour = "black", 
-                   pattern_spacing = 0.02,
-                   pattern_alpha = 0.8) + 
+                     pattern_spacing = 0.03,
+                     pattern_alpha = 1) + 
     scale_y_reordered() +
@@
     scale_fill_manual(values = cols_modeltype) +
-  scale_pattern_manual(values=c('none', 'stripe')) +
-  scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw()
+    scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses weather covariates?') +
+    scale_pattern_type_manual(values=c(NA, NA), name = 'Uses weather covariates?') +
+    theme_bw() + 
+    guides(pattern = guide_legend(override.aes = list(fill = "white")),
+           fill = guide_legend(override.aes = list(pattern = "none")))
 
-ggpubr::ggarrange(top10_temp, all_temperature, nrow = 1, 
+ggpubr::ggarrange(top10_temp, all_temp, nrow = 1, 
                   common.legend = T,

### 2023-11-24 Updates/2023-11-17 figures_draft.Rmd
plot bias and sd

--- 2023-11-22 Updates/2023-11-17 figures_draft.Rmd
@@
   left_join(model_meta) |> 
-  mutate(model_id = fct_reorder(model_id, mean_crps)) |> 
+  mutate(model_id = fct_reorder(model_id, mean_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
   ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
@@
   scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw()
+  theme_bw() + 
+    guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
+                                  title = 'Uses weather covariates?'),
+           fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
 
-all_temperature <- temperature_skill |> 
+all_temp <- temperature_skill |> 
     group_by(model_id) |> 
@@
     left_join(model_meta) |> 
-  mutate(model_id = fct_reorder(model_id, mean_crps)) |> 
+    mutate(model_id = fct_reorder(model_id, mean_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
     ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
@@
     scale_fill_manual(values = cols_modeltype) +
-  scale_pattern_manual(values=c('none', 'stripe')) +
-  scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw()
+    scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses weather covariates?') +
+    scale_pattern_type_manual(values=c(NA, NA), name = 'Uses weather covariates?') +
+    theme_bw() + 
+    guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses weather covariates?'),
+           fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
 
-ggpubr::ggarrange(top10_temp, all_temperature, nrow = 1, 
+ggpubr::ggarrange(top10_temp, all_temp, nrow = 1, 
                   common.legend = T,
@@
 There are a lot of really bad empirical models.
+
 ## Figure 2
@@
   labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
-  scale_color_viridis_d(option = 'A',) +
+  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
   theme_bw()
@@
 ```
+
 Models beeating climatology at the shorter horizons (as expected) PRLA and PRPO always better than climatology.
@@
 Less degradation in performance of ML models across the horizon (xgboost and random forest) than process models.
+
+### How much is uncertainty?
+
+How does the sd vary - is this my some see a increase in crps skill?
+
+```{r horizon-score}
+# top 10 average models
+top_mods <- temperature_skill |> 
+  group_by(model_id) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  ungroup() |> 
+  slice_max(mean_crps, n = 10, na_rm = T) |> 
+  distinct(model_id)
+
+sd <- temperature_scores |> 
+  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
+  filter(model_id %in% top_mods$model_id, 
+         horizon > 0, 
+         horizon <=30) |>
+  group_by(model_id, horizon) |> 
+  summarise(mean_sd = mean(sd, na.rm = T)) |> 
+  ggplot(aes(x=horizon, 
+             y = mean_sd, colour = model_id)) +
+  geom_line()  +  
+  labs(y='mean standard deviation', x = 'horizon (days)') +
+  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
+  theme_bw()
+
+bias <- temperature_scores |> 
+  mutate(horizon = as_date(datetime) - as_date(reference_datetime),
+         abs_bias = abs(mean - observation)) |> 
+  filter(model_id %in% top_mods$model_id, 
+         horizon > 0, 
+         horizon <=30) |>
+  group_by(model_id, horizon) |> 
+  summarise(mean_abs_bias = mean(abs_bias, na.rm = T)) |> 
+  ggplot(aes(x=horizon, 
+             y = mean_abs_bias, colour = model_id)) +
+  geom_line()  +  
+  labs(y='mean absolute bias (oC)', x = 'horizon (days)') +
+  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
+  theme_bw() +
+  coord_cartesian(ylim = c(0, 2.5))
+
+ggpubr::ggarrange(sd, bias, labels = 'auto', align = 'v', nrow = 2)
+```
 

### 2023-11-24 R/get_forecast.R
remove pubdate column

@@ -20,7 +20,7 @@ get_forecast  <- function(theme,
     mutate(model_id = model_id,
            horizon = as_date(datetime) - as_date(forecast_date)) |> 
     filter(horizon <= h) |> 
-    select(-horizon)
+    select(-any_of(c('horizon', 'pubDate')))
   
   if (nrow(forecast) == 0) {
     message('No forecast for ', var, ' using ', model_id)

### 2023-11-24 Updates/2023-11-17 figures_draft.Rmd
plot DO and chlorophyll

@@
 
-# Supplementary figures
+## Supplementary figures
 
@@
 ```
+
+## Chlorophyll and oxygen patterns same?
+
+```{r read-WQ, echo = F}
+# Open dataset-------------
+chla_scores <- arrow::open_dataset("../scores_rescored") |>
+  filter(site_id %in% lake_sites,
+         # model_id != 'climatology',
+         model_id %in% model_meta$model_id,
+         variable == 'chla') |>
+  collect()
+
+# Calculate skill scores ------------
+chla_persistence <- arrow::open_dataset("../scores_rescored") |>
+  filter(model_id == 'persistenceRW', 
+         site_id %in% lake_sites,
+         variable == 'chla') |>
+  collect() |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  arrange(reference_datetime, datetime, site_id) |> 
+  pivot_wider(values_from = c(crps, logs),
+              names_from = model_id)
+
+chla_skill <- chla_scores |>
+  filter(model_id != 'persistenceRW') |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  full_join(chla_persistence) |>
+
+  # calcaulte the skill relative to persistence (- is bad, + is good)
+  mutate(skill_crps = crps_persistenceRW - crps,
+         skill_logs = logs_persistenceRW - logs,
+         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+  # consistent forecast period
+  filter(horizon <= 30,
+         horizon >= 1,
+         model_id != 'fARIMA') |> 
+  arrange(reference_datetime, site_id, datetime, model_id)
+
+#--------------------------------------#
+# Dissolved oxygen
+
+# Open dataset-------------
+oxygen_scores <- arrow::open_dataset("../scores_rescored") |>
+  filter(site_id %in% lake_sites,
+         # model_id != 'climatology',
+         model_id %in% model_meta$model_id,
+         variable == 'oxygen') |>
+  collect()
+
+# Calculate skill scores ------------
+oxygen_climatology <- arrow::open_dataset("../scores_rescored") |>
+  filter(model_id == 'climatology', 
+         site_id %in% lake_sites,
+         variable == 'oxygen') |>
+  collect() |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  arrange(reference_datetime, datetime, site_id) |> 
+  pivot_wider(values_from = c(crps, logs),
+              names_from = model_id)
+
+oxygen_skill <- oxygen_scores |>
+  filter(model_id != 'climatology') |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  full_join(oxygen_climatology) |>
+
+  # calcaulte the skill relative to persistence (- is bad, + is good)
+  mutate(skill_crps = crps_climatology - crps,
+         skill_logs = logs_climatology - logs,
+         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+  # consistent forecast period
+  filter(horizon <= 30,
+         horizon >= 1,
+         model_id != 'fARIMA') |> 
+  arrange(reference_datetime, site_id, datetime, model_id)
+
+#--------------------------------------#
+```
+
+```{r plot-WQ, echo = F}
+
+top10_chla <- 
+  chla_skill |> 
+  group_by(model_id) |> 
+  summarise(median_crps = median(skill_crps, na.rm = T)) |> 
+  ungroup() |>
+  slice_max(median_crps, n= 10, na_rm = T) |>
+  left_join(model_meta) |> 
+  mutate(model_id = fct_reorder(model_id, median_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
+  ggplot(aes(x=median_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
+  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
+                   stat = "identity",
+                   alpha = 0.8, 
+                   pattern_fill = "black",
+                   colour = "black", 
+                   pattern_spacing = 0.04,
+                   pattern_alpha = 1) + 
+  scale_y_reordered() +
+  labs(y='Model ID', x = 'median CRPS skill (vs climatology)', title = 'chla') +
+  geom_vline(xintercept = 0) +
+  scale_fill_manual(values = cols_modeltype) +
+  scale_pattern_manual(values=c('none', 'stripe')) +
+  scale_pattern_type_manual(values=c(NA, NA)) +
+  theme_bw() + 
+  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
+                                title = 'Uses weather covariates?'),
+         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
+
+all_chla <- chla_skill |> 
+  group_by(model_id) |> 
+  summarise(median_crps = median(skill_crps, na.rm = T)) |> 
+  left_join(model_meta) |> 
+  mutate(model_id = fct_reorder(model_id, median_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
+  ggplot(aes(x=median_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
+  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
+                   stat = "identity",
+                   alpha = 0.8, 
+                   pattern_fill = "black",
+                   colour = "black", 
+                   pattern_spacing = 0.03,
+                   pattern_alpha = 1) + 
+  scale_y_reordered() +
+  labs(y='Model ID', x = 'median CRPS skill (vs persistence)') +
+  geom_vline(xintercept = 0) +
+  scale_fill_manual(values = cols_modeltype) +
+  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses weather covariates?') +
+  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses weather covariates?') +
+  theme_bw() + 
+  guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses weather covariates?'),
+         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
+
+ggpubr::ggarrange(top10_chla, all_chla, nrow = 1, 
+                  common.legend = T,align = 'h', 
+                  legend = 'top', labels = 'auto')
+
+#---------------------------------------------------#
+
+top10_oxygen <- 
+  oxygen_skill |> 
+  group_by(model_id) |> 
+  summarise(median_crps = median(skill_crps, na.rm = T)) |> 
+  ungroup() |>
+  slice_max(median_crps, n= 10, na_rm = T) |>
+  left_join(model_meta) |> 
+  mutate(model_id = fct_reorder(model_id, median_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
+  ggplot(aes(x=median_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
+  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
+                   stat = "identity",
+                   alpha = 0.8, 
+                   pattern_fill = "black",
+                   colour = "black", 
+                   pattern_spacing = 0.04,
+                   pattern_alpha = 1) + 
+  scale_y_reordered() +
+  labs(y='Model ID', x = 'median CRPS skill (vs climatology)', title = 'oxygen') +
+  geom_vline(xintercept = 0) +
+  scale_fill_manual(values = cols_modeltype) +
+  scale_pattern_manual(values=c('none', 'stripe')) +
+  scale_pattern_type_manual(values=c(NA, NA)) +
+  theme_bw() + 
+  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
+                                title = 'Uses weather covariates?'),
+         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
+
+all_oxygen <- oxygen_skill |> 
+  group_by(model_id) |> 
+  summarise(median_crps = median(skill_crps, na.rm = T)) |> 
+  left_join(model_meta) |> 
+  mutate(model_id = fct_reorder(model_id, median_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
+  ggplot(aes(x=median_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
+  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
+                   stat = "identity",
+                   alpha = 0.8, 
+                   pattern_fill = "black",
+                   colour = "black", 
+                   pattern_spacing = 0.03,
+                   pattern_alpha = 1) + 
+  scale_y_reordered() +
+  labs(y='Model ID', x = 'median CRPS skill (vs climatology)') +
+  geom_vline(xintercept = 0) +
+  scale_fill_manual(values = cols_modeltype) +
+  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses weather covariates?') +
+  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses weather covariates?') +
+  theme_bw() + 
+  guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses weather covariates?'),
+         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
+
+ggpubr::ggarrange(top10_oxygen, all_oxygen, nrow = 1, 
+                  common.legend = T, align = 'h', 
+                  legend = 'top', labels = 'auto')
+```

### 2023-11-24 02_chla_skills_scores.R
plot top models by type and covariates

@@ -15,7 +15,6 @@ googlesheets4::gs4_deauth()
 model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")
 
 # Open dataset-------------
-# calculate chla skill scores
 chla_scores <- arrow::open_dataset("scores") |>
   filter(site_id %in% lake_sites,
          # model_id != 'climatology',
@@ -31,7 +30,6 @@ chla_persistence <- arrow::open_dataset("scores") |>
   collect() |> 
   select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
   arrange(reference_datetime, datetime, site_id) |> 
-  mutate(bias = observation - mean) |> 
   pivot_wider(values_from = c(crps, logs),
               names_from = model_id)
 
@@ -68,18 +66,58 @@ chla_skill |>
   scale_fill_manual(values = cols_modeltype) +
   theme_bw() 
 
-chla_skill |> 
-  group_by(model_id, site_id) |> 
+top10_chla <- 
+  chla_skill |> 
+  group_by(model_id) |> 
   summarise(median_crps = median(skill_crps, na.rm = T)) |> 
-  group_by(site_id) |> 
-  slice_max(median_crps, n= 5, na_rm = T) |> 
-  mutate(model_id = fct_reorder(model_id, desc(median_crps))) |>
+  ungroup() |>
+  slice_max(median_crps, n= 10, na_rm = T) |>
   left_join(model_meta) |> 
-  ggplot(aes(x=median_crps, 
-             y = reorder_within(x = model_id, within = site_id, by = median_crps),
-             fill = model_type)) +
-  geom_bar(stat = 'identity') +
+  mutate(model_id = fct_reorder(model_id, median_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
+  ggplot(aes(x=median_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
+  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
+                   stat = "identity",
+                   alpha = 0.8, 
+                   pattern_fill = "black",
+                   colour = "black", 
+                   pattern_spacing = 0.04,
+                   pattern_alpha = 1) + 
   scale_y_reordered() +
-  facet_wrap(~site_id, scales = 'free') +
[…]

### 2023-11-27 R/get_forecast.R
if statement for corrupt files

@@ -9,47 +9,54 @@ get_forecast  <- function(theme,
                                   paste0('reference_datetime=', forecast_date)),
                         endpoint_override= "data.ecoforecast.org")
   
-  forecast <- arrow::open_dataset(s3_model) |>
-    collect() |> 
-    filter(variable == var, 
-           datetime > forecast_date) |>
-    group_by(site_id) |> 
-    # remove sites that contain NAs
-    filter(!any(is.na(prediction))) |> 
-    ungroup() |> 
-    mutate(model_id = model_id,
-           horizon = as_date(datetime) - as_date(forecast_date)) |> 
-    filter(horizon <= h) |> 
-    select(-any_of(c('horizon', 'pubDate')))
+  forecast <- try(arrow::open_dataset(s3_model) |>
+    collect(), silent = T) 
   
-  if (nrow(forecast) == 0) {
-    message('No forecast for ', var, ' using ', model_id)
-    return(forecast)
+  if (class(forecast)[1] == 'try-error') {
+    message(paste0('error in retrieving forecast for ', model_id ))
+    return(NULL)
   } else {
-    # do some more data wrangling
-    message(model_id, ' read in')
+    forecast <-  forecast |> 
+      filter(variable == var, 
+             datetime > forecast_date) |>
+      group_by(site_id) |> 
+      # remove sites that contain NAs
+      filter(!any(is.na(prediction))) |> 
+      ungroup() |> 
+      mutate(model_id = model_id,
+             horizon = as_date(datetime) - as_date(forecast_date)) |> 
+      filter(horizon <= h) |> 
+      select(-any_of(c('horizon', 'pubDate')))
     
-    if (forecast$family[1] != 'sample') {
-      message('generate ensemble of 100')
+    if (nrow(forecast) == 0) {
+      message('No forecast for ', var, ' using ', model_id)
+      return(forecast)
+    } else {
+      # do some more data wrangling
+      message(model_id, ' read in')
       
-      forecast <- forecast |> 
-        select(datetime, site_id, variable, family, parameter, prediction, model_id) |> 
-        pivot_wider(names_from = parameter,
-                    values_from = prediction, 
-                    id_cols = c(datetime, site_id, model_id)) |> 
+      if (forecast$family[1] != 'sample') {
+        message('generate ensemble of 100')
         
-        group_by(site_id, datetime, model_id) |> 
-        # sample from the distribution based on the mean and sd
[…]

### 2023-11-27 R/scoring_function_arrow.R
add scoring function

@@ -1,0 +1,44 @@
+generate_forecast_score_arrow <- function(targets_file,
+                                          forecast_df,
+                                          use_s3 = FALSE,
+                                          bucket = NULL,
+                                          endpoint = NULL,
+                                          local_directory = NULL){
+  
+  
+  if(use_s3){
+    if(is.null(bucket) | is.null(endpoint)){
+      stop("scoring function needs bucket and endpoint if use_s3=TRUE")
+    }
+    vars <- FLAREr:::arrow_env_vars()
+    output_directory <- arrow::s3_bucket(bucket = bucket,
+                                         endpoint_override =  endpoint)
+    FLAREr:::unset_arrow_vars(vars)
+  }else{
+    if(is.null(local_directory)){
+      stop("scoring function needs local_directory if use_s3=FALSE")
+    }
+    output_directory <- arrow::SubTreeFileSystem$create(local_directory)
+  }
+  
+  target <- readr::read_csv(targets_file, show_col_types = FALSE) 
+  
+  df <- forecast_df %>%
+    score4cast::standardize_forecast() %>%
+    dplyr::mutate(family = as.character(family)) |> 
+    score4cast::crps_logs_score(target) %>%
+    dplyr::mutate(horizon = datetime-lubridate::as_datetime(reference_datetime)) %>%
+    dplyr::mutate(horizon = as.numeric(lubridate::as.duration(horizon),
+                                       units = "seconds"),
+                  horizon = horizon / 86400) %>%
+    dplyr::mutate(depth = as.numeric(str_split_fixed(site_id, "-", 2)[,2]),
+                  site_id = str_split_fixed(site_id, "-", 2)[,1])
+  
+  
+  # reference_datetime_format <- "%Y-%m-%d %H:%M:%S"
+  
+  df <- df |> dplyr::mutate(reference_datetime = as_date(reference_datetime))
+  
+  arrow::write_dataset(df, path = output_directory, partitioning = c("model_id","site_id", "reference_datetime", "datetime"))
+  
+}

### 2023-11-27 AGU_analysis/analyse_scores.R
analyse all_submissions mme

@@ -1,0 +1,21 @@
+knitr::opts_chunk$set(echo = TRUE)
+library(tidyverse)
+library(arrow)
+library(lubridate)
+
+# read in the all submissions ensemble for the lake sites
+lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
+
+all_ensemble_scores <- arrow::open_dataset("scores_rescored") |>
+  filter(site_id %in% lake_sites,
+         model_id == 'all_submissions',
+         variable == 'temperature') |>
+  collect()
+
+all_ensemble_scores |> 
+  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
+  group_by(horizon, site_id, model_id) |> 
+  summarise(mean_crps = mean(crps, na.rm = T)) |> 
+  ggplot(aes(x=horizon, y= mean_crps)) +
+  geom_line(aes(colour = model_id)) +
+  facet_wrap(~site_id)

### 2023-11-27 AGU_analysis/model_means.R
change to date partition

@@ -73,7 +73,7 @@ if (!dir.exists(score_path)) {
 
 #### Read in forecasts ####
 # Open the dataset of forecast parquets to be scored
-forecast_parquets <- './Forecasts/parquet'
+forecast_parquets <- './Forecasts/parquet/model_id=all_submissions'
 open_parquets <- arrow::open_dataset(forecast_parquets)
 
 
@@ -81,22 +81,21 @@ open_parquets <- arrow::open_dataset(forecast_parquets)
 models <- 'all_submissions'
 
 # vector of unique reference_datetimes
-unique_reftime <- open_parquets |>
-  distinct(reference_datetime) |>
+unique_date <- open_parquets |>
+  distinct(date) |>
   collect() |>
-  arrange(reference_datetime) |>
-  pull(reference_datetime)
+  arrange(date) |>
+  pull(date)
 
 # get all combinations of model and reference_datetime
-to_score <- expand.grid(model_id = models, reference_datetime = unique_reftime)
+to_score <- expand.grid(model_id = models, date = unique_date)
 
 
 #### Regular scoring function ####
 for (i in 1:nrow(to_score)) {
   # subset the model and reference datetime
   forecast_df <- open_parquets|>
-    dplyr::filter(model_id == to_score$model_id[i],
-                  reference_datetime == to_score$reference_datetime[i]) |>
+    dplyr::filter(date == to_score$date[i]) |>
     collect()
 
   if (nrow(forecast_df) != 0) {

### 2023-11-27 R/scoring_function_arrow.R

@@ -37,8 +37,8 @@ generate_forecast_score_arrow <- function(targets_file,
   
   # reference_datetime_format <- "%Y-%m-%d %H:%M:%S"
   
-  df <- df |> dplyr::mutate(reference_datetime = as_date(reference_datetime))
+  df <- df |> dplyr::mutate(date = as.character(datetime))
   
-  arrow::write_dataset(df, path = output_directory, partitioning = c("model_id","site_id", "reference_datetime", "datetime"))
+  arrow::write_dataset(df, path = output_directory, partitioning = c("model_id","site_id", "date"))
   
 }

### 2023-11-27 AGU_analysis/model_means.R
mount at highest level

@@ -73,7 +73,7 @@ if (!dir.exists(score_path)) {
 
 #### Read in forecasts ####
 # Open the dataset of forecast parquets to be scored
-forecast_parquets <- './Forecasts/parquet/model_id=all_submissions'
+forecast_parquets <- './Forecasts/parquet/'
 open_parquets <- arrow::open_dataset(forecast_parquets)
 
 
@@ -82,6 +82,7 @@ models <- 'all_submissions'
 
 # vector of unique reference_datetimes
 unique_date <- open_parquets |>
+  filter(model_id == 'all_submissions') |> 
   distinct(date) |>
   collect() |>
   arrange(date) |>
@@ -95,7 +96,8 @@ to_score <- expand.grid(model_id = models, date = unique_date)
 for (i in 1:nrow(to_score)) {
   # subset the model and reference datetime
   forecast_df <- open_parquets|>
-    dplyr::filter(date == to_score$date[i]) |>
+    dplyr::filter(model_id == to_score$model_id[i],
+                  date == to_score$date[i]) |>
     collect()
 
   if (nrow(forecast_df) != 0) {

### 2024-01-05 Updates/2023-11-17 figures_draft.Rmd
recaulcute skill and save plots

@@
   # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = crps_climatology - crps,
-         skill_logs = logs_climatology - logs,
+  mutate(skill_crps = 1 - (crps/crps_climatology),
+         skill_logs = 1 - (logs/logs_climatology),
          horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
@@
   scale_y_reordered() +
-  labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
+  labs(y='Model', x = 'mean CRPS skill (vs climatology)') +
   geom_vline(xintercept = 0) +
@@
 
-all_temp <- temperature_skill |> 
+all_temp <- 
+  temperature_skill |> 
   group_by(model_id) |> 
@@
   ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
+  # geom_rect(aes(ymin = 'flareGOTM_noDA', 
+  #               ymax ='flareGLM', 
+  #               xmin = -Inf, 
+  #               xmax = Inf)) +
   geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
@@
   scale_y_reordered() +
-    labs(y='Model ID', x = 'mean CRPS skill (vs climatology)') +
+  labs(y='Model', x = 'mean CRPS skill (vs climatology)') +
   geom_vline(xintercept = 0) +
@@
 
-ggpubr::ggarrange(top10_temp, all_temp, nrow = 1, 
+ggpubr::ggarrange(top10_temp, all_temp, nrow = 1, align = 'hv',
                   common.legend = T,
-                  legend = 'top', labels = 'auto')
+                  legend = 'top', labels = 'auto') |> 
+  ggsave(filename = './Plots/Figure1.png', width = 25, height = 10, units = 'cm')
 
@@
 ```{r model_type-site}
-temperature_skill |> 
+clim_mean <- temperature_climatology |> 
+  group_by(site_id) |> 
+  summarise(mean_score = mean(crps_climatology, na.rm = T))  |> 
+  arrange(mean_score)
+
+n_model_type <- temperature_skill |> 
+  distinct(model_id) |> 
+  left_join(model_meta) |>
+  group_by(model_type) |> 
+  summarise(n = n())  |> 
+  mutate(site_id = 'TOOK', mean_skill = -3.5) |> 
+  filter(model_type != 'null')
+
+(temperature_skill |> 
   left_join(model_meta, by = 'model_id') |>
@@
              colour = model_type)) +
-  geom_vline(xintercept = 0, linetype = 'dashed') +
+  geom_vline(xintercept = 0, linetype = 'solid') +
+  geom_label(data = n_model_type, aes(label = paste0('n = ', n)))+ 
+  # geom_point(data = clim_mean, aes(x = mean_score, y = site_id), colour = 'black') +
   geom_boxplot() +
   facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
-  labs(y='model_id', x = 'mean CRPS score') +
+  labs(y='Site', x = 'mean CRPS skill (vs climatology)') +
   scale_colour_manual(values = cols_modeltype) +
-  theme_bw()
+  theme_bw()) |> 
+  ggsave(filename = './Plots/Figure2.png', width = 15, height = 15, units = 'cm')
 
-temperature_climatology |> 
-  group_by(site_id) |> 
-  summarise(mean_score = mean(crps_climatology, na.rm = T))  |> 
-  arrange(mean_score)
 ```
@@
   geom_line()  +  
-  labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
+  labs(y='mean CRPS skill (vs climatology)', x = 'Horizon (days)') +
   scale_color_viridis_d(option = 'H') +
@@
   geom_line()  +  
-  labs(y='mean CRPS skill (vs climatology)', x = 'horizon (days)') +
-  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
+  labs(y='mean CRPS skill (vs climatology)', x = 'Horizon (days)') +
+  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9, name = 'Model') +
   theme_bw()
 
-ggpubr::ggarrange(A, B, labels = 'auto', align = 'v', nrow = 2)
+ggpubr::ggarrange(A, B, labels = 'auto', align = 'v', nrow = 2) |> 
+  ggsave(filename = 'Plots/Figure3.png', width = 15, height = 17, units = 'cm')
 ```
@@
   geom_line()  +  
-  labs(y='mean standard deviation', x = 'horizon (days)') +
+  labs(y='mean standard deviation', x = 'Horizon (days)') +
   scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
@@
   geom_line()  +  
-  labs(y='mean absolute bias (oC)', x = 'horizon (days)') +
+  labs(y='mean absolute bias', x = 'Horizon (days)') +
   scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
@@
 
-ggpubr::ggarrange(sd, bias, labels = 'auto', align = 'v', nrow = 2)
+ggpubr::ggarrange(sd, bias, labels = 'auto', align = 'v', nrow = 2) |> 
+  ggsave(filename = 'Plots/Figure4.png')
 ```

### 2024-01-15 Updates/2023-11-17 figures_draft.Rmd
reliability and observation plots

@@
 library(ggpattern)
+library(ggpubr)
 ```
@@
 
-lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
+lake_sites <- c('TOOK', 'PRLA','PRPO', 'CRAM', 'LIRO','BARC', 'SUGG')
 cols_modeltype <- c('empirical' = "#482576FF",
@@
                     'ML' = '#BBDF27FF',
-                    'null' = '#43BF71FF')
+                    'null' = '#A9A9A9')
 # read in the metadata from the google sheet
@@
   # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = 1 - (crps/crps_climatology),
-         skill_logs = 1 - (logs/logs_climatology),
+  mutate(skill_crps = crps_climatology - crps,
+         skill_logs = logs_climatology - logs,
          horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
@@
                    pattern_fill = "black",
-                   colour = "black", 
+                   colour = "black", width = 0.8,
                    pattern_spacing = 0.03,
@@
   summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
   ggplot(aes(x=horizon, 
@@
 
-B <- temperature_skill |> 
+B <- 
+  temperature_skill |> 
   filter(model_id %in% top_mods$model_id) |> 
@@
   ggplot(aes(x=horizon, 
-             y = mean_crps, colour = model_id)) +
+             y = mean_crps, colour = fct_reorder(model_id, mean_crps, .desc = T))) +
   geom_hline(yintercept = 0) + 
@@
   labs(y='mean CRPS skill (vs climatology)', x = 'Horizon (days)') +
-  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9, name = 'Model') +
+  scale_color_brewer(name = 'Model', palette = 'Paired') +
   theme_bw()
@@
   slice_max(mean_crps, n = 10, na_rm = T) |> 
+  mutate(model_id = fct_reorder(model_id, mean_crps, .desc = T)) |> 
   distinct(model_id)
@@
          horizon > 0, 
+         horizon != 15,
          horizon <=30) |>
@@
   summarise(mean_sd = mean(sd, na.rm = T)) |> 
+  mutate(model_id = factor(model_id, levels = levels(top_mods$model_id))) |> 
   ggplot(aes(x=horizon, 
              y = mean_sd, colour = model_id)) +
-  geom_line()  +  
+  geom_line()+
   labs(y='mean standard deviation', x = 'Horizon (days)') +
-  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
+  scale_color_brewer(name = 'Model', palette = 'Paired') +
   theme_bw()
@@
   summarise(mean_abs_bias = mean(abs_bias, na.rm = T)) |> 
+  mutate(model_id = factor(model_id, levels = levels(top_mods$model_id))) |> 
   ggplot(aes(x=horizon, 
              y = mean_abs_bias, colour = model_id)) +
-  geom_line()  +  
+  geom_line()+
   labs(y='mean absolute bias', x = 'Horizon (days)') +
-  scale_color_viridis_d(option = 'A', begin = 0.1, end =0.9) +
+  scale_color_brewer(name = 'Model', palette = 'Paired') +
   theme_bw() +
@@
 
-ggpubr::ggarrange(sd, bias, labels = 'auto', align = 'v', nrow = 2) |> 
-  ggsave(filename = 'Plots/Figure4.png')
+ggpubr::ggarrange(sd, bias, labels = 'auto', align = 'v', nrow = 2, common.legend = T, legend = 'right') |> 
+  ggsave(filename = 'Plots/Figure4.png', height = 15, width = 14, units = 'cm')
 ```
@@
   filter(horizon <= 30) |> 
+  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
   ggplot(aes(x=horizon, y=mean_crps, colour = model_id)) +
@@
 
+## Figure 5: reliability plots
+
+```{r}
+source('../R/reliability_plot.R')
+
+good_80 <- 
+  plot_reliability(scores = temperature_scores,
+                            models = top_mods$model_id, 
+                            CI = 80, 
+                            order = levels(top_mods$model_id)) +
+  scale_color_brewer(name = 'Model', palette = 'Paired') +
+  theme_bw() +
+  scale_y_continuous(limits = c(45, 100))
+
+good_95 <- plot_reliability(scores = temperature_scores,
+                            models = top_mods$model_id, 
+                            CI = 95, 
+                            order = levels(top_mods$model_id)) +
+  scale_color_brewer(name = 'Model', palette = 'Paired') +
+  theme_bw() +
+  scale_y_continuous(limits = c(45, 100))
+
+bad_mods <- model_meta |> 
+  filter(!model_id %in% top_mods$model_id) |> 
+  pull(model_id)
+
+bad_80 <- 
+  plot_reliability(scores = temperature_scores,
+                   models = bad_mods, 
+                   CI = 80, 
+                   order = bad_mods) +
+  scale_colour_discrete(name = 'Model') +
+  theme_bw() +
+  scale_y_continuous(limits = c(0, 100))
+
+bad_95 <- plot_reliability(scores = temperature_scores,
+                            models = bad_mods, 
+                            CI = 95, 
+                            order = bad_mods) +
+  scale_colour_discrete(name = 'Model') +
+  theme_bw() +
+  scale_y_continuous(limits = c(0, 100))
+
+ggarrange(ggarrange(good_80, good_95, bad_80, bad_95, align = 'hv', legend= 'none'), nrow = 2, align = 'hv') |> 
+  ggsave(filename = 'Plots/Figure5.png', height = 25, width = 18, units = 'cm')
+```
+
 ## site differences in observations?
 
+```{r target-plot}
+targets <- readr::read_csv("https://data.ecoforecast.org/neon4cast-targets/aquatics/aquatics-targets.csv.gz") |> 
+  na.omit()
+
+layout_design <- "
+  A#
+  BC
+  DE
+  FG
+"
+
+# is 2023 a stand out year?
+obs <- targets |> 
+  filter(variable == 'temperature',
+         site_id %in% lake_sites) |> 
+  mutate(doy = yday(datetime),
+         year = year(datetime),
+         is_2023 = ifelse(year == 2023, 2023, 2022),
+         site_id = factor(site_id, levels = lake_sites)) |> 
+  ggplot(aes(x=doy, y=observation, colour = is_2023)) +
+  geom_jitter(show.legend = F, aes(alpha = is_2023), size = 0.7, height = 0.2) +
+  facet_manual(~factor(site_id, levels = lake_sites),
+                 design = layout_design) +
+  scale_colour_viridis_c(begin = 0, end = 0.6) +
+  labs(x= 'Day of year', y = 'Observation (oC)') +
+  scale_alpha(range = c(0.5, 0.8))+
+  scale_x_continuous(expand = c(0.01,0)) +
+  scale_y_continuous(expand = c(0,0),
+                     limits = c(0,35)) +
+  theme_bw()
+
+ggsave(obs, filename = 'Plots/Figure6.png', width = 20, height=15, units = 'cm')
+
+```
+
 ```{r obs-cv}
-temperature_scores |> 
-  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-  filter(between(horizon, 0,30)) |> 
-  group_by(site_id, variable, horizon) |> 
+cv_obs <- targets |> 
+  mutate(site_id = factor(site_id, levels = lake_sites), 
+         year = year(datetime)) |> 
+  filter(variable == 'temperature',
+         site_id %in% lake_sites,
+         year < 2024) |> 
+  group_by(site_id, year, variable) |> 
   summarise(cv = sd(observation, na.rm = T)/ mean(observation, na.rm = T)) |> 
-  ggplot(aes(x=horizon, y= cv, colour = site_id)) +
+  ggplot(aes(x=year, y= cv, colour = site_id)) +
+  geom_line() +
   geom_point() +
@@
   theme_bw()
+
+ggarrange(obs, cv_obs)
+```
+## Reliability plots
+
+```{r}
+layout_design <- "
+  A#
+  BC
+  DE
+  FG
+"
+
+temperature_scores |> 
+  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+  filter(model_id %in% top_mods$model_id, 
+         horizon >0,
+         horizon <=30,
+         !is.na(observation)) |> 
+  mutate(cf = between(observation, quantile10, quantile90),
+         site_id = factor(site_id, levels = lake_sites), 
+         model_id = factor(model_id, levels = levels(top_mods$model_id))) |> 
+  group_by(horizon, cf, site_id, model_id) |> 
+  summarise(n = n()) |> 
+  pivot_wider(names_from = cf, names_prefix = 'within_', values_from = n, values_fill = 0) |> 
+  mutate(perc = within_TRUE/(within_FALSE + within_TRUE)*100) |> 
+  ggplot(aes(x=horizon, y=perc, colour = model_id)) +
+  geom_hline(yintercept = 80, colour = 'grey3', linetype = 'dashed')+
+  geom_line(alpha = 0.8) +
+  scale_color_brewer(name = 'Model', palette = 'Paired') +
+  labs(y = 'Percentage of observations within\n95% confidence intervals', x='Horizon (days)') +
+  facet_manual(~factor(site_id, levels = lake_sites),
+                 design = layout_design) +
+  scale_x_continuous(expand = c(0,0)) +
+  theme_bw()
+
+temperature_scores |> 
+  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+  filter(model_id %in% top_mods$model_id, 
+         horizon >0,
+         horizon <=30,
+         !is.na(observation)) |> 
+  mutate(cf = between(observation, quantile02.5, quantile97.5),
+         site_id = factor(site_id, levels = lake_sites), 
+         model_id = factor(model_id, levels = levels(top_mods$model_id))) |> 
+  group_by(horizon, cf, site_id, model_id) |> 
+  summarise(n = n()) |> 
+  pivot_wider(names_from = cf, names_prefix = 'within_', values_from = n, values_fill = 0) |> 
+  mutate(perc = within_TRUE/(within_FALSE + within_TRUE)*100) |> 
+  ggplot(aes(x=horizon, y=perc, colour = model_id)) +
+  geom_hline(yintercept = 95, colour = 'grey3', linetype = 'dashed')+
+  geom_line(alpha = 0.8) +
+  scale_color_brewer(name = 'Model', palette = 'Paired') +
+  labs(y = 'Percentage of observations within\n95% confidence intervals', x='Horizon (days)') +
+  facet_manual(~factor(site_id, levels = lake_sites),
+                 design = layout_design) +
+  scale_x_continuous(expand = c(0,0)) +
+  theme_bw()
+
 ```
@@
   summarise(mean_skill = mean(skill_crps, na.rm = T)) |> 
+  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
   ggplot(aes(x = mean_skill, 
@@
   summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
+  mutate(site_id = factor(site_id, levels = rev(lake_sites))) |> 
   ggplot(aes(x = mean_skill, 

### 2024-01-15 Updates/2023-11-17 figures_draft.Rmd
omit day 15

@@
   filter(site_id %in% lake_sites,
-         # model_id != 'climatology',
+         horizon != 15,
          model_id %in% model_meta$model_id,
@@
          site_id %in% lake_sites,
+         horizon != 15,
          variable == 'temperature') |>

### 2024-02-02 01_get_model_list.R
specify which scores to get

@@ -13,14 +13,8 @@ model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/
 
 
 # What and when?
-forecast_dates <- as.character(seq.Date(as_date('2023-01-01'), Sys.Date(), 'day'))
+forecast_dates <- as.character(seq.Date(as_date('2023-01-01'), as_date('2023-12-31'), 'day'))
 variables <- c('temperature', 'oxygen', 'chla')
-
-# Build/update local database using the recent_submissions
-
-# Grab the data from the score bucket
-scores_s3 <- arrow::s3_bucket("neon4cast-scores/parquet/aquatics/",
-                              endpoint_override= "data.ecoforecast.org", anonymous = T)
 
 
 # build a local parquet database from which to query
@@ -29,6 +23,6 @@ scores_s3 |>
   dplyr::filter(reference_datetime %in% forecast_dates,
                 variable %in% variables,
                 model_id %in% model_meta$model_id) |>
-  arrow::write_dataset("scores_rescored", partitioning=c("model_id", "site_id"))
+  arrow::write_dataset("scores_2023", partitioning=c("model_id", "site_id"))
 
 #====================================================================#

### 2024-02-02 Updates/2023-11-17 figures_draft.Rmd
drafts

--- 2024-01-15 Updates/2023-11-17 figures_draft.Rmd
@@
 # start with lake temperatures
-temperature_scores <- arrow::open_dataset("../scores_rescored") |>
+temperature_scores <- arrow::open_dataset("../scores_2023") |>
   filter(site_id %in% lake_sites,
-         # model_id != 'climatology',
          model_id %in% model_meta$model_id,
@@
 # Calculate skill scores ------------
-temperature_climatology <- arrow::open_dataset("../scores_rescored") |>
+temperature_climatology <- arrow::open_dataset("../scores_2023") |>
   filter(model_id == 'climatology', 
@@
          horizon >= 1, 
-         model_id != 'fARIMA') |> 
+         horizon != 15) |> 
   arrange(reference_datetime, site_id, datetime, model_id) 
@@
 
-```{r}
+```{r eval = F}
 NOAA_scores <- arrow::open_dataset("../scores") |>
@@
                             order = levels(top_mods$model_id)) +
-  scale_color_brewer(name = 'Model', palette = 'Paired') +
+  scale_color_brewer(name = 'Top Models', palette = 'Paired') +
   theme_bw() +
-  scale_y_continuous(limits = c(45, 100))
+  scale_y_continuous(limits = c(0, 100))
 
@@
                    order = bad_mods) +
-  scale_colour_discrete(name = 'Model') +
   theme_bw() +
-  scale_y_continuous(limits = c(0, 100))
+  scale_colour_discrete(name = '') +
+  scale_y_continuous(limits = c(0, 100)) + 
+  theme(legend.position = 'bottom') + 
+  guides(colour = guide_legend(nrow = 5))
 
@@
                             order = bad_mods) +
-  scale_colour_discrete(name = 'Model') +
   theme_bw() +
@@
 
-ggarrange(ggarrange(good_80, good_95, bad_80, bad_95, align = 'hv', legend= 'none'), nrow = 2, align = 'hv') |> 
-  ggsave(filename = 'Plots/Figure5.png', height = 25, width = 18, units = 'cm')
+ggarrange(ggarrange(ggarrange(get_legend(good_80), NULL, nrow = 2),
+          ggarrange(good_80, good_95, bad_80, bad_95, align = 'hv', legend= 'none'), ncol = 2, 
+          widths = c(1,5)),
+          ggarrange(NULL, get_legend(bad_80), ncol = 2, widths = c(1,5)),
+          ncol = 1, heights = c(5, 1)) |> 
+  ggsave(filename = 'Plots/Figure5.png', height = 17.6, width = 22, units = 'cm')
 ```

### 2024-02-02 AGU_analysis/analyse_scores.R

@@ -12,12 +12,12 @@ all_ensemble_scores <- arrow::open_dataset("scores_rescored/") |>
          variable == 'temperature') |>
   collect() 
 
-all_ensemble_scores |> 
+top10 <- all_ensemble_scores |> 
   mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |>
-  group_by(model_id, site_id) |> 
+  group_by(model_id) |> 
   summarise(mean_crps = mean(crps, na.rm = T)) |> 
-  group_by(site_id) |> 
-  slice_min(mean_crps, n= 5)
+  ungroup() |> 
+  slice_min(mean_crps, n= 10)
 
 googlesheets4::gs4_deauth()
 model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")
@@ -155,8 +155,10 @@ design <- "
 ranks |> 
   full_join(all_vals) |> 
   mutate(proportion = ifelse(is.na(proportion), 0, proportion)) |> 
-  filter(model_id %in% c('all_submissions', 'cb_prophet', 'GLEON_physics', 'flareGLM', 
-                         'fARIMA','tg_randfor', 'xgboost_parallel')) |> 
+  filter(model_id %in% c('all_submissions', 'cb_prophet', 'flareGLM', 'xgboost_parallel', 
+                         'air2waterSat_2', 'tg_ets', 'tg_tbats')) |> 
+  mutate(model_id = factor(model_id, levels = c('all_submissions', 'cb_prophet', 'flareGLM', 'xgboost_parallel', 
+                                      'air2waterSat_2', 'tg_ets', 'tg_tbats'))) |> 
   mutate(group_rank = cut(rank,
                           breaks = ceiling(seq(0, 31, length.out =11)),
                           labels = seq(1,10, length.out = 10))) |> 
@@ -175,7 +177,7 @@ ranks |>
         panel.spacing.y = unit(1.1, 'lines'), 
         legend.position = 'none', panel.grid.major.y = element_blank()) +
   scale_x_continuous(name = '% of forecasts', breaks = seq(0, 0.3, 0.1), labels = seq(0, 30, 10)) +
-  scale_y_discrete(breaks = c(1,10), labels = c('best 10 %', 'worst 10 %'), name = '')
-#------------------------------------------
[…]

### 2024-02-02 R/reliability_plot.R
function to generate reliability plot

@@ -1,0 +1,50 @@
+plot_reliability <- function(scores, models, CI, order = levels(models)) {
+  
+  if (CI == 95) {
+    plot <- scores |> 
+      mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+      filter(model_id %in% models,
+             horizon >0,
+             horizon != 15, 
+             horizon <=30,
+             !is.na(observation)) |> 
+      mutate(is_in = between(observation, quantile02.5, quantile97.5)) |> 
+      group_by(horizon, is_in, model_id) |> 
+      summarise(n = n()) |> 
+      pivot_wider(names_from = is_in, names_prefix = 'within_', values_from = n, values_fill = 0) |> 
+      mutate(perc = within_TRUE/(within_FALSE + within_TRUE)*100, 
+             model_id = factor(model_id, levels = order)) |> 
+      ggplot(aes(x=horizon, y=perc, colour = model_id)) +
+      geom_hline(yintercept = 95, colour = 'grey3', linetype = 'dashed')+
+      geom_line(alpha = 0.8) +
+      labs(y = 'Percentage of observations within\n95% confidence intervals', x='Horizon (days)') +
+      annotate('text', x = 25, y = 100, label = 'underconfident') +
+      annotate('text', x = 25, y = 75, label = 'overconfident')
+  }
+  if (CI == 80) {
+    plot <- scores |> 
+      mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+      filter(model_id %in% models,
+             horizon >0,
+             horizon != 15, 
+             horizon <=30,
+             !is.na(observation)) |> 
+      mutate(is_in = between(observation, quantile10, quantile90)) |> 
+      group_by(horizon, is_in, model_id) |> 
+      summarise(n = n()) |> 
+      pivot_wider(names_from = is_in, names_prefix = 'within_', values_from = n, values_fill = 0) |> 
+      mutate(perc = within_TRUE/(within_FALSE + within_TRUE)*100, 
+             model_id = factor(model_id, levels = order)) |> 
+      ggplot(aes(x=horizon, y=perc, colour = model_id)) +
+      geom_hline(yintercept = 80, colour = 'grey3', linetype = 'dashed')+
+      geom_line(alpha = 0.8) +
+      labs(y = 'Percentage of observations within\n80% confidence intervals', x='Horizon (days)') +
[…]

### 2024-02-02 04_manuscript_plots.Rmd
manuscript plots

--- 2024-02-02 Updates/2023-11-17 figures_draft.Rmd
@@
 ---
-title: "figures?"
+title: "Manuscript figures"
 output: html_document
-date: "2023-11-18"
+date: "2024-01-24"
 ---
@@
 library(ggpubr)
+library(ggh4x)
 ```
@@
 ## Read in the data and model metadata
-
 ```{r read-in-data}
@@
 # start with lake temperatures
-temperature_scores <- arrow::open_dataset("../scores_2023") |>
+temperature_scores <- arrow::open_dataset("./scores_2023") |>
   filter(site_id %in% lake_sites,
@@
 # Calculate skill scores ------------
-temperature_climatology <- arrow::open_dataset("../scores_2023") |>
+temperature_climatology <- arrow::open_dataset("./scores_2023") |>
   filter(model_id == 'climatology', 
@@
 
-Top 10 models and all models coloured by type and showing if they include covariates or not. 
 ```{r model-type, fig.width=10}
@@
   ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
-  # geom_rect(aes(ymin = 'flareGOTM_noDA', 
-  #               ymax ='flareGLM', 
-  #               xmin = -Inf, 
-  #               xmax = Inf)) +
+
   geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
@@
 
-9 of 10 top 10 models use covariates, and process models and ML are the most common best performers. 
-There are a lot of really bad empirical models.
-
 ## Figure 2
-How does model type performance, on avergae, differ by site.
-
-This uses all model_ids (maybe get rid of the 'worst') performers.
-
-```{r model_type-site}
-clim_mean <- temperature_climatology |> 
-  group_by(site_id) |> 
-  summarise(mean_score = mean(crps_climatology, na.rm = T))  |> 
-  arrange(mean_score)
-
-n_model_type <- temperature_skill |> 
-  distinct(model_id) |> 
-  left_join(model_meta) |>
-  group_by(model_type) |> 
-  summarise(n = n())  |> 
-  mutate(site_id = 'TOOK', mean_skill = -3.5) |> 
-  filter(model_type != 'null')
-
-(temperature_skill |> 
-  left_join(model_meta, by = 'model_id') |>
-  filter(model_type != 'null') |> 
-  group_by(model_type, model_id, site_id) |> 
-  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
-  ggplot(aes(x = mean_skill, 
-             y = site_id, 
-             colour = model_type)) +
-  geom_vline(xintercept = 0, linetype = 'solid') +
-  geom_label(data = n_model_type, aes(label = paste0('n = ', n)))+ 
-  # geom_point(data = clim_mean, aes(x = mean_score, y = site_id), colour = 'black') +
-  geom_boxplot() +
-  facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
-  labs(y='Site', x = 'mean CRPS skill (vs climatology)') +
-  scale_colour_manual(values = cols_modeltype) +
-  theme_bw()) |> 
-  ggsave(filename = './Plots/Figure2.png', width = 15, height = 15, units = 'cm')
-
-```
-
-Gains at PRLA and PRPO relative to climatology (these have the worst climatology - see table), especially using process-models and ML.
-
-The best climatology sites (BARC, SUGG) are difficult to beat even though perform well overall.
-
-## Figure 3
-How does the performance vary across horizon? 
-Do the models perform better consistently across the horizon at each site?
-How do the best performing individual models perform?
-
 ```{r horizon-score}
@@
   distinct(model_id)
-
-A <- temperature_skill |> 
-  filter(model_id %in% top_mods$model_id) |> 
-  group_by(site_id, horizon) |> 
-  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
-  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
-  ggplot(aes(x=horizon, 
-             y = mean_crps, colour = site_id)) +
-  geom_hline(yintercept = 0) + 
-  geom_line()  +  
-  labs(y='mean CRPS skill (vs climatology)', x = 'Horizon (days)') +
-  scale_color_viridis_d(option = 'H') +
-  theme_bw()
 
@@
 
-ggpubr::ggarrange(A, B, labels = 'auto', align = 'v', nrow = 2) |> 
-  ggsave(filename = 'Plots/Figure3.png', width = 15, height = 17, units = 'cm')
+ggpubr::ggarrange(B) |> 
+  ggsave(filename = 'Plots/Figure2.png', width = 15, height = 17, units = 'cm')
 ```
 
-Models beeating climatology at the shorter horizons (as expected) PRLA and PRPO always better than climatology.
-
-Individual model performance decreases relative to climatology mostly except for random forest and lasso regression which improve after short horizons (may be related to the precision rather than accuracy).
-Less degradation in performance of ML models across the horizon (xgboost and random forest) than process models.
-
-### How much is uncertainty?
-
-How does the sd vary - is this my some see a increase in crps skill?
-
-```{r horizon-score}
+## Figure 3
+```{r bias-sd}
 # top 10 average models
@@
 ggpubr::ggarrange(sd, bias, labels = 'auto', align = 'v', nrow = 2, common.legend = T, legend = 'right') |> 
-  ggsave(filename = 'Plots/Figure4.png', height = 15, width = 14, units = 'cm')
+  ggsave(filename = 'Plots/Figure3.png', height = 15, width = 14, units = 'cm')
 ```
 
-## Figure 4 - NOAA scores
+## Figure 4
 
-can this be related to the weather forecasts?
-
-There is only one month when the observations overlap for the observational data...
-But still the performance is better vs the observations than the stage 3
-
-```{r eval = F}
-NOAA_scores <- arrow::open_dataset("../scores") |>
-  filter(site_id %in% lake_sites,
-         model_id %in% c('NOAA_daily', 'NOAA_daily_obs')) |>
-  collect()
-
-NOAA_scores |> 
-  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
-  select(site_id, crps, horizon, reference_datetime, model_id) |> 
-  pivot_wider(names_from = site_id, values_from = crps, id_cols = horizon:model_id) |> 
-  filter(reference_datetime > as_date('2023-05-25'),
-         reference_datetime < as_date('2023-07-31')) |> 
-  # only a 1 month overlap....
-  pivot_longer(cols = BARC:TOOK, names_to = 'site_id', values_to = 'crps') |> 
-  group_by(model_id, site_id, horizon) |> 
-  summarise(mean_crps = mean(crps, na.rm = T)) |> 
-  filter(horizon <= 30) |> 
-  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
-  ggplot(aes(x=horizon, y=mean_crps, colour = model_id)) +
-  geom_line() +
-  facet_wrap(~site_id) +
-  theme_bw() +
-  labs(y = 'CRPS score (oC)', title = 'NOAA air temperature forecast')
-
-```
-
-## Figure 5: reliability plots
-
-```{r}
-source('../R/reliability_plot.R')
+```{r reliability}
+source('./R/reliability_plot.R')
 
@@
           ncol = 1, heights = c(5, 1)) |> 
-  ggsave(filename = 'Plots/Figure5.png', height = 17.6, width = 22, units = 'cm')
+  ggsave(filename = 'Plots/Figure4.png', height = 17.6, width = 22, units = 'cm')
 ```
 
-## site differences in observations?
+## Figure 5
+```{r model_type-site}
+n_model_type <- temperature_skill |> 
+  distinct(model_id) |> 
+  left_join(model_meta) |>
+  group_by(model_type) |> 
+  summarise(n = n())  |> 
+  mutate(site_id = 'TOOK', mean_skill = -3.5) |> 
+  filter(model_type != 'null')
+
+# top 10 average models
+top_mods <- temperature_skill |> 
+  group_by(model_id) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  ungroup() |> 
+  slice_max(mean_crps, n = 10, na_rm = T) |> 
+  distinct(model_id)
+
+A <- temperature_skill |> 
+  left_join(model_meta, by = 'model_id') |>
+  filter(model_type != 'null') |> 
+  group_by(model_type, model_id, site_id) |> 
+  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
+  ggplot(aes(x = mean_skill, 
+             y = site_id, 
+             colour = model_type)) +
+  geom_vline(xintercept = 0, linetype = 'solid') +
+  geom_label(data = n_model_type, aes(label = paste0('n = ', n)))+ 
+  geom_boxplot() +
+  facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
+  labs(y='Site', x = 'mean CRPS skill (vs climatology)') +
+  scale_colour_manual(values = cols_modeltype) +
+  theme_bw()
+
+B <- temperature_skill |> 
+  filter(model_id %in% top_mods$model_id) |> 
+  group_by(site_id, horizon) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
+  ggplot(aes(x=horizon, 
+             y = mean_crps, colour = site_id)) +
+  geom_hline(yintercept = 0) + 
+  geom_line()  +  
+  labs(y='mean CRPS skill (vs climatology)', x = 'Horizon (days)') +
+  scale_color_viridis_d(option = 'H') +
+  theme_bw()
+
+ggarrange(A, B, widths = c(1,1)) |> 
+  ggsave(filename = './Plots/Figure5.png', width = 25, height = 15, units = 'cm')
+```
+
+## Figure 6
 
@@
   filter(variable == 'temperature',
-         site_id %in% lake_sites) |> 
+         site_id %in% lake_sites, 
+         year(datetime) < 2024) |> 
   mutate(doy = yday(datetime),
@@
                  design = layout_design) +
-  scale_colour_viridis_c(begin = 0, end = 0.6) +
-  labs(x= 'Day of year', y = 'Observation (oC)') +
+  scale_colour_viridis_c(option = 'mako', begin = 0.1, end = 0.8) +
+  labs(x= 'Day of year', y = 'Observation (°C)') +
   scale_alpha(range = c(0.5, 0.8))+
@@
 
-ggsave(obs, filename = 'Plots/Figure6.png', width = 20, height=15, units = 'cm')
-
-```
-
-```{r obs-cv}
-cv_obs <- targets |> 
+cv_obs <- 
+  targets |> 
   mutate(site_id = factor(site_id, levels = lake_sites), 
@@
   summarise(cv = sd(observation, na.rm = T)/ mean(observation, na.rm = T)) |> 
-  ggplot(aes(x=year, y= cv, colour = site_id)) +
-  geom_line() +
+  mutate(is_2023 = ifelse(year == 2023, '2023', 'other years' )) |> 
+  ggplot(aes(y = site_id, x = cv, colour = is_2023)) +
   geom_point() +
-  scale_colour_viridis_d(option = 'turbo') +
-  theme_bw()
+  scale_colour_viridis_d(option = 'mako', begin = 0.8, end = 0.1, name = '') +
+  scale_y_discrete(limits = rev)+
+  theme_bw() +
+  labs(x = 'Coefficient of variation', y = 'Site')
 
-ggarrange(obs, cv_obs)
-```
-## Reliability plots
-
-```{r}
-layout_design <- "
-  A#
-  BC
-  DE
-  FG
-"
-
-temperature_scores |> 
-  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-  filter(model_id %in% top_mods$model_id, 
-         horizon >0,
-         horizon <=30,
-         !is.na(observation)) |> 
-  mutate(cf = between(observation, quantile10, quantile90),
-         site_id = factor(site_id, levels = lake_sites), 
-         model_id = factor(model_id, levels = levels(top_mods$model_id))) |> 
-  group_by(horizon, cf, site_id, model_id) |> 
-  summarise(n = n()) |> 
-  pivot_wider(names_from = cf, names_prefix = 'within_', values_from = n, values_fill = 0) |> 
-  mutate(perc = within_TRUE/(within_FALSE + within_TRUE)*100) |> 
-  ggplot(aes(x=horizon, y=perc, colour = model_id)) +
-  geom_hline(yintercept = 80, colour = 'grey3', linetype = 'dashed')+
-  geom_line(alpha = 0.8) +
-  scale_color_brewer(name = 'Model', palette = 'Paired') +
-  labs(y = 'Percentage of observations within\n95% confidence intervals', x='Horizon (days)') +
-  facet_manual(~factor(site_id, levels = lake_sites),
-                 design = layout_design) +
-  scale_x_continuous(expand = c(0,0)) +
-  theme_bw()
-
-temperature_scores |> 
-  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-  filter(model_id %in% top_mods$model_id, 
-         horizon >0,
-         horizon <=30,
-         !is.na(observation)) |> 
-  mutate(cf = between(observation, quantile02.5, quantile97.5),
-         site_id = factor(site_id, levels = lake_sites), 
-         model_id = factor(model_id, levels = levels(top_mods$model_id))) |> 
-  group_by(horizon, cf, site_id, model_id) |> 
-  summarise(n = n()) |> 
-  pivot_wider(names_from = cf, names_prefix = 'within_', values_from = n, values_fill = 0) |> 
-  mutate(perc = within_TRUE/(within_FALSE + within_TRUE)*100) |> 
-  ggplot(aes(x=horizon, y=perc, colour = model_id)) +
-  geom_hline(yintercept = 95, colour = 'grey3', linetype = 'dashed')+
-  geom_line(alpha = 0.8) +
-  scale_color_brewer(name = 'Model', palette = 'Paired') +
-  labs(y = 'Percentage of observations within\n95% confidence intervals', x='Horizon (days)') +
-  facet_manual(~factor(site_id, levels = lake_sites),
-                 design = layout_design) +
-  scale_x_continuous(expand = c(0,0)) +
-  theme_bw()
+ggarrange(obs, cv_obs, widths = c(1, 1.5)) |> 
+  ggsave(filename = 'Plots/Figure6.png', width = 20, height=10, units = 'cm')
 
@@
 
-## Supplementary figures
-
-```{r echo = F}
-temperature_skill |> 
-  left_join(model_meta, by = 'model_id') |>
-  filter(model_type != 'null') |> 
-  group_by(model_type, model_id, site_id) |> 
-  summarise(mean_skill = mean(skill_crps, na.rm = T)) |> 
-  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
-  ggplot(aes(x = mean_skill, 
-             y = site_id, 
-             colour = model_type)) +
-  geom_vline(xintercept = 0, linetype = 'dashed') +
-  geom_boxplot() +
-  facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
-  labs(y='model_id', x = 'mean CRPS score', title = 'all dates') +
-  scale_colour_manual(values = cols_modeltype) +
-  theme_bw()
-
-temperature_skill |> 
-  left_join(model_meta, by = 'model_id') |>
-  filter(model_type != 'null',
-         reference_datetime > as_date('2023-06-28'),
-         reference_datetime < as_date('2023-09-18')) |> 
-  group_by(model_type, model_id, site_id) |> 
-  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
-  mutate(site_id = factor(site_id, levels = rev(lake_sites))) |> 
-  ggplot(aes(x = mean_skill, 
-             y = site_id, 
-             colour = model_type)) +
-  geom_vline(xintercept = 0, linetype = 'dashed') +
-  geom_boxplot() +
-  facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
-  labs(y='model_id', x = 'mean CRPS score', title = 'overlapping only') +
-  scale_colour_manual(values = cols_modeltype) +
-  theme_bw() 
+```{r}
+targets |> 
+  filter(variable == 'temperature',
+         site_id %in% lake_sites, 
+         year(datetime) < 2024) |> 
+  mutate(doy = yday(datetime),
+         year = year(datetime),
+         site_id = factor(site_id, levels = lake_sites), 
+         is_2023 = ifelse(year == 2023, T, F))  |> 
+  group_by(site_id, is_2023) |> 
+  summarise(mean = mean(observation),
+            sd = sd(observation)) |> 
+  filter(site_id  == 'BARC')
 ```
-
-## Chlorophyll and oxygen patterns same?
-
-```{r read-WQ, echo = F}
-# Open dataset-------------
-chla_scores <- arrow::open_dataset("../scores_rescored") |>
-  filter(site_id %in% lake_sites,
-         # model_id != 'climatology',
-         model_id %in% model_meta$model_id,
-         variable == 'chla') |>
-  collect()
-
-# Calculate skill scores ------------
-chla_persistence <- arrow::open_dataset("../scores_rescored") |>
-  filter(model_id == 'persistenceRW', 
-         site_id %in% lake_sites,
-         variable == 'chla') |>
-  collect() |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  arrange(reference_datetime, datetime, site_id) |> 
-  pivot_wider(values_from = c(crps, logs),
-              names_from = model_id)
-
-chla_skill <- chla_scores |>
-  filter(model_id != 'persistenceRW') |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  full_join(chla_persistence) |>
-
-  # calcaulte the skill relative to persistence (- is bad, + is good)
-  mutate(skill_crps = crps_persistenceRW - crps,
-         skill_logs = logs_persistenceRW - logs,
-         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-  # consistent forecast period
-  filter(horizon <= 30,
-         horizon >= 1,
-         model_id != 'fARIMA') |> 
-  arrange(reference_datetime, site_id, datetime, model_id)
-
-#--------------------------------------#
-# Dissolved oxygen
-
-# Open dataset-------------
-oxygen_scores <- arrow::open_dataset("../scores_rescored") |>
-  filter(site_id %in% lake_sites,
-         # model_id != 'climatology',
-         model_id %in% model_meta$model_id,
-         variable == 'oxygen') |>
-  collect()
-
-# Calculate skill scores ------------
-oxygen_climatology <- arrow::open_dataset("../scores_rescored") |>
-  filter(model_id == 'climatology', 
-         site_id %in% lake_sites,
-         variable == 'oxygen') |>
-  collect() |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  arrange(reference_datetime, datetime, site_id) |> 
-  pivot_wider(values_from = c(crps, logs),
-              names_from = model_id)
-
-oxygen_skill <- oxygen_scores |>
-  filter(model_id != 'climatology') |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  full_join(oxygen_climatology) |>
-
-  # calcaulte the skill relative to persistence (- is bad, + is good)
-  mutate(skill_crps = crps_climatology - crps,
-         skill_logs = logs_climatology - logs,
-         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-  # consistent forecast period
-  filter(horizon <= 30,
-         horizon >= 1,
-         model_id != 'fARIMA') |> 
-  arrange(reference_datetime, site_id, datetime, model_id)
-
-#--------------------------------------#
-```
-
-```{r plot-WQ, echo = F}
-
-top10_chla <- 
-  chla_skill |> 
-  group_by(model_id) |> 
-  summarise(median_crps = median(skill_crps, na.rm = T)) |> 
-  ungroup() |>
-  slice_max(median_crps, n= 10, na_rm = T) |>
-  left_join(model_meta) |> 
-  mutate(model_id = fct_reorder(model_id, median_crps),
-         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
-  ggplot(aes(x=median_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
-  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
-                   stat = "identity",
-                   alpha = 0.8, 
-                   pattern_fill = "black",
-                   colour = "black", 
-                   pattern_spacing = 0.04,
-                   pattern_alpha = 1) + 
-  scale_y_reordered() +
-  labs(y='Model ID', x = 'median CRPS skill (vs climatology)', title = 'chla') +
-  geom_vline(xintercept = 0) +
-  scale_fill_manual(values = cols_modeltype) +
-  scale_pattern_manual(values=c('none', 'stripe')) +
-  scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw() + 
-  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
-                                title = 'Uses weather covariates?'),
-         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
-
-all_chla <- chla_skill |> 
-  group_by(model_id) |> 
-  summarise(median_crps = median(skill_crps, na.rm = T)) |> 
-  left_join(model_meta) |> 
-  mutate(model_id = fct_reorder(model_id, median_crps),
-         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
-  ggplot(aes(x=median_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
-  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
-                   stat = "identity",
-                   alpha = 0.8, 
-                   pattern_fill = "black",
-                   colour = "black", 
-                   pattern_spacing = 0.03,
-                   pattern_alpha = 1) + 
-  scale_y_reordered() +
-  labs(y='Model ID', x = 'median CRPS skill (vs persistence)') +
-  geom_vline(xintercept = 0) +
-  scale_fill_manual(values = cols_modeltype) +
-  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses weather covariates?') +
-  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses weather covariates?') +
-  theme_bw() + 
-  guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses weather covariates?'),
-         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
-
-ggpubr::ggarrange(top10_chla, all_chla, nrow = 1, 
-                  common.legend = T,align = 'h', 
-                  legend = 'top', labels = 'auto')
-
-#---------------------------------------------------#
-
-top10_oxygen <- 
-  oxygen_skill |> 
-  group_by(model_id) |> 
-  summarise(median_crps = median(skill_crps, na.rm = T)) |> 
-  ungroup() |>
-  slice_max(median_crps, n= 10, na_rm = T) |>
-  left_join(model_meta) |> 
-  mutate(model_id = fct_reorder(model_id, median_crps),
-         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
-  ggplot(aes(x=median_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
-  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
-                   stat = "identity",
-                   alpha = 0.8, 
-                   pattern_fill = "black",
-                   colour = "black", 
-                   pattern_spacing = 0.04,
-                   pattern_alpha = 1) + 
-  scale_y_reordered() +
-  labs(y='Model ID', x = 'median CRPS skill (vs climatology)', title = 'oxygen') +
-  geom_vline(xintercept = 0) +
-  scale_fill_manual(values = cols_modeltype) +
-  scale_pattern_manual(values=c('none', 'stripe')) +
-  scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw() + 
-  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
-                                title = 'Uses weather covariates?'),
-         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
-
-all_oxygen <- oxygen_skill |> 
-  group_by(model_id) |> 
-  summarise(median_crps = median(skill_crps, na.rm = T)) |> 
-  left_join(model_meta) |> 
-  mutate(model_id = fct_reorder(model_id, median_crps),
-         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
-  ggplot(aes(x=median_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
-  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
-                   stat = "identity",
-                   alpha = 0.8, 
-                   pattern_fill = "black",
-                   colour = "black", 
-                   pattern_spacing = 0.03,
-                   pattern_alpha = 1) + 
-  scale_y_reordered() +
-  labs(y='Model ID', x = 'median CRPS skill (vs climatology)') +
-  geom_vline(xintercept = 0) +
-  scale_fill_manual(values = cols_modeltype) +
-  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses weather covariates?') +
-  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses weather covariates?') +
-  theme_bw() + 
-  guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses weather covariates?'),
-         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
-
-ggpubr::ggarrange(top10_oxygen, all_oxygen, nrow = 1, 
-                  common.legend = T, align = 'h', 
-                  legend = 'top', labels = 'auto')
-```

### 2024-02-05 04_manuscript_plots.Rmd
moving panels and other edits

@@
 cols_modeltype <- c('empirical' = "#482576FF",
-                    'multi-model ensemble' =  "#35608DFF",
+                    'MME' =  "#35608DFF",
                     'process' = "#21908CFF",
@@
   left_join(model_meta) |> 
-  mutate(model_id = fct_reorder(model_id, mean_crps),
-         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
-  ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
-  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
+  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'), 
+         uses_AT = str_detect(NOAA_var, 'air_temperature')) |> 
+  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +    # Different pattern for each group
+  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
                    stat = "identity",
@@
   scale_y_reordered() +
-  labs(y='Model', x = 'mean CRPS skill (vs climatology)') +
+  labs(y='Model', x = 'mean relative skill (°C)') +
   geom_vline(xintercept = 0) +
@@
   guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
-                                title = 'Uses weather covariates?'),
+                                title = 'Uses air temperature as a covariate?'),
          fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
@@
   left_join(model_meta) |> 
-  mutate(model_id = fct_reorder(model_id, mean_crps),
-         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no')) |> 
-  ggplot(aes(x=mean_crps, y = model_id, fill = model_type)) +    # Different pattern for each group
+  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'), 
+         uses_AT = str_detect(NOAA_var, 'air_temperature')) |> 
+  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +    # Different pattern for each group
 
-  geom_bar_pattern(aes(pattern = as.factor(uses_NOAA)),
+  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
                    stat = "identity",
@@
   scale_y_reordered() +
-  labs(y='Model', x = 'mean CRPS skill (vs climatology)') +
+  labs(y='Model', x = 'mean relative skill (°C)') +
   geom_vline(xintercept = 0) +
   scale_fill_manual(values = cols_modeltype) +
-  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses weather covariates?') +
-  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses weather covariates?') +
+  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses air temperature as a covariate?') +
+  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses air temperature as a covariate?') +
   theme_bw() + 
-  guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses weather covariates?'),
+  guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses air temperature as a covariate?'),
          fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
@@
 ## Figure 2
-```{r horizon-score}
+```{r skill-bias-sd}
 # top 10 average models
@@
   slice_max(mean_crps, n = 10, na_rm = T) |> 
-  distinct(model_id)
+  left_join(select(model_meta, model_id, manuscript_name)) |> 
+  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps, .desc = T)) |> 
+  distinct(manuscript_name, model_id)
 
-B <- 
+skill <- 
   temperature_skill |> 
-  filter(model_id %in% top_mods$model_id) |> 
+  filter(model_id %in% top_mods$model_id, 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
   group_by(model_id, horizon) |> 
   summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  left_join(select(model_meta, model_id, manuscript_name)) |> 
+  mutate(manuscript_name = factor(manuscript_name, levels = levels(top_mods$manuscript_name))) |> 
   ggplot(aes(x=horizon, 
-             y = mean_crps, colour = fct_reorder(model_id, mean_crps, .desc = T))) +
-  geom_hline(yintercept = 0) + 
+             y = mean_crps, colour = manuscript_name)) +
+  geom_hline(yintercept = 0, colour = 'grey40' ) + 
   geom_line()  +  
-  labs(y='mean CRPS skill (vs climatology)', x = 'Horizon (days)') +
+  labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
   scale_color_brewer(name = 'Model', palette = 'Paired') +
   theme_bw()
-
-ggpubr::ggarrange(B) |> 
-  ggsave(filename = 'Plots/Figure2.png', width = 15, height = 17, units = 'cm')
-```
-
-## Figure 3
-```{r bias-sd}
-# top 10 average models
-top_mods <- temperature_skill |> 
-  group_by(model_id) |> 
-  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
-  ungroup() |> 
-  slice_max(mean_crps, n = 10, na_rm = T) |> 
-  mutate(model_id = fct_reorder(model_id, mean_crps, .desc = T)) |> 
-  distinct(model_id)
 
@@
   summarise(mean_sd = mean(sd, na.rm = T)) |> 
-  mutate(model_id = factor(model_id, levels = levels(top_mods$model_id))) |> 
+  left_join(select(model_meta, model_id, manuscript_name)) |> 
+  mutate(manuscript_name = factor(manuscript_name, levels = levels(top_mods$manuscript_name))) |> 
   ggplot(aes(x=horizon, 
-             y = mean_sd, colour = model_id)) +
+             y = mean_sd, colour = manuscript_name)) +
   geom_line()+
-  labs(y='mean standard deviation', x = 'Horizon (days)') +
+  labs(y='mean standard deviation (°C)', x = 'Horizon (days)') +
   scale_color_brewer(name = 'Model', palette = 'Paired') +
@@
          horizon != 15,
-         horizon <=30) |>
+         horizon <= 30) |>
   group_by(model_id, horizon) |> 
   summarise(mean_abs_bias = mean(abs_bias, na.rm = T)) |> 
-  mutate(model_id = factor(model_id, levels = levels(top_mods$model_id))) |> 
+  left_join(select(model_meta, model_id, manuscript_name)) |> 
+  mutate(manuscript_name = factor(manuscript_name, levels = levels(top_mods$manuscript_name))) |> 
   ggplot(aes(x=horizon, 
-             y = mean_abs_bias, colour = model_id)) +
+             y = mean_abs_bias, colour = manuscript_name)) +
   geom_line()+
-  labs(y='mean absolute bias', x = 'Horizon (days)') +
+  labs(y='mean absolute bias (°C)', x = 'Horizon (days)') +
   scale_color_brewer(name = 'Model', palette = 'Paired') +
@@
 
-ggpubr::ggarrange(sd, bias, labels = 'auto', align = 'v', nrow = 2, common.legend = T, legend = 'right') |> 
-  ggsave(filename = 'Plots/Figure3.png', height = 15, width = 14, units = 'cm')
+ggpubr::ggarrange(skill, 
+                  ggarrange(as_ggplot(get_legend(bias)), NULL, widths = c(0.5, 1.5)),  
+                  sd, 
+                  bias, 
+                  legend = F, align = 'hv') |> 
+  ggsave(filename = 'Plots/Figure2.png', height = 16, width = 16, units = 'cm')
 ```
 
-## Figure 4
+## Figure 3
 
@@
   plot_reliability(scores = temperature_scores,
-                            models = top_mods$model_id, 
+                   models = top_mods, # model_ids and manuscript names
                    CI = 80, 
-                            order = levels(top_mods$model_id)) +
+                   order = levels(top_mods$manuscript_name)) +
   scale_color_brewer(name = 'Top Models', palette = 'Paired') +
@@
 good_95 <- plot_reliability(scores = temperature_scores,
-                            models = top_mods$model_id, 
+                            models = top_mods, 
                             CI = 95, 
-                            order = levels(top_mods$model_id)) +
+                            order = levels(top_mods$manuscript_name)) +
   scale_color_brewer(name = 'Top Models', palette = 'Paired') +
@@
 
+# get the list of bad models
 bad_mods <- model_meta |> 
   filter(!model_id %in% top_mods$model_id) |> 
-  pull(model_id)
+  select(model_id, manuscript_name)
 
@@
                    CI = 80, 
-                   order = bad_mods) +
+                   order = bad_mods$manuscript_name) +
   theme_bw() +
@@
                             CI = 95, 
-                            order = bad_mods) +
+                            order = bad_mods$manuscript_name) +
   theme_bw() +
@@
           ncol = 1, heights = c(5, 1)) |> 
-  ggsave(filename = 'Plots/Figure4.png', height = 17.6, width = 22, units = 'cm')
+  ggsave(filename = 'Plots/Figure3.png', height = 17.6, width = 23, units = 'cm')
 ```
 
-## Figure 5
+## Figure 4
 ```{r model_type-site}
@@
   summarise(n = n())  |> 
-  mutate(site_id = 'TOOK', mean_skill = -3.5) |> 
+  mutate(site_id = 'TOOK', mean_skill = -3.25) |> 
   filter(model_type != 'null')
@@
   geom_boxplot() +
-  facet_wrap(~model_type, scales = 'free_y',nrow = 4) +
-  labs(y='Site', x = 'mean CRPS skill (vs climatology)') +
-  scale_colour_manual(values = cols_modeltype) +
+  facet_wrap(~model_type, scales = 'free_y',nrow = 2) +
+  labs(y='Site', x = 'mean relative skill (°C)') +
+  scale_colour_manual(values = cols_modeltype, name = 'Model Type') +
   theme_bw()
@@
              y = mean_crps, colour = site_id)) +
-  geom_hline(yintercept = 0) + 
+  geom_hline(yintercept = 0, colour = 'grey40') + 
   geom_line()  +  
-  labs(y='mean CRPS skill (vs climatology)', x = 'Horizon (days)') +
-  scale_color_viridis_d(option = 'H') +
+  labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
+  scale_color_viridis_d(option = 'H', name  = 'Site') +
   theme_bw()
 
-ggarrange(A, B, widths = c(1,1)) |> 
-  ggsave(filename = './Plots/Figure5.png', width = 25, height = 15, units = 'cm')
+cowplot::plot_grid(A, B, nrow = 2, rel_heights = c(1.5, 1), align = 'v', axis = 'lr', labels = 'auto') |> 
+  ggsave(filename = './Plots/Figure4.png', width = 19, height = 16, units = 'cm')
 ```
 
-## Figure 6
+## Figure 5
 
@@
          year = year(datetime),
-         is_2023 = ifelse(year == 2023, 2023, 2022),
+         is_2023 = ifelse(year == 2023, '2023', 'other years' ),
          site_id = factor(site_id, levels = lake_sites)) |> 
   ggplot(aes(x=doy, y=observation, colour = is_2023)) +
-  geom_jitter(show.legend = F, aes(alpha = is_2023), size = 0.7, height = 0.2) +
+  geom_jitter(aes(alpha = is_2023), size = 0.7, height = 0.2) +
   facet_manual(~factor(site_id, levels = lake_sites),
                design = layout_design) +
-  scale_colour_viridis_c(option = 'mako', begin = 0.1, end = 0.8) +
+  scale_colour_viridis_d(option = 'mako', begin = 0.8, end = 0.1, name ='') +
   labs(x= 'Day of year', y = 'Observation (°C)') +
-  scale_alpha(range = c(0.5, 0.8))+
+  scale_alpha_manual(values = c(0.7, 0.5), name = '', guide = F)+
   scale_x_continuous(expand = c(0.01,0)) +
@@
                      limits = c(0,35)) +
-  theme_bw()
+  theme_bw() + 
+  theme(legend.position = c(0.75, 1))
 
@@
   ggplot(aes(y = site_id, x = cv, colour = is_2023)) +
-  geom_point() +
+  geom_point(show.legend = F) +
   scale_colour_viridis_d(option = 'mako', begin = 0.8, end = 0.1, name = '') +
@@
 
-ggarrange(obs, cv_obs, widths = c(1, 1.5)) |> 
-  ggsave(filename = 'Plots/Figure6.png', width = 20, height=10, units = 'cm')
+ggarrange(obs, cv_obs, widths = c(1, 1.5), labels = 'auto') |> 
+  ggsave(filename = 'Plots/Figure5.png', width = 20, height=10, units = 'cm')
 

### 2024-02-05 R/reliability_plot.R
use manuscript names

@@ -2,19 +2,19 @@ plot_reliability <- function(scores, models, CI, order = levels(models)) {
   
   if (CI == 95) {
     plot <- scores |> 
+      inner_join(models) |> 
       mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-      filter(model_id %in% models,
-             horizon >0,
+      filter(horizon >0,
              horizon != 15, 
              horizon <=30,
              !is.na(observation)) |> 
       mutate(is_in = between(observation, quantile02.5, quantile97.5)) |> 
-      group_by(horizon, is_in, model_id) |> 
+      group_by(horizon, is_in, model_id, manuscript_name) |> 
       summarise(n = n()) |> 
       pivot_wider(names_from = is_in, names_prefix = 'within_', values_from = n, values_fill = 0) |> 
       mutate(perc = within_TRUE/(within_FALSE + within_TRUE)*100, 
-             model_id = factor(model_id, levels = order)) |> 
-      ggplot(aes(x=horizon, y=perc, colour = model_id)) +
+             manuscript_name = factor(manuscript_name, levels = order)) |> 
+      ggplot(aes(x=horizon, y=perc, colour = manuscript_name)) +
       geom_hline(yintercept = 95, colour = 'grey3', linetype = 'dashed')+
       geom_line(alpha = 0.8) +
       labs(y = 'Percentage of observations within\n95% confidence intervals', x='Horizon (days)') +
@@ -23,19 +23,19 @@ plot_reliability <- function(scores, models, CI, order = levels(models)) {
   }
   if (CI == 80) {
     plot <- scores |> 
+      inner_join(models) |> 
       mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-      filter(model_id %in% models,
-             horizon >0,
+      filter(horizon >0,
              horizon != 15, 
              horizon <=30,
              !is.na(observation)) |> 
       mutate(is_in = between(observation, quantile10, quantile90)) |> 
-      group_by(horizon, is_in, model_id) |> 
+      group_by(horizon, is_in, model_id, manuscript_name) |> 
       summarise(n = n()) |> 
       pivot_wider(names_from = is_in, names_prefix = 'within_', values_from = n, values_fill = 0) |> 
       mutate(perc = within_TRUE/(within_FALSE + within_TRUE)*100, 
-             model_id = factor(model_id, levels = order)) |> 
[…]

### 2024-02-08 04_manuscript_plots.Rmd
update observation figure

@@
 temperature_scores <- arrow::open_dataset("./scores_2023") |>
+  dplyr::filter(!(model_id %in% c('flareGOTM','flareGOTM_noDA', 'flare_ler', 'flare_ler_baseline') &
+                  site_id == 'TOOK')) |> 
   filter(site_id %in% lake_sites,
@@
 # is 2023 a stand out year?
+targets_2023 <- targets |> 
+  filter(variable == 'temperature',
+         site_id %in% lake_sites, 
+         year(datetime) == 2023) |> 
+  mutate(month = as.factor(month(datetime))) |> 
+  group_by(month, site_id) |> 
+  summarise(mean_obs = mean(observation, na.rm = T),
+            n = n()) |> 
+  mutate(mean_obs = ifelse(n < 18, NA, mean_obs))
+
 obs <- targets |> 
@@
          site_id %in% lake_sites,
-           year(datetime) < 2024) |> 
-  mutate(doy = yday(datetime),
-         year = year(datetime),
-         is_2023 = ifelse(year == 2023, '2023', 'other years' ),
-         site_id = factor(site_id, levels = lake_sites)) |> 
-  ggplot(aes(x=doy, y=observation, colour = is_2023)) +
-  geom_jitter(aes(alpha = is_2023), size = 0.7, height = 0.2) +
+         year(datetime) < 2023) |> 
+  mutate(month = as.factor(month(datetime))) |> 
+  ggplot(aes(x = month, y= observation), colour = "#26172AFF") +
+  geom_boxplot() +
   facet_manual(~factor(site_id, levels = lake_sites),
                design = layout_design) +
-  scale_colour_viridis_d(option = 'mako', begin = 0.8, end = 0.1, name ='') +
-  labs(x= 'Day of year', y = 'Observation (°C)') +
-  scale_alpha_manual(values = c(0.7, 0.5), name = '', guide = F)+
-  scale_x_continuous(expand = c(0.01,0)) +
-  scale_y_continuous(expand = c(0,0),
-                     limits = c(0,35)) +
+  geom_point(data = targets_2023, aes(x = month, y = mean_obs), colour = '#60CEACFF') +
+  scale_x_discrete(name = 'Month', labels = month.abb) +
+  scale_y_continuous(name = 'Water temperature (°C)') +
   theme_bw() +
-  theme(legend.position = c(0.75, 1))
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
 
-cv_obs <- 
-  targets |> 
-  mutate(site_id = factor(site_id, levels = lake_sites), 
-         year = year(datetime)) |> 
-  filter(variable == 'temperature',
-         site_id %in% lake_sites,
-         year < 2024) |> 
-  group_by(site_id, year, variable) |> 
-  summarise(cv = sd(observation, na.rm = T)/ mean(observation, na.rm = T)) |> 
-  mutate(is_2023 = ifelse(year == 2023, '2023', 'other years' )) |> 
-  ggplot(aes(y = site_id, x = cv, colour = is_2023)) +
-  geom_point(show.legend = F) +
-  scale_colour_viridis_d(option = 'mako', begin = 0.8, end = 0.1, name = '') +
-  scale_y_discrete(limits = rev)+
-  theme_bw() +
-  labs(x = 'Coefficient of variation', y = 'Site')
-
-ggarrange(obs, cv_obs, widths = c(1, 1.5), labels = 'auto') |> 
-  ggsave(filename = 'Plots/Figure5.png', width = 20, height=10, units = 'cm')
+ggsave(obs, filename = 'Plots/Figure5.png', width = 10, height=13, units = 'cm')
 

### 2024-03-06 04_manuscript_plots.Rmd
change colours

@@
 ```{r read-in-data}
+lake_sites <- c('TOOK', 'PRLA','PRPO', 'CRAM', 'LIRO','BARC', 'SUGG')
 
-lake_sites <- c('TOOK', 'PRLA','PRPO', 'CRAM', 'LIRO','BARC', 'SUGG')
-cols_modeltype <- c('empirical' = "#482576FF",
-                    'MME' =  "#35608DFF",
-                    'process' = "#21908CFF",
-                    'ML' = '#BBDF27FF',
-                    'null' = '#A9A9A9')
 # read in the metadata from the google sheet
@@
 
+baselines <- model_meta |> 
+  filter(model_id %in% c('persistenceRW', 'climatology'))
 # Open dataset-------------
@@
   arrange(reference_datetime, site_id, datetime, model_id)
+
+source('R/get_colours.R') # returns model_id_colours and cols_modeltypes
+
 ```
@@
   ungroup() |>
-  slice_max(mean_crps, n= 10, na_rm = T) |>
+  filter(mean_crps > 0) |>
   left_join(model_meta) |> 
@@
                    pattern_fill = "black",
-                   colour = "black", 
+                   pattern_colour = "black", 
                    pattern_spacing = 0.04,
-                   pattern_alpha = 1) + 
+                   pattern_alpha = 1,
+                   colour = 'black') + 
   scale_y_reordered() +
@@
                    alpha = 0.8, 
+                   colour = 'black',
                    pattern_fill = "black",
-                   colour = "black", width = 0.8,
+                   pattern_colour = "black", width = 0.8,
                    pattern_spacing = 0.03,
@@
 ```{r skill-bias-sd}
-# top 10 average models
-top_mods <- temperature_skill |> 
-  group_by(model_id) |> 
-  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
-  ungroup() |> 
-  slice_max(mean_crps, n = 10, na_rm = T) |> 
-  left_join(select(model_meta, model_id, manuscript_name)) |> 
-  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps, .desc = T)) |> 
-  distinct(manuscript_name, model_id)
-
 skill <- 
   temperature_skill |> 
-  filter(model_id %in% top_mods$model_id, 
+  filter(model_id %in% c(top_mods$model_id) & model_id != 'persistenceRW', 
          horizon > 0, 
@@
   left_join(select(model_meta, model_id, manuscript_name)) |> 
-  mutate(manuscript_name = factor(manuscript_name, levels = levels(top_mods$manuscript_name))) |> 
+  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name))) |> 
   ggplot(aes(x=horizon, 
@@
   labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
-  scale_color_brewer(name = 'Model', palette = 'Paired') +
+  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
   theme_bw()
@@
   mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
-  filter(model_id %in% top_mods$model_id, 
+  filter(model_id %in% top_mods$model_id & model_id != 'persistenceRW',  
          horizon > 0, 
@@
   labs(y='mean standard deviation (°C)', x = 'Horizon (days)') +
-  scale_color_brewer(name = 'Model', palette = 'Paired') +
+  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
   theme_bw()
@@
          abs_bias = abs(mean - observation)) |> 
-  filter(model_id %in% top_mods$model_id, 
+  filter(model_id %in% top_mods$model_id & model_id != 'persistenceRW', 
          horizon > 0, 
@@
   labs(y='mean absolute bias (°C)', x = 'Horizon (days)') +
-  scale_color_brewer(name = 'Model', palette = 'Paired') +
+  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
   theme_bw() +
@@
                   bias, 
-                  legend = F, align = 'hv') |> 
-  ggsave(filename = 'Plots/Figure2.png', height = 16, width = 16, units = 'cm')
+                  legend = F, align = 'hv', 
+                  labels = c('a','', 'b', 'c')) |> 
+  ggsave(filename = 'Plots/Figure2.png', height = 17, width = 16, units = 'cm')
 ```
@@
                    CI = 80, 
-                   order = levels(top_mods$manuscript_name)) +
-  scale_color_brewer(name = 'Top Models', palette = 'Paired') +
+                   order = levels(model_id_colours$manuscript_name)) +
+  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
   theme_bw() +
-  scale_y_continuous(limits = c(0, 100))
+  scale_y_continuous(limits = c(45, 100))
 
@@
                             CI = 95, 
-                            order = levels(top_mods$manuscript_name)) +
-  scale_color_brewer(name = 'Top Models', palette = 'Paired') +
+                            order = levels(model_id_colours$manuscript_name)) +
+  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
   theme_bw() +
-  scale_y_continuous(limits = c(0, 100))
+  scale_y_continuous(limits = c(45, 100))
 
@@
     theme_bw() +
-  scale_colour_discrete(name = '') +
+    scale_colour_manual(name = '', values = rep('grey50', 20)) +
     scale_y_continuous(limits = c(0, 100)) + 
-  theme(legend.position = 'bottom') + 
-  guides(colour = guide_legend(nrow = 5))
+    theme(legend.position = 'none')
 
@@
   theme_bw() +
-  scale_y_continuous(limits = c(0, 100))
+  scale_y_continuous(limits = c(0, 100))  +
+  scale_colour_manual(name = '', values = rep('grey50', 20))
 
-ggarrange(ggarrange(ggarrange(get_legend(good_80), NULL, nrow = 2),
-          ggarrange(good_80, good_95, bad_80, bad_95, align = 'hv', legend= 'none'), ncol = 2, 
-          widths = c(1,5)),
-          ggarrange(NULL, get_legend(bad_80), ncol = 2, widths = c(1,5)),
-          ncol = 1, heights = c(5, 1)) |> 
+ggarrange(ggarrange(get_legend(good_80), NULL, nrow = 2),
+          ggarrange(good_80, good_95, bad_80, bad_95, align = 'hv', legend= 'none', labels = 'auto'), 
+          ncol = 2, 
+          widths = c(1,5)) |> 
   ggsave(filename = 'Plots/Figure3.png', height = 17.6, width = 23, units = 'cm')
@@
   ungroup() |> 
-  slice_max(mean_crps, n = 10, na_rm = T) |> 
+  filter(mean_crps > 0) |> 
   distinct(model_id)
@@
   labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
-  scale_color_viridis_d(option = 'H', name  = 'Site') +
+  scale_colour_brewer(palette = 'Dark2')+
   theme_bw()

### 2024-03-06 R/reliability_plot.R
increase size of text annotation

@@ -2,7 +2,7 @@ plot_reliability <- function(scores, models, CI, order = levels(models)) {
   
   if (CI == 95) {
     plot <- scores |> 
-      inner_join(models) |> 
+      inner_join(models, by = 'model_id') |> 
       mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
       filter(horizon >0,
              horizon != 15, 
@@ -17,13 +17,13 @@ plot_reliability <- function(scores, models, CI, order = levels(models)) {
       ggplot(aes(x=horizon, y=perc, colour = manuscript_name)) +
       geom_hline(yintercept = 95, colour = 'grey3', linetype = 'dashed')+
       geom_line(alpha = 0.8) +
-      labs(y = 'Percentage of observations within\n95% confidence intervals', x='Horizon (days)') +
-      annotate('text', x = 25, y = 100, label = 'underconfident') +
-      annotate('text', x = 25, y = 75, label = 'overconfident')
+      labs(y = 'Percentage of observations within\n95% confidence intervals', x='Horizon (days)')  +
+      annotate('text', x = 30, y = 100, label = 'underconfident', size = 5, hjust = 1) +
+      annotate('text', x = 30, y = 75, label = 'overconfident', size = 5, hjust = 1)
   }
   if (CI == 80) {
     plot <- scores |> 
-      inner_join(models) |> 
+      inner_join(models, by = 'model_id') |> 
       mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
       filter(horizon >0,
              horizon != 15, 
@@ -39,8 +39,8 @@ plot_reliability <- function(scores, models, CI, order = levels(models)) {
       geom_hline(yintercept = 80, colour = 'grey3', linetype = 'dashed')+
       geom_line(alpha = 0.8) +
       labs(y = 'Percentage of observations within\n80% confidence intervals', x='Horizon (days)') +
-      annotate('text', x = 25, y = 100, label = 'underconfident') +
-      annotate('text', x = 25, y = 75, label = 'overconfident')
+      annotate('text', x = 30, y = 100, label = 'underconfident', size = 5, hjust = 1) +
+      annotate('text', x = 30, y = 75, label = 'overconfident', size = 5, hjust = 1)
   }
   if (CI != 80 & CI != 95) {
     message('CI must be 80 or 95')

### 2024-03-06 04_manuscript_plots.Rmd
type  --> class

@@
                                 title = 'Uses air temperature as a covariate?'),
-         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
+         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class'))
 
@@
   guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses air temperature as a covariate?'),
-         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model type'))
+         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class'))
 
@@
   theme_bw() +
-  coord_cartesian(ylim = c(0, 2.5))
+  coord_cartesian(ylim = c(0, 2))
 
@@
   geom_vline(xintercept = 0, linetype = 'solid') +
-  geom_label(data = n_model_type, aes(label = paste0('n = ', n)))+ 
+  geom_text(data = n_model_type, aes(label = paste0('n = ', n)))+ 
   geom_boxplot() +
@@
   labs(y='Site', x = 'mean relative skill (°C)') +
-  scale_colour_manual(values = cols_modeltype, name = 'Model Type') +
+  scale_colour_manual(values = cols_modeltype, name = 'Model class') +
   theme_bw()
@@
   labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
-  scale_colour_brewer(palette = 'Dark2')+
+  scale_colour_brewer(palette = 'Dark2', name = 'Site')+
   theme_bw()

### 2024-03-19 04b_supplementary_plots.Rmd
SI plots

---
title: "supplementary plots"
output: html_document
date: "2024-02-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(arrow)
library(lubridate)
library(tidytext)
library(ggpattern)
library(ggpubr)
library(ggh4x)
```

## Read in the data and model metadata
```{r read-in-data}
targets <- readr::read_csv("https://data.ecoforecast.org/neon4cast-targets/aquatics/aquatics-targets.csv.gz") |> 
  na.omit()

lake_sites <- c('TOOK', 'PRLA','PRPO', 'CRAM', 'LIRO','BARC', 'SUGG')
cols_modeltype <- c('empirical' = "#482576FF",
                    'MME' =  "#35608DFF",
                    'process' = "#21908CFF",
                    'ML' = '#BBDF27FF',
                    'null' = '#A9A9A9')
# read in the metadata from the google sheet
googlesheets4::gs4_deauth()
model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")

# Open dataset-------------
# start with lake temperatures
temperature_scores <- arrow::open_dataset("./scores_2023") |>
  dplyr::filter(!(model_id %in% c('flareGOTM','flareGOTM_noDA', 'flare_ler', 'flare_ler_baseline') &
                  site_id == 'TOOK')) |> 
  filter(site_id %in% lake_sites,
         model_id %in% model_meta$model_id,
         variable == 'temperature') |>
  collect()

# Calculate skill scores ------------
temperature_climatology <- arrow::open_dataset("./scores_2023") |>
  filter(model_id == 'climatology', 
         site_id %in% lake_sites,
         variable == 'temperature') |>
  collect() |> 
  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
  arrange(reference_datetime, datetime, site_id) |> 
  pivot_wider(values_from = c(crps, logs),
              names_from = model_id)

temperature_skill <- temperature_scores |>
  filter(model_id != 'climatology') |> 
  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
  full_join(temperature_climatology) |>

  # calcaulte the skill relative to climatology (- is bad, + is good)
  mutate(skill_crps = crps_climatology - crps,
         skill_logs = logs_climatology - logs,
         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  # consistent forecast period
  filter(horizon <= 30,
         horizon >= 1, 
         horizon != 15) |> 
  arrange(reference_datetime, site_id, datetime, model_id) 
```
## Figure s1 - targets timeseries

```{r}
targets_ts <- targets |> 
  filter(year(datetime) < 2023) |> 
  ggplot(aes(x=datetime, y=variable, colour = variable)) +
  geom_point(show.legend = F) +
  facet_wrap(~site_id, ncol = 5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave(plot = targets_ts, 'Plots/FigureS1.png', width = 20, height = 15, units = 'cm')
```

## Figure S2 - all models, skill over horizon

```{r}
horizon_skill <- 
  temperature_skill |> 
  filter(!(model_id %in% c(top_mods$model_id)), 
         horizon > 0, 
         horizon != 15,
         horizon <=30) |> 
  group_by(model_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  left_join(select(model_meta, model_id, manuscript_name)) |> 
  # mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name))) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = manuscript_name, linetype = manuscript_name)) +
  geom_hline(yintercept = 0, colour = 'grey40' ) + 
  geom_line(show.legend = F)  +  
  labs(y='Mean relative skill (°C)', x = 'Horizon (days)') +
  theme_bw()

ggsave(plot = horizon_skill, 'Plots/FigureS2.png', width = 10, height = 10, units = 'cm')
```

## Figure S3 - performance of null models

```{r}
baseline_forecasts <- arrow::open_dataset("scores_2023") |>
  filter(site_id %in% lake_sites,
         model_id %in% c('persistenceRW', 'climatology')) |>
  collect()

baseline_forecasts |> 
  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  group_by(site_id, model_id, horizon) |> 
  summarise(crps_median = median(crps, na.rm = T), .groups = 'drop') |> 
  ggplot(aes(x=horizon, y=crps_median, colour = site_id)) +
  geom_line() +
  facet_wrap(~model_id)

baseline_forecasts |> 
  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
  group_by(site_id, model_id, reference_datetime) |> 
  summarise(crps_median = median(crps, na.rm = T), .groups = 'drop') |> 
  ggplot(aes(x = as_date(reference_datetime), y=crps_median, colour = site_id)) +
  geom_line() +
  facet_wrap(~model_id, scale = 'free_y')

```

## Figure S4

Plotting only dates where all sites are evaluated shows the same pattern as when all dates are considered. 
```{r overlapping-periods}
n_model_type <- temperature_skill |> 
  distinct(model_id) |> 
  left_join(model_meta) |>
  group_by(model_type) |> 
  summarise(n = n())  |> 
  mutate(site_id = 'TOOK', mean_skill = -3.25) |> 
  filter(model_type != 'null')

# top 10 average models
top_mods <- temperature_skill |> 
  group_by(model_id) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  ungroup() |> 
  slice_max(mean_crps, n = 10, na_rm = T) |> 
  distinct(model_id)

# what duration have the same dates, including TOOK
date_same_A <-  temperature_skill |> 
  na.omit() |> 
  group_by(site_id) |> 
  summarise(first_date = min(datetime), 
            last_date = max(datetime)) |> 
  summarise(first_date = max(first_date),
            last_date = min(last_date))

# what duration have the same dates, including TOOK
date_same_B <- temperature_skill |> 
  filter(site_id != 'TOOK') |> 
  na.omit()  |> 
  group_by(site_id) |> 
  summarise(first_date = min(datetime), 
            last_date = max(datetime)) |> 
  summarise(first_date = max(first_date),
            last_date = min(last_date))

A <- temperature_skill |> 
  left_join(model_meta, by = 'model_id') |>
  filter(model_type != 'null', 
         between(datetime, date_same_B$first_date, date_same_B$last_date)) |> 
  group_by(model_type, model_id, site_id) |> 
  summarise(mean_skill = mean(skill_crps, na.rm = T)) |>  
  ggplot(aes(x = mean_skill, 
             y = site_id, 
             colour = model_type)) +
  geom_vline(xintercept = 0, linetype = 'solid') +
  geom_label(data = n_model_type, aes(label = paste0('n = ', n)))+ 
  geom_boxplot() +
  facet_wrap(~model_type, scales = 'free_y',nrow = 2) +
  labs(y='Site', x = 'mean relative skill (°C)') +
  scale_colour_manual(values = cols_modeltype, name = 'Model Type') +
  theme_bw()

B <- temperature_skill |> 
  filter(model_id %in% top_mods$model_id, 
         between(datetime, date_same_B$first_date, date_same_B$last_date)) |> 
  group_by(site_id, horizon) |> 
  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
  ggplot(aes(x=horizon, 
             y = mean_crps, colour = site_id)) +
  geom_hline(yintercept = 0, colour = 'grey40') + 
  geom_line()  +  
  labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
  scale_color_viridis_d(option = 'H', name  = 'Site') +
  theme_bw()

cowplot::plot_grid(A, B, nrow = 2, rel_heights = c(1.5, 1), align = 'v', axis = 'lr', labels = 'auto') |> 
  ggsave(filename = './Plots/FigureS4.png', width = 19, height = 16, units = 'cm')
```

### 2024-03-19 04_manuscript_plots.Rmd
update axis labels

@@
   scale_y_reordered() +
-  labs(y='Model', x = 'mean relative skill (°C)') +
+  labs(y='Model', x = 'Mean relative skill (°C)') +
   geom_vline(xintercept = 0) +
@@
   geom_line()  +  
-  labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
+  labs(y='Mean relative skill (°C)', x = 'Horizon (days)') +
   scale_colour_manual(name = 'Model',values = model_id_colours$value) +
@@
   geom_line()+
-  labs(y='mean standard deviation (°C)', x = 'Horizon (days)') +
+  labs(y='Mean standard deviation (°C)', x = 'Horizon (days)') +
   scale_colour_manual(name = 'Model',values = model_id_colours$value) +
@@
   geom_line()+
-  labs(y='mean absolute bias (°C)', x = 'Horizon (days)') +
+  labs(y='Mean absolute bias (°C)', x = 'Horizon (days)') +
   scale_colour_manual(name = 'Model',values = model_id_colours$value) +
@@
   facet_wrap(~model_type, scales = 'free_y',nrow = 2) +
-  labs(y='Site', x = 'mean relative skill (°C)') +
+  labs(y='Site', x = 'Mean relative skill (°C)') +
   scale_colour_manual(values = cols_modeltype, name = 'Model class') +
@@
   geom_line()  +  
-  labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
+  labs(y='Mean relative skill (°C)', x = 'Horizon (days)') +
   scale_colour_brewer(palette = 'Dark2', name = 'Site')+

### 2024-03-19 01a_inventory.R
inventory for wt models

@@ -2,16 +2,20 @@ library(tidyverse)
 # inventory
 googlesheets4::gs4_deauth()
 model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing") |> 
-  mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0)) |> 
-  filter(!is.na(model_type))
+  mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0)) 
+
+
+# water temperature models only
+wt_meta <- model_meta |>
+  filter(str_detect(variable, 'temperature'))
 
 # types of uncertainty represented
 # Number of models that contain each type
-model_meta |> 
+wt_meta |> 
   summarise(across(contains('unc'), ~ sum(.x, na.rm = T)))
 
 # How many sources are represented in each model?
-model_meta |> 
+wt_meta |> 
   select(model_id, contains('unc_')) |> 
   mutate(unc_n = rowSums(pick(where(is.numeric)))) |> 
   group_by(unc_n) |> 
@@ -27,7 +31,7 @@ NOAA_vars <- c('air_temperature',
                'eastward_wind',
                'northward_wind')
 
-NOAA_summary <- model_meta |> 
+NOAA_summary <- wt_meta |> 
   select(model_id, variable) 
 
 
@@ -36,7 +40,7 @@ for (i in 1:length(NOAA_vars)) {
   
   NOAA_summary <- 
     NOAA_summary |> 
-    mutate(detect_var = str_detect(model_meta$NOAA_var, detect_var)) 
+    mutate(detect_var = str_detect(wt_meta$NOAA_var, detect_var)) 
   
   colnames(NOAA_summary)[2 + i] <- detect_var
 }
@@ -52,22 +56,13 @@ NOAA_summary |>
   group_by(variable, vars_n) |> 
   summarise(n = n())
 
-# Aspects of DA
-# do the forecasts contain initial conditions?
-model_meta |> 
-  summarise(initial_cond = sum(initial_cond, na.rm = T))
-
-# do the forecast models update parameters?
-model_meta |> 
-  summarise(updates_parameters = sum(updates_parameters, na.rm = T))
-
-# is the forecast 'dynamic'?
-model_meta |> 
-  summarise(dynamic = sum(dynamic, na.rm = T))
 
[…]

### 2024-03-19 R/get_colours.R
helper script to get colour palettes set up for plotting

@@ -1,0 +1,45 @@
+# top 10 average models
+top_mods <- temperature_skill |> 
+  group_by(model_id) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  ungroup() |> 
+  filter(mean_crps > 0) |> 
+  left_join(select(model_meta, model_id, manuscript_name, model_type)) |> 
+  arrange(model_type) |> 
+  # also include the baseline models
+  full_join(select(baselines, model_id, manuscript_name, model_type)) 
+
+
+# define the colours
+top_process <- top_mods |> filter(model_type == 'process') |> pull(model_id)
+process_cols <- RColorBrewer::brewer.pal(n = 9, name = 'YlGnBu')[3:6]
+names(process_cols) <- top_process
+
+top_MME <- top_mods |> filter(model_type == 'MME') |> pull(model_id)
+MME_cols <- RColorBrewer::brewer.pal(n = 9, name = 'YlOrRd')[c(4,5)]
+names(MME_cols) <- top_MME
+
+
+top_empirical <- top_mods |> filter(model_type == 'empirical') |> pull(model_id)
+empirical_cols <- RColorBrewer::brewer.pal(n = 9, name = 'Reds')[6]
+names(empirical_cols) <- top_empirical
+
+
+top_ML <- top_mods |> filter(model_type == 'ML') |> pull(model_id)
+ML_cols <- RColorBrewer::brewer.pal(n = 6, name = 'Purples')[c(2,4,6)]
+names(ML_cols) <- top_ML
+
+# make into a dataframe
+model_id_colours <- enframe(c(process_cols, MME_cols, empirical_cols, ML_cols)) |> 
+  full_join(top_mods, by = join_by(name == model_id)) |> 
+  mutate(value = ifelse(is.na(value), 'grey20', value),
+         mean_crps = ifelse(is.na(mean_crps), 0, mean_crps),
+         manuscript_name = fct_reorder(manuscript_name, mean_crps, .desc = T)) |> 
+  arrange(desc(mean_crps))
+
+# colours for model types 
+cols_modeltype <- c('empirical' = "#C42503FF" ,
+                    'MME' = "#FDAD35FF"  ,
+                    'process' = "#23C3E4FF" ,
+                    'ML' = "#3E378FFF",
+                    'null' = '#A9A9A9')

### 2024-03-19 04_manuscript_plots.Rmd
add forecast summary

@@
 
+```
+
+## Summary stats
+```{r}
+# total forecasts
+distinct_forecasts <- arrow::open_dataset("./scores_2023") |> 
+  distinct(model_id, site_id, variable, reference_datetime) |> 
+  collect()
+
+# variable forecasts
+distinct_forecasts |> 
+  group_by(variable ) |>
+  summarise(n = n()) |> 
+  mutate(percent = 100*(n/nrow(distinct_forecasts)))
+
+# lake temp forecasts
+distinct_forecasts |> 
+  filter(site_id %in% lake_sites,
+         variable == 'temperature') |> 
+  summarise(n = n()) 
 ```

### 2024-04-03 04_manuscript_plots.Rmd
add DO and chl to Fig2, update Figure 5 to deltas

--- 2024-03-19 04_manuscript_plots.Rmd
@@
   filter(model_id %in% c('persistenceRW', 'climatology'))
-# Open dataset-------------
-# start with lake temperatures
-temperature_scores <- arrow::open_dataset("./scores_2023") |>
-  dplyr::filter(!(model_id %in% c('flareGOTM','flareGOTM_noDA', 'flare_ler', 'flare_ler_baseline') &
-                  site_id == 'TOOK')) |> 
-  filter(site_id %in% lake_sites,
-         model_id %in% model_meta$model_id,
-         variable == 'temperature') |>
-  collect()
-
-# Calculate skill scores ------------
-temperature_climatology <- arrow::open_dataset("./scores_2023") |>
-  filter(model_id == 'climatology', 
-         site_id %in% lake_sites,
-         variable == 'temperature') |>
-  collect() |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  arrange(reference_datetime, datetime, site_id) |> 
-  pivot_wider(values_from = c(crps, logs),
-              names_from = model_id)
-
-temperature_skill <- temperature_scores |>
-  filter(model_id != 'climatology') |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  full_join(temperature_climatology) |>
-
-  # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = crps_climatology - crps,
-         skill_logs = logs_climatology - logs,
-         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-  # consistent forecast period
-  filter(horizon <= 30,
-         horizon >= 1, 
-         horizon != 15) |> 
-  arrange(reference_datetime, site_id, datetime, model_id)
+source('get_data.R')
 
@@
 
-## Figure 1
+## Figure 2
 
@@
   scale_y_reordered() +
-  labs(y='Model', x = 'mean relative skill (°C)') +
+  labs(y='Model', x = 'Mean relative skill (°C)', subtitle = 'Water temperature') +
   geom_vline(xintercept = 0) +
@@
   ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +   # Different pattern for each group
-
+  annotate('rect', xmax = Inf, xmin = -Inf, ymin = 19.5, ymax =  Inf, fill = 'grey', alpha = 0.2) + 
   geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
@@
   scale_y_reordered() +
-  labs(y='Model', x = 'Mean relative skill (°C)') +
+  labs(y='Model', x = 'Mean relative skill (°C)', subtitle = 'Water temperature') +
   geom_vline(xintercept = 0) +
@@
                   legend = 'top', labels = 'auto') |> 
-  ggsave(filename = './Plots/Figure1.png', width = 25, height = 10, units = 'cm')
+  ggsave(filename = './Plots/Figure2.png', width = 25, height = 11, units = 'cm')
+
+all_DO <- oxygen_skill |> 
+  group_by(model_id) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  left_join(model_meta) |> 
+  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'), 
+         uses_AT = str_detect(NOAA_var, 'air_temperature')) |> 
+  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +    # Different pattern for each group
+  annotate('rect', xmax = Inf, xmin = -Inf, ymin = 9.5, ymax =  Inf, fill = 'grey', alpha = 0.2) + 
+  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
+                   stat = "identity",
+                   alpha = 0.8, 
+                   pattern_fill = "black",
+                   pattern_colour = "black", 
+                   pattern_spacing = 0.04,
+                   pattern_alpha = 1,
+                   colour = 'black') + 
+  scale_y_reordered() +
+  labs(y='Model', x = 'Mean relative skill (mg/L)', subtitle = 'Dissolved oxygen') +
+  geom_vline(xintercept = 0) +
+  scale_fill_manual(values = cols_modeltype) +
+  scale_pattern_manual(values=c('none', 'stripe')) +
+  scale_pattern_type_manual(values=c(NA, NA)) +
+  theme_bw() +
+  theme(legend.position = 'none')
+
+all_chla <- chla_skill |> 
+  group_by(model_id) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  left_join(model_meta) |> 
+  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
+         ast = ifelse(mean_crps < -100, '*', ''),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'), 
+         uses_AT = as.factor(str_detect(NOAA_var, 'air_temperature')),
+         mean_crps = ifelse(mean_crps < -100, -100, mean_crps)) |> 
+  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +
+  annotate('rect', xmax = Inf, xmin = -Inf, ymin = 19.5, ymax =  Inf, fill = 'grey', alpha = 0.2) + 
+  geom_bar_pattern(aes(pattern = uses_AT),# Different pattern for each group
+                   stat = "identity",
+                   alpha = 0.8, 
+                   colour = 'black',
+                   pattern_fill = "black",
+                   pattern_colour = "black", width = 0.8,
+                   pattern_spacing = 0.03,
+                   pattern_alpha = 1) + 
+  scale_y_reordered() +
+  labs(y='Model', x = 'Mean relative skill (ug/L)', subtitle = 'Chlorophyll a') +
+  geom_vline(xintercept = 0) +
+  scale_fill_manual(values = cols_modeltype) +
+  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses air temperature as a covariate?') +
+  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses air temperature as a covariate?') +
+  theme_bw()+
+  theme(legend.position = 'none')  +
+  coord_cartesian(xlim = c(-10, 5)) +
+  geom_text(aes(y=manuscript_name, x= 0.5, label = ast), inherit.aes = F)
+
+ggpubr::ggarrange(top10_temp, all_temp, all_DO, all_chla, nrow = 2, ncol= 2, align = 'hv',
+                  common.legend = T,
+                  legend = 'top', labels = 'auto') |> 
+  ggsave(filename = 'Plots/test.png', height = 20, width = 25, units = 'cm')
 
@@
 
-## Figure 2
+## Figure 3
 ```{r skill-bias-sd}
@@
                   labels = c('a','', 'b', 'c')) |> 
-  ggsave(filename = 'Plots/Figure2.png', height = 17, width = 16, units = 'cm')
+  ggsave(filename = 'Plots/Figure3.png', height = 17, width = 16, units = 'cm')
 ```
 
-## Figure 3
+## Figure 4
 
@@
   theme_bw() +
-  scale_y_continuous(limits = c(45, 100))
+  scale_y_continuous(limits = c(54, 100))
 
@@
           widths = c(1,5)) |> 
-  ggsave(filename = 'Plots/Figure3.png', height = 17.6, width = 23, units = 'cm')
+  ggsave(filename = 'Plots/Figure4.png', height = 17.6, width = 23, units = 'cm')
 ```
 
-## Figure 4
+## Figure 5
 ```{r model_type-site}
@@
 cowplot::plot_grid(A, B, nrow = 2, rel_heights = c(1.5, 1), align = 'v', axis = 'lr', labels = 'auto') |> 
-  ggsave(filename = './Plots/Figure4.png', width = 19, height = 16, units = 'cm')
+  ggsave(filename = './Plots/Figure5.png', width = 19, height = 16, units = 'cm')
 ```
 
-## Figure 5
+## Figure 6
 
@@
 targets <- readr::read_csv("https://data.ecoforecast.org/neon4cast-targets/aquatics/aquatics-targets.csv.gz") |> 
-  na.omit()
+  na.omit() |> 
+  filter(year(datetime) <= 2023)
 
@@
 
-obs <- targets |> 
+obs_diff <-  
+  targets |> 
   filter(variable == 'temperature',
-         site_id %in% lake_sites,
-         year(datetime) < 2023) |> 
-  mutate(month = as.factor(month(datetime))) |> 
-  ggplot(aes(x = month, y= observation), colour = "#26172AFF") +
-  geom_boxplot() +
-  facet_manual(~factor(site_id, levels = lake_sites),
-               design = layout_design) +
-  geom_point(data = targets_2023, aes(x = month, y = mean_obs), colour = '#60CEACFF') +
-  scale_x_discrete(name = 'Month', labels = month.abb) +
-  scale_y_continuous(name = 'Water temperature (°C)') +
+         site_id %in% lake_sites) |> 
+  mutate(month = month(datetime),
+         period = ifelse(year(datetime) == 2023, 'focal', 'historic'),
+        site_id = factor(site_id, levels = lake_sites)) |> 
+  group_by(month, site_id, period) |> 
+  summarise(median_obs = median(observation, na.rm = T),
+            n = n()) |> 
+  mutate(median_obs = ifelse(n < 18, NA, median_obs)) |> 
+  pivot_wider(names_from = period, values_from = median_obs, id_cols = c(month, site_id)) |> 
+  mutate(delta = focal - historic,
+         month = as_date(paste('2023', month, '01', sep = '-'))) |> 
+  ggplot()  +
+  geom_hline(yintercept = 0) +
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
+  annotate("rect", 
+           xmax = rep(as_date('2024-01-01'),8), 
+           xmin = rep(as_date('2022-01-01'), 8),
+           ymin = c(-Inf, -Inf, -Inf, -Inf,Inf,Inf,Inf,Inf), 
+           ymax = c(-4,-3,-2,-1,1,2,3,4), 
+           fill = 'grey', alpha = 0.15) +
+  geom_line(aes(x=month, y=delta, colour = site_id)) +
+  geom_point(aes(x=month, y=delta, colour = site_id)) +
+  scale_x_date(date_labels = "%b", breaks = '1 month') +
+  labs(y = 'Difference in median monthly water temperature (°C)', 
+       x = 'Month') +
+  coord_cartesian(xlim = c(as_date('2023-01-01'), 
+                           as_date('2023-12-01')),
+                  ylim = c(-2, 5.2),
+                  expand = T) +
+  theme(panel.grid.minor.x = element_blank()) +
+  scale_colour_brewer(palette = 'Dark2', name = 'Site')
 
-ggsave(obs, filename = 'Plots/Figure5.png', width = 10, height=13, units = 'cm')
+ggsave(obs_diff, filename = 'Plots/Figure6.png', width = 13, height=10, units = 'cm')
 
 ```
-
-```{r}
-targets |> 
-  filter(variable == 'temperature',
-         site_id %in% lake_sites, 
-         year(datetime) < 2024) |> 
-  mutate(doy = yday(datetime),
-         year = year(datetime),
-         site_id = factor(site_id, levels = lake_sites), 
-         is_2023 = ifelse(year == 2023, T, F))  |> 
-  group_by(site_id, is_2023) |> 
-  summarise(mean = mean(observation),
-            sd = sd(observation)) |> 
-  filter(site_id  == 'BARC')
-```

### 2024-04-03 04b_supplementary_plots.Rmd
baseline forecast with rigt colours

@@
 
-baseline_forecasts |> 
-  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+baseline1<-baseline_forecasts |> 
+  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime)),
+         site_id = factor(site_id, levels = lake_sites)) |> 
   group_by(site_id, model_id, horizon) |> 
@@
   geom_line() +
-  facet_wrap(~model_id)
+  facet_wrap(~model_id) +
+  scale_colour_brewer(palette = 'Dark2', name = 'Site') +
+  labs(y='Median CRPS (°C)', x = 'Horizon (days)') +
+  theme_bw()
 
-baseline_forecasts |> 
+baseline2<-baseline_forecasts |> 
   mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
@@
   geom_line() +
-  facet_wrap(~model_id, scale = 'free_y')
+  facet_wrap(~model_id, scale = 'free_y') +
+  scale_colour_brewer(palette = 'Dark2', name = 'Site') +
+  labs(y='Median CRPS (°C)', x = 'Forecast date') +
+  theme_bw()
 
+ggarrange(baseline1, baseline2, nrow = 2) |> 
+  ggsave(filename = './Plots/FigureS3.png', width = 19, height = 16, units = 'cm')
 ```
@@
   labs(y='mean relative skill (°C)', x = 'Horizon (days)') +
-  scale_color_viridis_d(option = 'H', name  = 'Site') +
+  scale_colour_brewer(palette = 'Dark2', name = 'Site') +
   theme_bw()

### 2024-04-04 04_manuscript_plots.Rmd
minor edits

@@
   mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
-         ast = ifelse(mean_crps < -100, '*', ''),
+         ast = ifelse(mean_crps < -100, '>>>', ''),
          uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'), 
@@
   scale_colour_manual(values = cols_modeltype, name = 'Model class') +
-  theme_bw()
+  theme_bw()  +
+  theme(legend.position = 'bottom')
 
@@
 
-cowplot::plot_grid(A, B, nrow = 2, rel_heights = c(1.5, 1), align = 'v', axis = 'lr', labels = 'auto') |> 
+cowplot::plot_grid(A, B, nrow = 2, rel_heights = c(1.5, 1), align = 'v', axis = 'tlbr', labels = 'auto') |> 
   ggsave(filename = './Plots/Figure5.png', width = 19, height = 16, units = 'cm')

### 2024-04-11 04b_supplementary_plots.Rmd

@@
 targets_ts <- targets |> 
-  filter(year(datetime) < 2023) |> 
+  filter(year(datetime) < 2023,
+         site_id %in% lake_sites) |> 
   ggplot(aes(x=datetime, y=variable, colour = variable)) +
   geom_point(show.legend = F) +
-  facet_wrap(~site_id, ncol = 5) +
+  facet_wrap(~site_id, ncol = 3) +
   theme_bw() +

### 2024-04-11 01a_inventory.R
inventory all models

@@ -4,22 +4,60 @@ googlesheets4::gs4_deauth()
 model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing") |> 
   mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0)) 
 
+lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
+# How many models?
+model_meta |>
+  mutate(tw = str_detect(variable, 'temperature'),
+         do = str_detect(variable, 'oxygen'),
+         chla = str_detect(variable, 'chla')) |> 
+  summarise(sum(tw), sum(do), sum(chla))
 
-# water temperature models only
-wt_meta <- model_meta |>
-  filter(str_detect(variable, 'temperature'))
+model_meta |> 
+  group_by(variable) |> 
+  summarise(n())
 
-# types of uncertainty represented
-# Number of models that contain each type
-wt_meta |> 
-  summarise(across(contains('unc'), ~ sum(.x, na.rm = T)))
 
-# How many sources are represented in each model?
-wt_meta |> 
-  select(model_id, contains('unc_')) |> 
-  mutate(unc_n = rowSums(pick(where(is.numeric)))) |> 
-  group_by(unc_n) |> 
-  summarise(n = n())
+# model classes by variable
+
+model_meta |> 
+  filter(str_detect(variable, 'temperature')) |> 
+  group_by(model_type) |> 
+  summarise(n())
+
+model_meta |> 
+  filter(str_detect(variable, 'oxygen')) |> 
+  group_by(model_type) |> 
+  summarise(n())
+
+model_meta |> 
+  filter(str_detect(variable, 'chla')) |> 
+  group_by(model_type) |> 
+  summarise(n())
+
+
+# How many forecasts?
+all_lake_forecasts <- arrow::open_dataset('scores_2023') |> 
+  filter(model_id %in% model_meta$model_id,
+         site_id %in% lake_sites) |> 
+  distinct(reference_datetime, variable, 
+           site_id, model_id) |> 
+  collect()
+
+all_lake_forecasts |> 
+  filter(variable == 'temperature') |> 
+  summarise(n())
+
+# use AT as a covariate?
+model_meta |> 
+  select(model_id, NOAA_var, variable) |>
[…]

### 2024-04-11 01_get_model_list.R
update database only lakes

@@ -15,13 +15,14 @@ model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/
 # What and when?
 forecast_dates <- as.character(seq.Date(as_date('2023-01-01'), as_date('2023-12-31'), 'day'))
 variables <- c('temperature', 'oxygen', 'chla')
-
+sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
 
 # build a local parquet database from which to query
 scores_s3 |>
   arrow::open_dataset() |>
   dplyr::filter(reference_datetime %in% forecast_dates,
                 variable %in% variables,
+                site_id %in% sites, 
                 model_id %in% model_meta$model_id) |>
   arrow::write_dataset("scores_2023", partitioning=c("model_id", "site_id"))
 

### 2024-04-12 R/get_colours.R
update to PB

@@ -11,7 +11,7 @@ top_mods <- temperature_skill |>
 
 
 # define the colours
-top_process <- top_mods |> filter(model_type == 'process') |> pull(model_id)
+top_process <- top_mods |> filter(model_type == 'process-based') |> pull(model_id)
 process_cols <- RColorBrewer::brewer.pal(n = 9, name = 'YlGnBu')[3:6]
 names(process_cols) <- top_process
 
@@ -40,6 +40,6 @@ model_id_colours <- enframe(c(process_cols, MME_cols, empirical_cols, ML_cols)) 
 # colours for model types 
 cols_modeltype <- c('empirical' = "#C42503FF" ,
                     'MME' = "#FDAD35FF"  ,
-                    'process' = "#23C3E4FF" ,
+                    'process-based' = "#23C3E4FF" ,
                     'ML' = "#3E378FFF",
                     'null' = '#A9A9A9')

### 2024-04-15 02_chla_skills_scores.R
remove unused code

@@ -1,123 +1,0 @@
-
-library(tidyverse)
-library(arrow)
-library(lubridate)
-library(tidytext)
-
-lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
-cols_modeltype <- c('empirical' = "#482576FF",
-                    'multi-model ensemble' =  "#35608DFF",
-                    'process' = "#21908CFF",
-                    'ML' = '#BBDF27FF',
-                    'null' = '#43BF71FF')
-# read in the metadata from the google sheet
-googlesheets4::gs4_deauth()
-model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")
-
-# Open dataset-------------
-chla_scores <- arrow::open_dataset("scores") |>
-  filter(site_id %in% lake_sites,
-         # model_id != 'climatology',
-         model_id %in% model_meta$model_id,
-         variable == 'chla') |>
-  collect()
-
-# Calculate skill scores ------------
-chla_persistence <- arrow::open_dataset("scores") |>
-  filter(model_id == 'persistenceRW', 
-         site_id %in% lake_sites,
-         variable == 'chla') |>
-  collect() |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  arrange(reference_datetime, datetime, site_id) |> 
-  pivot_wider(values_from = c(crps, logs),
-              names_from = model_id)
-
-chla_skill <- chla_scores |>
-  filter(model_id != 'persistenceRW') |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  full_join(chla_persistence) |> 
-  
-  # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = crps_persistenceRW - crps,
-         skill_logs = logs_persistenceRW - logs,
-         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-  # consistent forecast period
-  filter(horizon <= 30,
-         horizon >= 1,
-         model_id != 'fARIMA') |> 
-  arrange(reference_datetime, site_id, datetime, model_id) 
-#--------------------------------------#
-
-chla_skill |> 
-  group_by(model_id) |> 
[…]

### 2024-04-15 02_oxygen_skill_scores.R

@@ -1,67 +1,0 @@
-
-library(tidyverse)
-library(arrow)
-library(lubridate)
-library(tidytext)
-
-lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
-cols_modeltype <- c('empirical' = "#482576FF",
-                    'multi-model ensemble' =  "#35608DFF",
-                    'process' = "#21908CFF",
-                    'ML' = '#BBDF27FF',
-                    'null' = '#43BF71FF')
-# read in the metadata from the google sheet
-googlesheets4::gs4_deauth()
-model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")
-
-# Open dataset-------------
-# calculate DO skill scores
-DO_scores <- arrow::open_dataset("scores") |>
-  filter(site_id %in% lake_sites,
-         # model_id != 'climatology',
-         model_id %in% model_meta$model_id,
-         variable == 'oxygen') |>
-  collect()
-
-# Calculate skill scores ------------
-DO_climatology <- arrow::open_dataset("scores") |>
-  filter(model_id == 'climatology', 
-         site_id %in% lake_sites,
-         variable == 'oxygen') |>
-  collect() |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  arrange(reference_datetime, datetime, site_id) |> 
-  pivot_wider(values_from = c(crps, logs),
-              names_from = model_id)
-
-DO_skill <- DO_scores |>
-  filter(model_id != 'climatology') |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  full_join(DO_climatology) |> 
-  
-  # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = crps_climatology - crps,
-         skill_logs = logs_climatology - logs,
-         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-  # consistent forecast period
-  filter(horizon <= 30,
-         horizon >= 1,
-         model_id != 'fARIMA') |> 
-  arrange(reference_datetime, site_id, datetime, model_id) 
-#------------------------------------#
-
-DO_skill |> 
[…]

### 2024-04-15 02_temperature_skill_scores.R

@@ -1,194 +1,0 @@
-
-library(tidyverse)
-library(arrow)
-library(lubridate)
-library(tidytext)
-
-lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
-cols_modeltype <- c('empirical' = "#482576FF",
-          'multi-model ensemble' =  "#35608DFF",
-          'process' = "#21908CFF",
-          'ML' = '#BBDF27FF',
-          'null' = '#43BF71FF')
-# read in the metadata from the google sheet
-googlesheets4::gs4_deauth()
-model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")
-
-# Open dataset-------------
-# start with lake temperatures
-temperature_scores <- arrow::open_dataset("scores") |>
-  filter(site_id %in% lake_sites,
-         # model_id != 'climatology',
-         model_id %in% model_meta$model_id,
-         variable == 'temperature') |>
-  collect()
-
-
-# Calculate skill scores ------------
-temperature_climatology <- arrow::open_dataset("scores") |>
-  filter(model_id == 'climatology', 
-         site_id %in% lake_sites,
-         variable == 'temperature') |>
-  collect() |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  arrange(reference_datetime, datetime, site_id) |> 
-  pivot_wider(values_from = c(crps, logs),
-              names_from = model_id)
-
-temperature_skill <- temperature_scores |>
-  filter(model_id != 'climatology') |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  full_join(temperature_climatology) |> 
-  
-  # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = crps_climatology - crps,
-         skill_logs = logs_climatology - logs,
-         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-  # consistent forecast period
-  filter(horizon <= 30,
-         horizon >= 1,
-         model_id != 'fARIMA') |> 
[…]

### 2024-04-15 03_plot_NOAA.R

@@ -1,23 +1,0 @@
-library(tidyverse)
-library(arrow)
-library(lubridate)
-
-lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
-
-# Open dataset-------------
-# start with lake temperatures
-NOAA_scores <- arrow::open_dataset("scores") |>
-  filter(site_id %in% lake_sites,
-         model_id %in% c('NOAA_daily', 'NOAA_daily_obs')) |>
-  collect()
-
-NOAA_scores |> 
-  mutate(horizon = as_date(datetime) - as_date(reference_datetime)) |> 
-  group_by(model_id, site_id, horizon) |> 
-  summarise(mean_crps = mean(crps, na.rm = T)) |> 
-  filter(horizon <= 30) |> 
-  ggplot(aes(x=horizon, y=mean_crps, colour = model_id)) +
-  geom_line() +
-  facet_wrap(~site_id) +
-  theme_bw() +
-  labs(y = 'CRPS score (oC)', title = 'NOAA air temperature forecast')

### 2024-04-15 03_score_NOAA.R

@@ -1,169 +1,0 @@
-library(tidyverse)
-library(score4cast)
-library(lubridate)
-library(neonUtilities)
-source('R/score_weather.R')
-source('R/ignore_sigpipe.R')
-
-# --- stage2 vs stage3 ----
-# want to score all this years forecasts
-# 1. Check the date
-dates <- seq.Date(as_date('2023-01-01'), (Sys.Date() - days(1)), 'day')
-
-# 2. Check if the forecast has already been scored
-scored_forecasts <-
-  arrow::open_dataset('scores') |>
-  filter(model_id == 'NOAA_daily') |>
-  distinct(reference_datetime, site_id) |>
-  collect() |>
-  mutate(reference_datetime = as_date(reference_datetime))
-
-# 3. Generate list of forecasts to be scored
-lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
-
-site_dates <- expand.grid(reference_datetime = dates,
-                          site_id = lake_sites)
-
-to_score <- site_dates |> 
-  anti_join(scored_forecasts) 
-
-# forecast from a couple of dates are missing
-
-to_score <- to_score |> 
-  filter(!reference_datetime %in% as_date(c('2023-01-06', '2023-01-19', '2023-05-22')))
-
-# Loop through all site-date combos to score
-for (i in 1:nrow(to_score)) {
-  # get the forecast
-  forecast_date <- to_score$reference_datetime[i]
-  site <- to_score$site_id[i]
-
-  stage_2 <- neon4cast::noaa_stage2(start_date = forecast_date) |>
-    filter(site_id == site,
-           variable == 'air_temperature') |>
-    collect()|> 
-    mutate(datetime = as_date(datetime)) |> 
-    group_by(parameter, reference_datetime, datetime, site_id, variable, family) |> 
-    summarise(prediction = mean(prediction))
-
-  # get the observations
-  stage_3 <- neon4cast::noaa_stage3() |>
-    filter(site_id  == site,
-           datetime %in% unique(stage_2$datetime),
-           variable == 'air_temperature') |>
-    collect() |>
-
-    # compare with the mean prediction (as if it were the observation)
-    mutate(datetime = as_date(datetime)) |> 
-    group_by(datetime, site_id, variable) |>
-    summarise(prediction = mean(prediction), .groups = 'drop') |>
-    rename(observation = prediction)
-
-
-  scores <- score_weather_forecast(forecast = stage_2, target = stage_3) |>
[…]

### 2024-04-15 02a_manuscript_plots.Rmd
rename files

--- 2024-04-11 04_manuscript_plots.Rmd
@@
   mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
-         ast = ifelse(mean_crps < -100, '>>>', ''),
+         ast = ifelse(mean_crps < -100, '>>', ''),
          uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'), 
@@
   ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +
-  annotate('rect', xmax = Inf, xmin = -Inf, ymin = 19.5, ymax =  Inf, fill = 'grey', alpha = 0.2) + 
+  annotate('rect', xmax = Inf, xmin = -Inf, ymin = 20.5, ymax =  Inf, fill = 'grey', alpha = 0.2) + 
   geom_bar_pattern(aes(pattern = uses_AT),# Different pattern for each group
@@
   coord_cartesian(xlim = c(-10, 5)) +
-  geom_text(aes(y=manuscript_name, x= 0.5, label = ast), inherit.aes = F)
+  geom_text(aes(y=manuscript_name, x= 0.75, label = ast), inherit.aes = F, )
 
@@
                   legend = 'top', labels = 'auto') |> 
-  ggsave(filename = 'Plots/test.png', height = 20, width = 25, units = 'cm')
+  ggsave(filename = 'Plots/Figure2.png', height = 20, width = 25, units = 'cm')
 

### 2024-04-17 01_get_scores.R
comment out s3 instructions

@@ -1,8 +1,6 @@
-# Get model_id for this years submissions
 
 library(tidyverse)
 library(lubridate)
-`%nin%` = Negate(`%in%`)
 
 # Get from data publication on Zenodo
 save_loc <- here::here()
@@ -11,30 +9,29 @@ download.file(url = "https://sandbox.zenodo.org/records/46145/files/scores_2023.
               destfile = file.path(save_loc,"scores.zip"), method = "curl")
 unzip(file.path(save_loc,"scores.zip"))
 
-scores_s3 <- 'scores_2023'
+
 
 # if using S3...
 # scores_s3 <- arrow::s3_bucket("neon4cast-scores/parquet/aquatics/",
 #                               endpoint_override= "data.ecoforecast.org", anonymous = T)
-
-
-googlesheets4::gs4_deauth()
-model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing") |> 
-  mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0))
-
-
-# What and when?
-forecast_dates <- as.character(seq.Date(as_date('2023-01-01'), as_date('2023-12-31'), 'day'))
-variables <- c('temperature', 'oxygen', 'chla')
-sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
-
-# build a local parquet database from which to query
-scores_s3 |>
-  arrow::open_dataset() |>
-  dplyr::filter(reference_datetime %in% forecast_dates,
-                variable %in% variables,
-                site_id %in% sites, 
-                model_id %in% model_meta$model_id) |>
-  arrow::write_dataset("scores_2023", partitioning=c("model_id", "site_id"))
+# scores_s3 <- 'scores_2023'
+# googlesheets4::gs4_deauth()
+# model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing") |> 
+#   mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0))
+# 
+# 
+# # What and when?
[…]

### 2024-04-17 install_packages.R
add package installation

@@ -1,0 +1,10 @@
+install.packages('tidyverse')
+install.packages('arrow')
+install.packages('here')
+install.packages('lubridate')
+
+install.packages('ggpubr')
+install.packages('cowplot')
+install.packages('tidytext')
+install.packages('ggh4x')
+install.packages('ggpattern')

### 2024-04-17 01_get_scores.R
optional curl arg

@@ -6,7 +6,8 @@ library(lubridate)
 save_loc <- here::here()
 
 download.file(url = "https://sandbox.zenodo.org/records/46145/files/scores_2023.zip?download=1",
-              destfile = file.path(save_loc,"scores.zip"), method = "curl")
+              destfile = file.path(save_loc,"scores.zip"))
+#If this doesn't work try adding the  method = "curl" argument
 unzip(file.path(save_loc,"scores.zip"))
 
 

### 2024-04-17 01_get_scores.R
extend timeout limit

@@ -2,6 +2,8 @@
 library(tidyverse)
 library(lubridate)
 
+options(timeout = 1000)
+
 # Get from data publication on Zenodo
 save_loc <- here::here()
 

### 2024-04-17 R/get_data.R
update for new scores db

@@ -1,36 +1,27 @@
 
 # Open dataset-------------
 # Dissolved oxygen
-oxygen_scores <- arrow::open_dataset("./scores_2023") |>
-  filter(site_id %in% lake_sites,
-         model_id %in% model_meta$model_id,
-         variable == 'oxygen') |>
+oxygen_scores <- arrow::open_dataset("./scores") |>
+  filter(variable == 'oxygen') |>
   collect()
 
 
 # chlorophyll
-chla_scores <- arrow::open_dataset("./scores_2023") |>
-  filter(site_id %in% lake_sites,
-         model_id %in% model_meta$model_id,
-         variable == 'chla') |>
+chla_scores <- arrow::open_dataset("./scores") |>
+  filter(variable == 'chla') |>
   collect()
 
 
 # temperature
-temperature_scores <- arrow::open_dataset("./scores_2023") |>
-  dplyr::filter(!(model_id %in% c('flareGOTM','flareGOTM_noDA', 'flare_ler', 'flare_ler_baseline') &
-                    site_id == 'TOOK')) |> 
-  filter(site_id %in% lake_sites,
-         model_id %in% model_meta$model_id,
-         variable == 'temperature') |>
+temperature_scores <- arrow::open_dataset("./scores") |>
+  filter(variable == 'temperature') |>
   collect()
 
 
 # read in DOY scores ------------
 # temp
-temperature_climatology <- arrow::open_dataset("./scores_2023") |>
+temperature_climatology <- arrow::open_dataset("./scores") |>
   filter(model_id == 'climatology', 
-         site_id %in% lake_sites,
          variable == 'temperature') |>
   collect() |> 
   select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
@@ -39,9 +30,8 @@ temperature_climatology <- arrow::open_dataset("./scores_2023") |>
               names_from = model_id)
 
 # DO
-oxygen_climatology <- arrow::open_dataset("./scores_2023") |>
+oxygen_climatology <- arrow::open_dataset("./scores") |>
   filter(model_id == 'climatology', 
-         site_id %in% lake_sites,
          variable == 'oxygen') |>
   collect() |> 
   select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
@@ -50,9 +40,8 @@ oxygen_climatology <- arrow::open_dataset("./scores_2023") |>
               names_from = model_id)
 
 # Chorophyll
[…]

### 2024-04-17 01_get_scores.R
remove "bad" forecasts

@@ -17,9 +17,8 @@ unzip(file.path(save_loc,"scores.zip"))
 # if using S3...
 # scores_s3 <- arrow::s3_bucket("neon4cast-scores/parquet/aquatics/",
 #                               endpoint_override= "data.ecoforecast.org", anonymous = T)
-# scores_s3 <- 'scores_2023'
 # googlesheets4::gs4_deauth()
-# model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing") |> 
+# model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing") |>
 #   mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0))
 # 
 # 
@@ -28,13 +27,17 @@ unzip(file.path(save_loc,"scores.zip"))
 # variables <- c('temperature', 'oxygen', 'chla')
 # sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
 # 
+# 
+# bad_forecasts <- expand.grid(model_id = c('flareGOTM','flareGOTM_noDA', 'flare_ler', 'flare_ler_baseline'),
+#                              site_id = "TOOK")
 # # build a local parquet database from which to query
 # scores_s3 |>
 #   arrow::open_dataset() |>
 #   dplyr::filter(reference_datetime %in% forecast_dates,
 #                 variable %in% variables,
-#                 site_id %in% sites, 
+#                 site_id %in% sites,
 #                 model_id %in% model_meta$model_id) |>
-#   arrow::write_dataset("scores_2023", partitioning=c("model_id", "site_id"))
+#   anti_join(bad_forecasts) |> 
+#   arrow::write_dataset("scores", partitioning=c("model_id"))
 
 #====================================================================#

### 2024-04-17 01_get_scores.R
change out dir for unzipping

@@ -10,7 +10,8 @@ save_loc <- here::here()
 download.file(url = "https://sandbox.zenodo.org/records/46145/files/scores_2023.zip?download=1",
               destfile = file.path(save_loc,"scores.zip"))
 #If this doesn't work try adding the  method = "curl" argument
-unzip(file.path(save_loc,"scores.zip"))
+
+unzip(file.path(save_loc,"scores.zip"), exdir = 'scores')
 
 
 

### 2024-04-17 01_get_scores.R
update sandbox url

@@ -7,7 +7,7 @@ options(timeout = 1000)
 # Get from data publication on Zenodo
 save_loc <- here::here()
 
-download.file(url = "https://sandbox.zenodo.org/records/46145/files/scores_2023.zip?download=1",
+download.file(url = "https://sandbox.zenodo.org/records/46413/files/scores_archive.zip?download=1",
               destfile = file.path(save_loc,"scores.zip"))
 #If this doesn't work try adding the  method = "curl" argument
 

### 2024-04-17 01_get_scores.R
remove outdir from zip

@@ -11,7 +11,7 @@ download.file(url = "https://sandbox.zenodo.org/records/46413/files/scores_archi
               destfile = file.path(save_loc,"scores.zip"))
 #If this doesn't work try adding the  method = "curl" argument
 
-unzip(file.path(save_loc,"scores.zip"), exdir = 'scores')
+unzip(file.path(save_loc,"scores.zip"))
 
 
 

### 2024-04-17 02a_manuscript_plots.Rmd
check paths

@@
   filter(model_id %in% c('persistenceRW', 'climatology'))
-source('get_data.R')
+source('R/get_data.R')
 
@@
 # total forecasts
-distinct_forecasts <- arrow::open_dataset("./scores_2023") |> 
+distinct_forecasts <- arrow::open_dataset("./scores") |> 
   distinct(model_id, site_id, variable, reference_datetime) |> 
@@
          fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class'))
-
-ggpubr::ggarrange(top10_temp, all_temp, nrow = 1, align = 'hv',
-                  common.legend = T,
-                  legend = 'top', labels = 'auto') |> 
-  ggsave(filename = './Plots/Figure2.png', width = 25, height = 11, units = 'cm')
 

### 2024-04-18 02b_supplementary_plots.Rmd
update title
check paths

--- 2024-04-11 04b_supplementary_plots.Rmd
@@
 ---
-title: "supplementary plots"
+title: "Supplementary Figures"
 output: html_document
@@
 # start with lake temperatures
-temperature_scores <- arrow::open_dataset("./scores_2023") |>
+temperature_scores <- arrow::open_dataset("./scores") |>
   dplyr::filter(!(model_id %in% c('flareGOTM','flareGOTM_noDA', 'flare_ler', 'flare_ler_baseline') &
@@
 # Calculate skill scores ------------
-temperature_climatology <- arrow::open_dataset("./scores_2023") |>
+temperature_climatology <- arrow::open_dataset("./scores") |>
   filter(model_id == 'climatology', 
@@
 ```{r}
-baseline_forecasts <- arrow::open_dataset("scores_2023") |>
+baseline_forecasts <- arrow::open_dataset("scores") |>
   filter(site_id %in% lake_sites,

### 2024-04-18 01_get_scores.R

@@ -9,7 +9,7 @@ save_loc <- here::here()
 
 download.file(url = "https://sandbox.zenodo.org/records/46413/files/scores_archive.zip?download=1",
               destfile = file.path(save_loc,"scores.zip"))
-#If this doesn't work try adding the  method = "curl" argument
+#If this doesn't work try adding the method = "curl" argument
 
 unzip(file.path(save_loc,"scores.zip"))
 

### 2024-04-18 02a_manuscript_plots.Rmd
update inventory

@@
 
-```
-
-## Summary stats
-```{r}
-# total forecasts
-distinct_forecasts <- arrow::open_dataset("./scores") |> 
-  distinct(model_id, site_id, variable, reference_datetime) |> 
-  collect()
-
-# variable forecasts
-distinct_forecasts |> 
-  group_by(variable ) |>
-  summarise(n = n()) |> 
-  mutate(percent = 100*(n/nrow(distinct_forecasts)))
-
-# lake temp forecasts
-distinct_forecasts |> 
-  filter(site_id %in% lake_sites,
-         variable == 'temperature') |> 
-  summarise(n = n()) 
 ```

### 2024-04-18 01a_inventory.R

@@ -1,10 +1,14 @@
 library(tidyverse)
-# inventory
+
+
+# Model submissions -------------------------------------------------------
+# Read in the google sheet of model descriptions
 googlesheets4::gs4_deauth()
 model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing") |> 
   mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0)) 
 
 lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
+
 # How many models?
 model_meta |>
   mutate(tw = str_detect(variable, 'temperature'),
@@ -18,7 +22,6 @@ model_meta |>
 
 
 # model classes by variable
-
 model_meta |> 
   filter(str_detect(variable, 'temperature')) |> 
   group_by(model_type) |> 
@@ -35,18 +38,6 @@ model_meta |>
   summarise(n())
 
 
-# How many forecasts?
-all_lake_forecasts <- arrow::open_dataset('scores_2023') |> 
-  filter(model_id %in% model_meta$model_id,
-         site_id %in% lake_sites) |> 
-  distinct(reference_datetime, variable, 
-           site_id, model_id) |> 
-  collect()
-
-all_lake_forecasts |> 
-  filter(variable == 'temperature') |> 
-  summarise(n())
-
 # use AT as a covariate?
 model_meta |> 
   select(model_id, NOAA_var, variable) |>
@@ -57,65 +48,42 @@ model_meta |>
   summarise(sum(tw), sum(do), sum(chla))
 
 
-
-
-# Which weather variables are used in each model?
-NOAA_vars <- c('air_temperature',
-               'air_pressure',
-               'relative_humidity',
-               'surface_downwelling_longwave_flux_in_air',
-               'surface_downwelling_shortwave_flux_in_air', 
-               'precipitation_flux',
-               'eastward_wind',
-               'northward_wind')
-
-NOAA_summary <- wt_meta |> 
-  select(model_id, variable) 
-
-
-for (i in 1:length(NOAA_vars)) {
-  detect_var <- NOAA_vars[i]
-  
-  NOAA_summary <- 
-    NOAA_summary |> 
-    mutate(detect_var = str_detect(wt_meta$NOAA_var, detect_var)) 
-  
[…]

### 2024-04-19 02b_supplementary_plots.Rmd
add top mods

@@
 ```{r}
+top_mods <- temperature_skill |> 
+  group_by(model_id) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  ungroup() |> 
+  filter(mean_crps > 0) |> 
+  distinct(model_id)
+
 horizon_skill <- 

### 2024-04-19 02b_supplementary_plots.Rmd
add directory for plots

@@
   arrange(reference_datetime, site_id, datetime, model_id)
+
+if (!dir.exists('Plots')) {
+  dir.create('Plots')
+}
 ```

### 2024-04-22 02a_manuscript_plots.Rmd
add directory for plots
change size of figure 2

@@
 
+if (!dir.exists('Plots')) {
+  dir.create('Plots')
+}
 ```
@@
   scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw() + 
+  theme_bw(base_size = 10) + 
   guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
                                 title = 'Uses air temperature as a covariate?'),
-         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class'))
+         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class')) +
+  theme(legend.box = 'vertical')
 
@@
   scale_pattern_type_manual(values=c(NA, NA), name = 'Uses air temperature as a covariate?') +
-  theme_bw() + 
+  theme_bw(base_size = 10) + 
   guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses air temperature as a covariate?'),
-         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class'))
+         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class'))  +
+  theme(legend.box = 'vertical')
 
@@
   scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw() +
+  theme_bw(base_size = 10) + 
   theme(legend.position = 'none')
@@
   scale_pattern_type_manual(values=c(NA, NA), name = 'Uses air temperature as a covariate?') +
-  theme_bw()+
+  theme_bw(base_size = 10) + 
   theme(legend.position = 'none')  +
@@
                   legend = 'top', labels = 'auto') |> 
-  ggsave(filename = 'Plots/Figure2.png', height = 20, width = 25, units = 'cm')
+  ggsave(filename = 'Plots/Figure2.png', height = 20, width = 15.5, units = 'cm')
 

### 2024-04-25 02a_manuscript_plots.Rmd
remove chla

@@
 googlesheets4::gs4_deauth()
-model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")
+model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/16J_jBbYyMoPbJujBHnKbANXKIMCcu2UtjDHwnL5mUA0/edit?usp=sharing")
 
@@
   guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02), 
-                                title = 'Uses air temperature as a covariate?'),
+                                title = 'Uses air temperature\nas a covariate?'),
          fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class')) +
@@
 
-all_chla <- chla_skill |> 
-  group_by(model_id) |> 
-  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
-  left_join(model_meta) |> 
-  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
-         ast = ifelse(mean_crps < -100, '>>', ''),
-         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'), 
-         uses_AT = as.factor(str_detect(NOAA_var, 'air_temperature')),
-         mean_crps = ifelse(mean_crps < -100, -100, mean_crps)) |> 
-  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +
-  annotate('rect', xmax = Inf, xmin = -Inf, ymin = 20.5, ymax =  Inf, fill = 'grey', alpha = 0.2) + 
-  geom_bar_pattern(aes(pattern = uses_AT),# Different pattern for each group
-                   stat = "identity",
-                   alpha = 0.8, 
-                   colour = 'black',
-                   pattern_fill = "black",
-                   pattern_colour = "black", width = 0.8,
-                   pattern_spacing = 0.03,
-                   pattern_alpha = 1) + 
-  scale_y_reordered() +
-  labs(y='Model', x = 'Mean relative skill (ug/L)', subtitle = 'Chlorophyll a') +
-  geom_vline(xintercept = 0) +
-  scale_fill_manual(values = cols_modeltype) +
-  scale_pattern_manual(values=c('none', 'stripe'), name = 'Uses air temperature as a covariate?') +
-  scale_pattern_type_manual(values=c(NA, NA), name = 'Uses air temperature as a covariate?') +
-  theme_bw(base_size = 10) + 
-  theme(legend.position = 'none')  +
-  coord_cartesian(xlim = c(-10, 5)) +
-  geom_text(aes(y=manuscript_name, x= 0.75, label = ast), inherit.aes = F, )
-
-ggpubr::ggarrange(top10_temp, all_temp, all_DO, all_chla, nrow = 2, ncol= 2, align = 'hv',
-                  common.legend = T,
-                  legend = 'top', labels = 'auto') |> 
-  ggsave(filename = 'Plots/Figure2.png', height = 20, width = 15.5, units = 'cm')
+legend <- get_legend(top10_temp)
+ggpubr::ggarrange(top10_temp, ggarrange(as_ggplot(legend), NULL, ncol = 2, widths = c(1, 3)),
+                  all_temp, all_DO, 
+                  nrow = 2, ncol= 2, 
+                  align = 'hv', heights = c(1, 1.3),
+                  legend = 'none', labels = c('a', '', 'b', 'c')) |> 
+  ggsave(filename = 'Plots/Figure2.png', height = 20, width = 17, units = 'cm')
 

### 2024-04-25 00_get_scores.R

@@ -7,7 +7,7 @@ options(timeout = 1000)
 # Get from data publication on Zenodo
 save_loc <- here::here()
 
-download.file(url = "https://sandbox.zenodo.org/records/46413/files/scores_archive.zip?download=1",
+download.file(url = "https://sandbox.zenodo.org/records/48909/files/scores_v2.zip?download=1",
               destfile = file.path(save_loc,"scores.zip"))
 #If this doesn't work try adding the method = "curl" argument
 

### 2024-04-25 01a_inventory.R

@@ -4,7 +4,7 @@ library(tidyverse)
 # Model submissions -------------------------------------------------------
 # Read in the google sheet of model descriptions
 googlesheets4::gs4_deauth()
-model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing") |> 
+model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/16J_jBbYyMoPbJujBHnKbANXKIMCcu2UtjDHwnL5mUA0/edit?usp=sharing") |> 
   mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0)) 
 
 lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
@@ -12,9 +12,8 @@ lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
 # How many models?
 model_meta |>
   mutate(tw = str_detect(variable, 'temperature'),
-         do = str_detect(variable, 'oxygen'),
-         chla = str_detect(variable, 'chla')) |> 
-  summarise(sum(tw), sum(do), sum(chla))
+         do = str_detect(variable, 'oxygen')) |> 
+  summarise(sum(tw), sum(do))
 
 model_meta |> 
   group_by(variable) |> 
@@ -32,37 +31,28 @@ model_meta |>
   group_by(model_type) |> 
   summarise(n())
 
-model_meta |> 
-  filter(str_detect(variable, 'chla')) |> 
-  group_by(model_type) |> 
-  summarise(n())
-
 
 # use AT as a covariate?
 model_meta |> 
   select(model_id, NOAA_var, variable) |>
   mutate(tw = str_detect(variable, 'temperature'),
-         do = str_detect(variable, 'oxygen'),
-         chla = str_detect(variable, 'chla')) |> 
+         do = str_detect(variable, 'oxygen')) |> 
   filter(str_detect(NOAA_var, 'air_temp')) |>
-  summarise(sum(tw), sum(do), sum(chla))
+  summarise(sum(tw), sum(do))
 
 
 # types of uncertainty represented
 # Number of models that contain each type
 model_meta |>
-  mutate(tw = str_detect(variable, 'temperature'),
-         do = str_detect(variable, 'oxygen'),
[…]

### 2024-04-25 R/get_data.R

@@ -1,26 +1,19 @@
 
 # Open dataset-------------
 # Dissolved oxygen
-oxygen_scores <- arrow::open_dataset("./scores") |>
+oxygen_scores <- arrow::open_dataset("./scores_v2") |>
   filter(variable == 'oxygen') |>
   collect()
 
-
-# chlorophyll
-chla_scores <- arrow::open_dataset("./scores") |>
-  filter(variable == 'chla') |>
-  collect()
-
-
 # temperature
-temperature_scores <- arrow::open_dataset("./scores") |>
+temperature_scores <- arrow::open_dataset("./scores_v2") |>
   filter(variable == 'temperature') |>
   collect()
 
 
 # read in DOY scores ------------
 # temp
-temperature_climatology <- arrow::open_dataset("./scores") |>
+temperature_climatology <- arrow::open_dataset("./scores_v2") |>
   filter(model_id == 'climatology', 
          variable == 'temperature') |>
   collect() |> 
@@ -30,7 +23,7 @@ temperature_climatology <- arrow::open_dataset("./scores") |>
               names_from = model_id)
 
 # DO
-oxygen_climatology <- arrow::open_dataset("./scores") |>
+oxygen_climatology <- arrow::open_dataset("./scores_v2") |>
   filter(model_id == 'climatology', 
          variable == 'oxygen') |>
   collect() |> 
@@ -39,15 +32,6 @@ oxygen_climatology <- arrow::open_dataset("./scores") |>
   pivot_wider(values_from = c(crps, logs),
               names_from = model_id)
 
-# Chorophyll
-chla_climatology <- arrow::open_dataset("./scores") |>
-  filter(model_id == 'climatology', 
-         variable == 'chla') |>
-  collect() |> 
-  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
-  arrange(reference_datetime, datetime, site_id) |> 
-  pivot_wider(values_from = c(crps, logs),
-              names_from = model_id)
 #-------------------------#
 
 # Calculate relative skill ------
@@ -71,22 +55,6 @@ oxygen_skill <- oxygen_scores |>
   filter(model_id != 'climatology') |> 
   select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
   full_join(oxygen_climatology) |> 
-  
-  # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = crps_climatology - crps,
[…]

### 2024-04-29 02a_manuscript_plots.Rmd
restructure scripts

@@
 library(ggh4x)
+```
+## Get scores
+
+```{r}
+source('00_get_scores.R')
 ```
@@
 }
+```
+
+## Inventory
+
+```{r}
+
+# How many models?
+model_meta |>
+  mutate(tw = str_detect(variable, 'temperature'),
+         do = str_detect(variable, 'oxygen')) |> 
+  summarise(sum(tw), sum(do))
+
+model_meta |> 
+  group_by(variable) |> 
+  summarise(n())
+
+# model classes by variable
+model_meta |> 
+  filter(str_detect(variable, 'temperature')) |> 
+  group_by(model_type) |> 
+  summarise(n())
+
+model_meta |> 
+  filter(str_detect(variable, 'oxygen')) |> 
+  group_by(model_type) |> 
+  summarise(n())
+
+# use AT as a covariate?
+model_meta |> 
+  select(model_id, NOAA_var, variable) |>
+  mutate(tw = str_detect(variable, 'temperature'),
+         do = str_detect(variable, 'oxygen')) |> 
+  filter(str_detect(NOAA_var, 'air_temp')) |>
+  summarise(sum(tw), sum(do))
+
+# types of uncertainty represented
+# Number of models that contain each type
+model_meta |>
+  mutate(tw = str_detect(variable, 'temperature'))  |> 
+  filter(tw == T) |> 
+  summarise(across(contains('unc'), ~ sum(.x, na.rm = T)))
+
+# How many sources are represented in each model?
+model_meta |>
+  mutate(tw = str_detect(variable, 'temperature'))  |> 
+  filter(tw == T) |> 
+  select(model_id, contains('unc_')) |> 
+  mutate(unc_n = rowSums(pick(where(is.numeric)))) |> 
+  group_by(unc_n) |> 
+  summarise(n = n())
+
+# Scores database summary -------------------------------------------------
+# total forecasts - distinct model/site/var/refdate combos
+distinct_forecasts <- arrow::open_dataset("./scores_v2") |> 
+  distinct(model_id, site_id, variable, reference_datetime) |> 
+  collect()
+
+# variable forecasts
+distinct_forecasts |> 
+  group_by(variable ) |>
+  summarise(n = n()) |> 
+  mutate(percent = 100*(n/nrow(distinct_forecasts)))
+
 ```

### 2024-04-29 00_get_scores.R

@@ -7,7 +7,7 @@ options(timeout = 1000)
 # Get from data publication on Zenodo
 save_loc <- here::here()
 
-download.file(url = "https://sandbox.zenodo.org/records/48909/files/scores_v2.zip?download=1",
+download.file(url = "https://sandbox.zenodo.org/records/46413/files/scores_archive.zip?download=1",
               destfile = file.path(save_loc,"scores.zip"))
 #If this doesn't work try adding the method = "curl" argument
 
@@ -25,12 +25,13 @@ unzip(file.path(save_loc,"scores.zip"))
 # 
 # # What and when?
 # forecast_dates <- as.character(seq.Date(as_date('2023-01-01'), as_date('2023-12-31'), 'day'))
-# variables <- c('temperature', 'oxygen', 'chla')
+# variables <- c('temperature', 'oxygen')
 # sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
 # 
 # 
 # bad_forecasts <- expand.grid(model_id = c('flareGOTM','flareGOTM_noDA', 'flare_ler', 'flare_ler_baseline'),
-#                              site_id = "TOOK")
+#                              site_id = "TOOK",
+#                              variable = 'temperature')
 # # build a local parquet database from which to query
 # scores_s3 |>
 #   arrow::open_dataset() |>
@@ -38,7 +39,7 @@ unzip(file.path(save_loc,"scores.zip"))
 #                 variable %in% variables,
 #                 site_id %in% sites,
 #                 model_id %in% model_meta$model_id) |>
-#   anti_join(bad_forecasts) |> 
+#   anti_join(bad_forecasts) |>
 #   arrow::write_dataset("scores", partitioning=c("model_id"))
 
 #====================================================================#

### 2024-04-29 manuscript_plots.Rmd
restructure scripts

--- 2024-04-29 02a_manuscript_plots.Rmd
@@
 # total forecasts - distinct model/site/var/refdate combos
-distinct_forecasts <- arrow::open_dataset("./scores_v2") |> 
+distinct_forecasts <- arrow::open_dataset("./scores") |> 
   distinct(model_id, site_id, variable, reference_datetime) |> 

### 2024-04-29 supplementary_plots.Rmd
restructure scripts

--- 2024-04-19 02b_supplementary_plots.Rmd
@@
 # start with lake temperatures
-temperature_scores <- arrow::open_dataset("./scores") |>
-  dplyr::filter(!(model_id %in% c('flareGOTM','flareGOTM_noDA', 'flare_ler', 'flare_ler_baseline') &
-                  site_id == 'TOOK')) |> 
+temperature_scores <- arrow::open_dataset("./scores_v2") |> 
   filter(site_id %in% lake_sites,
@@
 # Calculate skill scores ------------
-temperature_climatology <- arrow::open_dataset("./scores") |>
+temperature_climatology <- arrow::open_dataset("./scores_v2") |>
   filter(model_id == 'climatology', 
@@
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
+  scale_colour_viridis_d(option = 'A', begin=0.2, end=0.9)
 

### 2024-04-29 01a_inventory.R

@@ -1,73 +1,0 @@
-library(tidyverse)
-
-
-# Model submissions -------------------------------------------------------
-# Read in the google sheet of model descriptions
-googlesheets4::gs4_deauth()
-model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/16J_jBbYyMoPbJujBHnKbANXKIMCcu2UtjDHwnL5mUA0/edit?usp=sharing") |> 
-  mutate(unc_driver = ifelse(uses_NOAA == 1, 1, 0)) 
-
-lake_sites <- c('BARC', 'CRAM', 'LIRO', 'PRLA', 'PRPO', 'SUGG', 'TOOK')
-
-# How many models?
-model_meta |>
-  mutate(tw = str_detect(variable, 'temperature'),
-         do = str_detect(variable, 'oxygen')) |> 
-  summarise(sum(tw), sum(do))
-
-model_meta |> 
-  group_by(variable) |> 
-  summarise(n())
-
-
-# model classes by variable
-model_meta |> 
-  filter(str_detect(variable, 'temperature')) |> 
-  group_by(model_type) |> 
-  summarise(n())
-
-model_meta |> 
-  filter(str_detect(variable, 'oxygen')) |> 
-  group_by(model_type) |> 
-  summarise(n())
-
-
-# use AT as a covariate?
-model_meta |> 
-  select(model_id, NOAA_var, variable) |>
-  mutate(tw = str_detect(variable, 'temperature'),
-         do = str_detect(variable, 'oxygen')) |> 
-  filter(str_detect(NOAA_var, 'air_temp')) |>
-  summarise(sum(tw), sum(do))
-
-
-# types of uncertainty represented
-# Number of models that contain each type
-model_meta |>
-  mutate(tw = str_detect(variable, 'temperature'))  |> 
-  filter(tw == T) |> 
-  summarise(across(contains('unc'), ~ sum(.x, na.rm = T)))
-
-
-
-# How many sources are represented in each model?
-model_meta |>
-  mutate(tw = str_detect(variable, 'temperature'))  |> 
-  filter(tw == T) |> 
-  select(model_id, contains('unc_')) |> 
-  mutate(unc_n = rowSums(pick(where(is.numeric)))) |> 
-  group_by(unc_n) |> 
-  summarise(n = n())
-
-
-# Scores database summary -------------------------------------------------
-# total forecasts - distinct model/site/var/refdate combos
[…]

### 2024-04-29 R/get_data.R
update paths

@@ -1,19 +1,19 @@
 
 # Open dataset-------------
 # Dissolved oxygen
-oxygen_scores <- arrow::open_dataset("./scores_v2") |>
+oxygen_scores <- arrow::open_dataset("./scores") |>
   filter(variable == 'oxygen') |>
   collect()
 
 # temperature
-temperature_scores <- arrow::open_dataset("./scores_v2") |>
+temperature_scores <- arrow::open_dataset("./scores") |>
   filter(variable == 'temperature') |>
   collect()
 
 
 # read in DOY scores ------------
 # temp
-temperature_climatology <- arrow::open_dataset("./scores_v2") |>
+temperature_climatology <- arrow::open_dataset("./scores") |>
   filter(model_id == 'climatology', 
          variable == 'temperature') |>
   collect() |> 
@@ -23,7 +23,7 @@ temperature_climatology <- arrow::open_dataset("./scores_v2") |>
               names_from = model_id)
 
 # DO
-oxygen_climatology <- arrow::open_dataset("./scores_v2") |>
+oxygen_climatology <- arrow::open_dataset("./scores") |>
   filter(model_id == 'climatology', 
          variable == 'oxygen') |>
   collect() |> 

### 2024-04-29 manuscript_plots.Rmd
add package install

@@
 library(ggh4x)
+source('install_packages.R')
 ```

### 2024-04-29 supplementary_plots.Rmd
update paths

--- 2024-04-19 02b_supplementary_plots.Rmd
@@
 temperature_scores <- arrow::open_dataset("./scores") |> 
-  dplyr::filter(!(model_id %in% c('flareGOTM','flareGOTM_noDA', 'flare_ler', 'flare_ler_baseline') &
-                  site_id == 'TOOK')) |> 
   filter(site_id %in% lake_sites,

### 2024-04-30 manuscript_plots.Rmd
update paths
use archive

--- 2024-04-29 manuscript_plots.Rmd
@@
 library(ggh4x)
+# source('install_packages.R')
 ```
+
 ## Get scores
@@
 ```{r}
-source('00_get_scores.R')
+if (!dir.exists('scores')) {
+   source('R/get_scores.R')
+}
 ```
@@
 # read in the metadata from the google sheet
-googlesheets4::gs4_deauth()
-model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/16J_jBbYyMoPbJujBHnKbANXKIMCcu2UtjDHwnL5mUA0/edit?usp=sharing")
+model_meta <- read_csv("https://zenodo.org/records/11087208/files/Model_descriptions.csv?download=1", na = 'character')
 
@@
   filter(model_id %in% c('persistenceRW', 'climatology'))
-source('R/get_data.R')
+source('R/load_data.R')
 
@@
 ```{r target-plot}
-targets <- readr::read_csv("https://data.ecoforecast.org/neon4cast-targets/aquatics/aquatics-targets.csv.gz") |> 
+targets <- readr::read_csv("targets/targets_2016-2023.csv") |> 
   na.omit() |> 

### 2024-04-30 supplementary_plots.Rmd
use archive

@@
 ```{r read-in-data}
-targets <- readr::read_csv("https://data.ecoforecast.org/neon4cast-targets/aquatics/aquatics-targets.csv.gz") |> 
+targets <- readr::read_csv("targets/targets_2016-2023.csv") |> 
   na.omit()
@@
                     'null' = '#A9A9A9')
-# read in the metadata from the google sheet
-googlesheets4::gs4_deauth()
-model_meta <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1oC7_w63wSCXNiHs1IK8AFGr0MG-NdjDAjwkfjvRZW-I/edit?usp=sharing")
+# read in the metadata 
+model_meta <- read_csv("https://zenodo.org/records/11087208/files/Model_descriptions.csv?download=1", na = 'character')
 

### 2024-04-30 R/get_scores.R
rename scripts

@@ -5,14 +5,16 @@ library(lubridate)
 options(timeout = 1000)
 
 # Get from data publication on Zenodo
-save_loc <- here::here()
 
-download.file(url = "https://sandbox.zenodo.org/records/46413/files/scores_archive.zip?download=1",
-              destfile = file.path(save_loc,"scores.zip"))
+download.file(url = "https://zenodo.org/records/11087208/files/scores.zip?download=1",
+              destfile = "scores.zip")
+download.file(url = "https://zenodo.org/records/11087208/files/targets.zip?download=1",
+              destfile = "targets.zip")
+
+
 #If this doesn't work try adding the method = "curl" argument
-
-unzip(file.path(save_loc,"scores.zip"))
-
+unzip("scores.zip")
+unzip("targets.zip")
 
 
 # if using S3...

### 2024-09-02 install_packages.R
new packages

@@ -8,3 +8,4 @@ install.packages('cowplot')
 install.packages('tidytext')
 install.packages('ggh4x')
 install.packages('ggpattern')
+install.packages('pals')

### 2024-09-02 manuscript_plots.Rmd
edits for review

@@
   geom_hline(yintercept = 0, colour = 'grey40' ) + 
-  geom_line()  +  
+  geom_line(linewidth = 0.8)  +  
   labs(y='Mean relative skill (°C)', x = 'Horizon (days)') +
@@
              y = mean_sd, colour = manuscript_name)) +
-  geom_line()+
+  geom_line(linewidth = 0.8)+
   labs(y='Mean standard deviation (°C)', x = 'Horizon (days)') +
@@
              y = mean_abs_bias, colour = manuscript_name)) +
-  geom_line()+
+  geom_line(linewidth = 0.8)+
   labs(y='Mean absolute bias (°C)', x = 'Horizon (days)') +
@@
                    order = levels(model_id_colours$manuscript_name)) +
-  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
+  scale_colour_manual(name = 'Top model',values = model_id_colours$value) +
+  scale_linetype_manual(name = 'Top model',values = model_id_colours$linetype) +
   theme_bw() +
-  scale_y_continuous(limits = c(54, 100))
+  scale_y_continuous(limits = c(54, 100)) +
+  labs(title = 'Top performing models') +
+  theme(title = element_text(size = 14))
 
@@
                             order = levels(model_id_colours$manuscript_name)) +
-  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
+  scale_colour_manual(name = 'Top model',values = model_id_colours$value) +
+  scale_linetype_manual(name = 'Top model',values = model_id_colours$linetype) +
   theme_bw() +
@@
                    CI = 80, 
-                 order = bad_mods$manuscript_name) +
+                   order = bad_mods$manuscript_name,
+                   alpha  = 0.6) +
   theme_bw()  +
-    scale_colour_manual(name = '', values = rep('grey50', 20)) +
-    scale_y_continuous(limits = c(0, 100)) + 
-    theme(legend.position = 'none')
+  scale_colour_manual(name = '', values = unname(pals::cols25(n= 20)))+ 
+  scale_linetype_manual(name = '', values = rep('solid', 20)) +
+  theme(legend.position = 'none') +
+  labs(title = 'All other submitted models') +
+  theme(title = element_text(size = 14))
 
@@
                            CI = 95, 
-                            order = bad_mods$manuscript_name) +
+                           order = bad_mods$manuscript_name,
+                           alpha  =0.6) +
   theme_bw() +
   scale_y_continuous(limits = c(0, 100))   +
-  scale_colour_manual(name = '', values = rep('grey50', 20))
+  scale_colour_manual(name = '', values = unname(pals::cols25(n= 20))) + 
+  scale_linetype_manual(name = '', values = rep('solid', 20))
 
-ggarrange(ggarrange(get_legend(good_80), NULL, nrow = 2),
-          ggarrange(good_80, good_95, bad_80, bad_95, align = 'hv', legend= 'none', labels = 'auto'), 
-          ncol = 2, 
-          widths = c(1,5)) |> 
+ggarrange(NULL, ggarrange(get_legend(good_80), NULL, nrow = 2),
+          ggarrange(good_80, good_95, bad_80, bad_95, align = 'hv', legend= 'none', labels = 'auto', hjust = -4, vjust = 3), 
+          ncol = 3, 
+          widths = c(0.1,1,5)) |> 
   ggsave(filename = 'Plots/Figure4.png', height = 17.6, width = 23, units = 'cm')
@@
 
+lake_order <- temperature_skill |> 
+  filter(model_id %in% top_mods$model_id) |> 
+  group_by(site_id, horizon) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) %>% 
+  filter(horizon == 30) %>% 
+  arrange(desc(mean_crps)) %>% 
+  pull(site_id)
+
 B <- temperature_skill |> 
@@
   summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
-  mutate(site_id = factor(site_id, levels = lake_sites)) |> 
+  mutate(site_id = factor(site_id, levels = lake_order)) |> 
   ggplot(aes(x=horizon, 
@@
   geom_hline(yintercept = 0, colour = 'grey40') + 
-  geom_line()  +  
+  geom_line(linewidth = 0.8)  +  
   labs(y='Mean relative skill (°C)', x = 'Horizon (days)') +

### 2024-09-02 supplementary_plots.Rmd
new SI figures comparisons with persistence

@@
 
+oxygen_scores <- arrow::open_dataset("./scores") |> 
+  filter(site_id %in% lake_sites,
+         model_id %in% model_meta$model_id,
+         variable == 'oxygen') |>
+  collect()
+
 # Calculate skill scores ------------
@@
   arrange(reference_datetime, site_id, datetime, model_id)
+
+# read in persistence scores ------------
+# temp
+temperature_persistence <- arrow::open_dataset("./scores") |>
+  filter(model_id == 'persistenceRW', 
+         variable == 'temperature') |>
+  collect() |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  arrange(reference_datetime, datetime, site_id) |> 
+  pivot_wider(values_from = c(crps, logs),
+              names_from = model_id)
+
+# DO
+oxygen_persistence <- arrow::open_dataset("./scores") |>
+  filter(model_id == 'persistenceRW', 
+         variable == 'oxygen') |>
+  collect() |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  arrange(reference_datetime, datetime, site_id) |> 
+  pivot_wider(values_from = c(crps, logs),
+              names_from = model_id)
+
+#-------------------------#
+
+# Calculate relative skill with persistence ------
+temperature_skill_RW <- temperature_scores |>
+  filter(model_id != 'persistenceRW') |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  full_join(temperature_persistence) |>
+
+  # calcaulte the skill relative to climatology (- is bad, + is good)
+  mutate(skill_crps = crps_persistenceRW - crps,
+         skill_logs = logs_persistenceRW - logs,
+         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+  # consistent forecast period
+  filter(horizon <= 30,
+         horizon >= 1, 
+         horizon != 15) |> 
+  arrange(reference_datetime, site_id, datetime, model_id)
+
+oxygen_skill_RW <- oxygen_scores |>
+  filter(model_id != 'persistenceRW') |> 
+  select(reference_datetime, datetime, crps, logs, site_id, model_id) |> 
+  full_join(oxygen_persistence) |>
+
+  # calcaulte the skill relative to persistenceRW (- is bad, + is good)
+  mutate(skill_crps = crps_persistenceRW - crps,
+         skill_logs = logs_persistenceRW - logs,
+         horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
+  # consistent forecast period
+  filter(horizon <= 30,
+         horizon >= 1, 
+         horizon != 15) |> 
+  arrange(reference_datetime, site_id, datetime, model_id)
+
+baselines <- model_meta |> 
+  filter(model_id %in% c('persistenceRW', 'climatology'))
+
+source('R/get_colours.R')
 
@@
 ```
+```{r}
+skill_summary_RW <- temperature_skill_RW |>
+  group_by(model_id) |>
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |>
+  ungroup() |>
+  filter(mean_crps > 0) |>
+  left_join(model_meta) |>
+  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'),
+         uses_AT = str_detect(NOAA_var, 'air_temperature')) |>
+  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +    # Different pattern for each group
+  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
+                   stat = "identity",
+                   alpha = 0.8,
+                   pattern_fill = "black",
+                   pattern_colour = "black",
+                   pattern_spacing = 0.04,
+                   pattern_alpha = 1,
+                   colour = 'black') +
+  scale_y_reordered() +
+  labs(y='Model', x = 'Mean relative skill (°C)') +
+  geom_vline(xintercept = 0) +
+  scale_fill_manual(values = cols_modeltype) +
+  scale_pattern_manual(values=c('none', 'stripe')) +
+  scale_pattern_type_manual(values=c(NA, NA)) +
+  theme_bw(base_size = 10) +
+  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02),
+                                title = 'Uses air temperature\nas a covariate?'),
+         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class')) +
+    theme(legend.box = 'horizontal')
+
+skill_summary_RW_nolegend <- skill_summary_RW +
+  theme(legend.position = 'none')
+
+skill_horizon_RW <- temperature_skill_RW |> 
+  filter(model_id %in% c(top_mods$model_id) & model_id != 'climatology', 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(model_id, horizon) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  left_join(select(model_meta, model_id, manuscript_name)) |> 
+  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name))) |> 
+  ggplot(aes(x=horizon, 
+             y = mean_crps, colour = manuscript_name)) +
+  geom_hline(yintercept = 0, colour = 'grey40' ) + 
+  geom_line(linewidth = 0.8)  +  
+  labs(y='Mean relative skill (°C)', x = 'Horizon (days)') +
+  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
+  theme_bw()
+
+ggpubr::ggarrange(ggarrange(NULL, as_ggplot(get_legend(skill_summary_RW)), NULL, NULL, widths =  c(0.2, 0.9,0.2, 1.1), nrow = 1),
+                  ggarrange(skill_summary_RW_nolegend,  NULL, skill_horizon_RW, align = 'h', widths =  c(0.9,0.05, 1.1), nrow = 1), 
+                  heights = c(0.5, 1),nrow = 2) |> 
+  ggsave(filename = 'Plots/FigureSnew.png', height = 17, width = 18, units = 'cm')
+```
+```{r}
+null_scores <- temperature_scores |> mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) %>% 
+  filter(model_id %in% baselines$model_id, 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(model_id, horizon) |> 
+  summarise(mean_crps = mean(crps, na.rm = T)) %>% 
+  mutate(model_id = factor(model_id, levels = c('climatology', 'persistenceRW'), labels = c("day-of-year", "persistence"))) %>% 
+  ggplot(aes(x=horizon,y = mean_crps, colour = model_id)) + 
+  geom_line() + 
+  scale_x_continuous(breaks = seq(2,30, 2)) + 
+  theme_bw() + 
+  labs(x= 'Horizon (days)', y = 'Mean CRPS (degrees C)') + 
+  scale_colour_viridis_d(name = 'Null Model', begin = 0.2, end=0.9) +
+  theme(panel.grid.minor = element_blank())
+
+ggsave(null_scores, filename = 'Plots/Figure_Snew2.png', height = 10, width = 15, units = 'cm')
+```
+
+```{r}
+
+temp_skill_RW_summary <- temperature_skill_RW |> 
+  filter(model_id %in% c(top_mods$model_id) & model_id != 'climatology', 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(model_id, horizon) |> 
+  summarise(mean_crpsRW = mean(skill_crps, na.rm = T)) |> 
+  left_join(select(model_meta, model_id, manuscript_name)) |> 
+  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name)))
+
+temp_skill_clim_summary <- temperature_skill |> 
+  filter(model_id %in% c(top_mods$model_id) & model_id != 'persistence', 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(model_id, horizon) |> 
+  summarise(mean_crpsDOY = mean(skill_crps, na.rm = T)) |> 
+  left_join(select(model_meta, model_id, manuscript_name)) |> 
+  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name)))
+
+null_model_beats <- full_join(temp_skill_clim_summary, temp_skill_RW_summary, 
+          by = join_by(model_id, horizon, manuscript_name)) %>% 
+  mutate(beats = ifelse(mean_crpsDOY > 0 & mean_crpsRW > 0, 'both',
+                        ifelse(mean_crpsDOY < 0 & mean_crpsRW > 0, 'RW',
+                               ifelse(mean_crpsDOY > 0 & mean_crpsRW < 0, 'DOY', 'none')))) %>% 
+  mutate(beats = factor(beats, levels = c("none", "RW", "DOY", "both"), labels = c("Neither", "Persistence", "Day-of-year", "Both"))) %>% 
+  na.omit() %>% 
+  ggplot(aes(x=horizon, y=manuscript_name, shape = beats, colour = beats)) + 
+  geom_point(size = 5) +
+  scale_shape_manual(values = c(0, 1, 4, 13), name = 'Null model that is out-performed') +
+  scale_colour_manual(values = c("grey90", "#455BCDFF", "#E1DD37FF", "#46F884FF"), name = 'Null model that is out-performed') +
+  scale_y_discrete(name = 'Model') +
+  scale_x_continuous(name = 'Horizon (days)', breaks = c(1,10, 20, 30), expand = c(0.02,0)) +
+  geom_hline(yintercept = seq(.5, 9.5, 1), linewidth = .2) +
+  theme(legend.position = "bottom", 
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8),
+        axis.ticks.y = element_blank())
+
+ggsave(null_model_beats, filename = 'Plots/Figure_Snew3.png', height = 10, width = 18, units = 'cm')
+
+```

### 2024-09-02 R/get_colours.R
minor edits for revisions

@@ -32,10 +32,11 @@ names(ML_cols) <- top_ML
 # make into a dataframe
 model_id_colours <- enframe(c(process_cols, MME_cols, empirical_cols, ML_cols)) |> 
   full_join(top_mods, by = join_by(name == model_id)) |> 
-  mutate(value = ifelse(is.na(value), 'grey20', value),
+  mutate(value = ifelse(is.na(value), '#A9A9A9', value),
          mean_crps = ifelse(is.na(mean_crps), 0, mean_crps),
          manuscript_name = fct_reorder(manuscript_name, mean_crps, .desc = T)) |> 
-  arrange(desc(mean_crps))
+  arrange(desc(mean_crps)) %>% 
+  mutate(linetype = ifelse(name == 'climatology', 'dotdash', 'solid'))
 
 # colours for model types 
 cols_modeltype <- c('empirical' = "#C42503FF" ,

### 2024-09-02 R/reliability_plot.R

@@ -1,4 +1,4 @@
-plot_reliability <- function(scores, models, CI, order = levels(models)) {
+plot_reliability <- function(scores, models, CI, order = levels(models), alpha = 0.8) {
   
   if (CI == 95) {
     plot <- scores |> 
@@ -14,9 +14,9 @@ plot_reliability <- function(scores, models, CI, order = levels(models)) {
       pivot_wider(names_from = is_in, names_prefix = 'within_', values_from = n, values_fill = 0) |> 
       mutate(perc = within_TRUE/(within_FALSE + within_TRUE)*100, 
              manuscript_name = factor(manuscript_name, levels = order)) |> 
-      ggplot(aes(x=horizon, y=perc, colour = manuscript_name)) +
-      geom_hline(yintercept = 95, colour = 'grey3', linetype = 'dashed')+
-      geom_line(alpha = 0.8) +
+      ggplot(aes(x=horizon, y=perc, colour = manuscript_name, linetype = manuscript_name)) +
+      geom_hline(yintercept = 95, colour = 'grey3', linetype = 'dashed', linewidth = 0.8)+
+      geom_line(linewidth = 0.8, alpha = alpha) +
       labs(y = 'Percentage of observations within\n95% confidence intervals', x='Horizon (days)')  +
       annotate('text', x = 30, y = 100, label = 'underconfident', size = 5, hjust = 1) +
       annotate('text', x = 30, y = 75, label = 'overconfident', size = 5, hjust = 1)
@@ -35,9 +35,9 @@ plot_reliability <- function(scores, models, CI, order = levels(models)) {
       pivot_wider(names_from = is_in, names_prefix = 'within_', values_from = n, values_fill = 0) |> 
       mutate(perc = within_TRUE/(within_FALSE + within_TRUE)*100, 
              manuscript_name = factor(manuscript_name, levels = order)) |> 
-      ggplot(aes(x=horizon, y=perc, colour = manuscript_name)) +
-      geom_hline(yintercept = 80, colour = 'grey3', linetype = 'dashed')+
-      geom_line(alpha = 0.8) +
+      ggplot(aes(x=horizon, y=perc, colour = manuscript_name, linetype = manuscript_name)) +
+      geom_hline(yintercept = 80, colour = 'grey3', linetype = 'dashed', linewidth = 0.8)+
+      geom_line(linewidth = 0.8, alpha = alpha) +
[…]

### 2024-09-03 supplementary_plots.Rmd
reverse the calculation of the relative skill metrics

@@
   # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = crps_climatology - crps,
-         skill_logs = logs_climatology - logs,
+  mutate(skill_crps = crps - crps_climatology,
+         skill_logs = logs - logs_climatology,
          horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
@@
   # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = crps_persistenceRW - crps,
-         skill_logs = logs_persistenceRW - logs,
+  mutate(skill_crps = crps - crps_persistenceRW,
+         skill_logs = logs - logs_persistenceRW,
          horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
@@
   # calcaulte the skill relative to persistenceRW (- is bad, + is good)
-  mutate(skill_crps = crps_persistenceRW - crps,
-         skill_logs = logs_persistenceRW - logs,
+  mutate(skill_crps = crps - crps_persistenceRW,
+         skill_logs = logs - logs_persistenceRW,
          horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 

### 2024-09-03 R/load_data.R

@@ -41,8 +41,8 @@ temperature_skill <- temperature_scores |>
   full_join(temperature_climatology) |> 
   
   # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = crps_climatology - crps,
-         skill_logs = logs_climatology - logs,
+  mutate(skill_crps = crps - crps_climatology,
+         skill_logs = logs - logs_climatology,
          horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
   # consistent forecast period
   filter(horizon <= 30,
@@ -57,8 +57,8 @@ oxygen_skill <- oxygen_scores |>
   full_join(oxygen_climatology) |> 
   
   # calcaulte the skill relative to climatology (- is bad, + is good)
-  mutate(skill_crps = crps_climatology - crps,
-         skill_logs = logs_climatology - logs,
+  mutate(skill_crps = crps - crps_climatology,
+         skill_logs = logs - logs_climatology,
          horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
   # consistent forecast period
   filter(horizon <= 30,

### 2024-09-03 manuscript_plots.Rmd
updates for review

@@
   ungroup() |>
-  filter(mean_crps > 0) |>
+  filter(mean_crps < 0) |>
   left_join(model_meta) |>
-  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
+  mutate(manuscript_name = fct_reorder(manuscript_name, desc(mean_crps)),
          uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'),
@@
   left_join(model_meta) |> 
-  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
+  mutate(manuscript_name = fct_reorder(manuscript_name, desc(mean_crps)),
          uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'), 
@@
   ungroup() |> 
-  filter(mean_crps > 0) |> 
+  filter(mean_crps < 0) |> 
   distinct(model_id)

### 2024-09-03 R/get_colours.R

@@ -3,7 +3,7 @@ top_mods <- temperature_skill |>
   group_by(model_id) |> 
   summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
   ungroup() |> 
-  filter(mean_crps > 0) |> 
+  filter(mean_crps < 0) |> 
   left_join(select(model_meta, model_id, manuscript_name, model_type)) |> 
   arrange(model_type) |> 
   # also include the baseline models
@@ -34,8 +34,8 @@ model_id_colours <- enframe(c(process_cols, MME_cols, empirical_cols, ML_cols)) 
   full_join(top_mods, by = join_by(name == model_id)) |> 
   mutate(value = ifelse(is.na(value), '#A9A9A9', value),
          mean_crps = ifelse(is.na(mean_crps), 0, mean_crps),
-         manuscript_name = fct_reorder(manuscript_name, mean_crps, .desc = T)) |> 
-  arrange(desc(mean_crps)) %>% 
+         manuscript_name = fct_reorder(manuscript_name, desc(mean_crps), .desc = T)) |> 
+  arrange(mean_crps) %>% 
   mutate(linetype = ifelse(name == 'climatology', 'dotdash', 'solid'))
 
 # colours for model types 

### 2024-09-03 manuscript_plots.Rmd
updates for review

@@
   summarise(n = n())  |> 
-  mutate(site_id = 'TOOK', mean_skill = -3.25) |> 
+  mutate(site_id = 'TOOK', mean_skill = 3.25) |> 
   filter(model_type != 'null')

### 2024-09-03 supplementary_plots.Rmd
updates for review

@@
   ungroup() |> 
-  slice_max(mean_crps, n = 10, na_rm = T) |> 
+  filter(mean_crps < 0) |> 
   distinct(model_id)
@@
   geom_vline(xintercept = 0, linetype = 'solid') +
-  geom_label(data = n_model_type, aes(label = paste0('n = ', n)))+ 
+  geom_text(data = n_model_type, aes(label = paste0('n = ', n)))+
   geom_boxplot() +
@@
   labs(y='Site', x = 'mean relative skill (°C)') +
-  scale_colour_manual(values = cols_modeltype, name = 'Model Type') +
+  scale_colour_manual(values = cols_modeltype, name = 'Model class') +
   theme_bw()
+
+lake_order <- temperature_skill |> 
+  filter(model_id %in% top_mods$model_id) |> 
+  group_by(site_id, horizon) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) %>% 
+  filter(horizon == 30) %>% 
+  arrange(desc(mean_crps)) %>% 
+  pull(site_id)
 
@@
   ungroup() |>
-  filter(mean_crps > 0) |>
+  filter(mean_crps < 0) |>
   left_join(model_meta) |>
@@
   theme_bw() + 
-  labs(x= 'Horizon (days)', y = 'Mean CRPS (degrees C)') + 
+  labs(x= 'Horizon (days)', y = 'Mean CRPS (°C)') + 
   scale_colour_viridis_d(name = 'Null Model', begin = 0.2, end=0.9) +
@@
           by = join_by(model_id, horizon, manuscript_name)) %>% 
-  mutate(beats = ifelse(mean_crpsDOY > 0 & mean_crpsRW > 0, 'both',
-                        ifelse(mean_crpsDOY < 0 & mean_crpsRW > 0, 'RW',
-                               ifelse(mean_crpsDOY > 0 & mean_crpsRW < 0, 'DOY', 'none')))) %>% 
+  mutate(beats = ifelse(mean_crpsDOY < 0 & mean_crpsRW < 0, 'both',
+                        ifelse(mean_crpsDOY > 0 & mean_crpsRW < 0, 'RW',
+                               ifelse(mean_crpsDOY < 0 & mean_crpsRW > 0, 'DOY', 'none')))) %>% 
   mutate(beats = factor(beats, levels = c("none", "RW", "DOY", "both"), labels = c("Neither", "Persistence", "Day-of-year", "Both"))) %>% 

### 2024-09-04 supplementary_plots.Rmd
updates for review

@@
 ```{r}
-skill_summary_RW <- temperature_skill_RW |>
+temp_skill_RW <- temperature_skill_RW |>
+  group_by(model_id) |>
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |>
+  ungroup() |>
+  filter(mean_crps < 0) |>
+  left_join(model_meta) |>
+  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
+         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'),
+         uses_AT = str_detect(NOAA_var, 'air_temperature')) |>
+  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +    # Different pattern for each group
+  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
+                   stat = "identity",
+                   alpha = 0.8,
+                   pattern_fill = "black",
+                   pattern_colour = "black",
+                   pattern_spacing = 0.04,
+                   pattern_alpha = 1,
+                   colour = 'black') +
+  scale_y_reordered() +
+  labs(y='Model', x = 'Mean relative skill (°C)') +
+  geom_vline(xintercept = 0) +
+  scale_fill_manual(values = cols_modeltype) +
+  scale_pattern_manual(values=c('none', 'stripe')) +
+  scale_pattern_type_manual(values=c(NA, NA)) +
+  theme_bw(base_size = 10) +
+  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02),
+                                title = 'Uses air temperature\nas a covariate?'),
+         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class')) +
+    theme(legend.box = 'horizontal')
+
+DO_skill_RW <- oxygen_skill_RW |>
   group_by(model_id) |>
@@
   ggsave(filename = 'Plots/FigureSnew.png', height = 17, width = 18, units = 'cm')
+
+ggsave(skill_horizon_RW, filename = 'Plots/FigureSnew1.png', height = 10, width = 13, units = 'cm')
 ```
+
 ```{r}
-null_scores <- temperature_scores |> mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) %>% 
+null_scores_T <- temperature_scores |>
+  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) %>% 
   filter(model_id %in% baselines$model_id, 
@@
   theme_bw() + 
-  labs(x= 'Horizon (days)', y = 'Mean CRPS (°C)') + 
+  labs(x= 'Horizon (days)', y = 'Mean CRPS (°C)', title = 'Water temperature') + 
   scale_colour_viridis_d(name = 'Null Model', begin = 0.2, end=0.9) +
@@
 
-ggsave(null_scores, filename = 'Plots/Figure_Snew2.png', height = 10, width = 15, units = 'cm')
+null_scores_DO <- oxygen_scores |>
+  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) %>% 
+  filter(model_id %in% baselines$model_id, 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(model_id, horizon) |> 
+  summarise(mean_crps = mean(crps, na.rm = T)) %>% 
+  mutate(model_id = factor(model_id, levels = c('climatology', 'persistenceRW'), labels = c("day-of-year", "persistence"))) %>% 
+  ggplot(aes(x=horizon,y = mean_crps, colour = model_id)) + 
+  geom_line() + 
+  scale_x_continuous(breaks = seq(2,30, 2)) + 
+  theme_bw() + 
+  labs(x= 'Horizon (days)', y = 'Mean CRPS (mg/L)', title = 'Dissolved oxygen') + 
+  scale_colour_viridis_d(name = 'Null Model', begin = 0.2, end=0.9) +
+  theme(panel.grid.minor = element_blank())
+
+ggarrange(null_scores_T, null_scores_DO, common.legend = T) %>% 
+  ggsave(filename = 'Plots/Figure_Snew2.png', height = 10, width = 18, units = 'cm')
 ```

### 2024-10-07 supplementary_plots.Rmd
edits for revisions

--- 2024-09-03 supplementary_plots.Rmd
@@
 
-  # calcaulte the skill relative to climatology (- is bad, + is good)
+  # calcaulte the skill relative to climatology (+ is bad, - is good)
   mutate(skill_crps = crps - crps_climatology,
@@
 
-## Figure S2 - all models, skill over horizon
+## Figure S2 - baseline performance
+```{r}
+null_scores_T <- temperature_scores |>
+  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) %>% 
+  filter(model_id %in% baselines$model_id, 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(model_id, horizon) |> 
+  summarise(mean_crps = mean(crps, na.rm = T)) %>% 
+  mutate(model_id = factor(model_id, levels = c('climatology', 'persistenceRW'), 
+                           labels = c("day-of-year", "persistence"))) %>% 
+  ggplot(aes(x=horizon,y = mean_crps, colour = model_id)) + 
+  geom_line() + 
+  scale_x_continuous(breaks = seq(2,30, 2)) + 
+  theme_bw() + 
+  labs(x= 'Horizon (days)', y = 'Mean CRPS (°C)', title = 'Water temperature') + 
+  scale_colour_viridis_d(name = 'Null Model', begin = 0.2, end=0.9) +
+  theme(panel.grid.minor = element_blank())
+
+null_scores_DO <- oxygen_scores |>
+  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) %>% 
+  filter(model_id %in% baselines$model_id, 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(model_id, horizon) |> 
+  summarise(mean_crps = mean(crps, na.rm = T)) %>% 
+  mutate(model_id = factor(model_id, levels = c('climatology', 'persistenceRW'), 
+                           labels = c("day-of-year", "persistence"))) %>% 
+  ggplot(aes(x=horizon,y = mean_crps, colour = model_id)) + 
+  geom_line() + 
+  scale_x_continuous(breaks = seq(2,30, 2)) + 
+  theme_bw() + 
+  labs(x= 'Horizon (days)', y = 'Mean CRPS (mg/L)', title = 'Dissolved oxygen') + 
+  scale_colour_viridis_d(name = 'Null Model', begin = 0.2, end=0.9) +
+  theme(panel.grid.minor = element_blank())
+
+ggarrange(null_scores_T, null_scores_DO, common.legend = T) %>% 
+  ggsave(filename = 'Plots/FigureS2.png', height = 10, width = 18, units = 'cm')
+```
+
+## Figure S3 - bad models, skill over horizon
 
@@
   ggplot(aes(x=horizon, 
-             y = mean_crps, colour = manuscript_name, linetype = manuscript_name)) +
+             y = mean_crps, colour = manuscript_name)) +
   geom_hline(yintercept = 0, colour = 'grey40' ) + 
-  geom_line(show.legend = F)  +  
+  geom_line()  +  
   labs(y='Mean relative skill (°C)', x = 'Horizon (days)') +
+  scale_colour_manual(name = '', values = unname(pals::cols25(n= 20))) + 
+  theme_bw() +
+  theme(legend.position = 'none')
+
+ggsave(plot = horizon_skill, 'Plots/FigureS3.png', width = 15, height = 15, units = 'cm')
+```
+## Figure S4 - skill vs persistence over horizon
+```{r}
+skill_horizon_RW <- temperature_skill_RW |> 
+  filter(model_id %in% c(top_mods$model_id) & model_id != 'climatology', 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(model_id, horizon) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
+  left_join(select(model_meta, model_id, manuscript_name)) |> 
+  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name))) |> 
+  ggplot(aes(x=horizon, 
+             y = mean_crps, colour = manuscript_name)) +
+  geom_hline(yintercept = 0, colour = 'grey40' ) + 
+  geom_line(linewidth = 0.8)  +  
+  labs(y='Mean relative skill (°C)', x = 'Horizon (days)') +
+  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
   theme_bw()
 
-ggsave(plot = horizon_skill, 'Plots/FigureS2.png', width = 10, height = 10, units = 'cm')
+ggsave(skill_horizon_RW, filename = 'Plots/FigureS4.png', height = 10, width = 13, units = 'cm')
 ```
 
-## Figure S3 - performance of null models
+## Figure S5 - summary of performance vs baselines
 
 ```{r}
-baseline_forecasts <- arrow::open_dataset("scores") |>
-  filter(site_id %in% lake_sites,
-         model_id %in% c('persistenceRW', 'climatology')) |>
-  collect()
 
-baseline1<-baseline_forecasts |> 
-  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime)),
-         site_id = factor(site_id, levels = lake_sites)) |> 
-  group_by(site_id, model_id, horizon) |> 
-  summarise(crps_median = median(crps, na.rm = T), .groups = 'drop') |> 
-  ggplot(aes(x=horizon, y=crps_median, colour = site_id)) +
-  geom_line() +
-  facet_wrap(~model_id) +
-  scale_colour_brewer(palette = 'Dark2', name = 'Site') +
-  labs(y='Median CRPS (°C)', x = 'Horizon (days)') +
-  theme_bw()
+temp_skill_RW_summary <- temperature_skill_RW |> 
+  filter(model_id %in% c(top_mods$model_id) & model_id != 'climatology', 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(model_id, horizon) |> 
+  summarise(mean_crpsRW = mean(skill_crps, na.rm = T)) |> 
+  left_join(select(model_meta, model_id, manuscript_name)) |> 
+  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name)))
 
-baseline2<-baseline_forecasts |> 
-  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) |> 
-  group_by(site_id, model_id, reference_datetime) |> 
-  summarise(crps_median = median(crps, na.rm = T), .groups = 'drop') |> 
-  ggplot(aes(x = as_date(reference_datetime), y=crps_median, colour = site_id)) +
-  geom_line() +
-  facet_wrap(~model_id, scale = 'free_y') +
-  scale_colour_brewer(palette = 'Dark2', name = 'Site') +
-  labs(y='Median CRPS (°C)', x = 'Forecast date') +
-  theme_bw()
+temp_skill_clim_summary <- temperature_skill |> 
+  filter(model_id %in% c(top_mods$model_id) & model_id != 'persistence', 
+         horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(model_id, horizon) |> 
+  summarise(mean_crpsDOY = mean(skill_crps, na.rm = T)) |> 
+  left_join(select(model_meta, model_id, manuscript_name)) |> 
+  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name)))
 
-ggarrange(baseline1, baseline2, nrow = 2) |> 
-  ggsave(filename = './Plots/FigureS3.png', width = 19, height = 16, units = 'cm')
+null_model_beats <- full_join(temp_skill_clim_summary, temp_skill_RW_summary, 
+          by = join_by(model_id, horizon, manuscript_name)) %>% 
+  mutate(beats = ifelse(mean_crpsDOY < 0 & mean_crpsRW < 0, 'both',
+                        ifelse(mean_crpsDOY > 0 & mean_crpsRW < 0, 'RW',
+                               ifelse(mean_crpsDOY < 0 & mean_crpsRW > 0, 'DOY', 'none')))) %>% 
+  mutate(beats = factor(beats, levels = c("none", "RW", "DOY", "both"), labels = c("Neither", "Persistence", "Day-of-year", "Both"))) %>% 
+  na.omit() %>% 
+  ggplot(aes(x=horizon, y=manuscript_name, shape = beats, colour = beats)) + 
+  geom_point(size = 5) +
+  scale_shape_manual(values = c(0, 1, 4, 13), name = 'Null model that is out-performed') +
+  scale_colour_manual(values = c("grey90", "#455BCDFF", "#E1DD37FF", "#46F884FF"), name = 'Null model that is out-performed') +
+  scale_y_discrete(name = 'Model') +
+  scale_x_continuous(name = 'Horizon (days)', breaks = c(1,10, 20, 30), expand = c(0.02,0)) +
+  geom_hline(yintercept = seq(.5, 9.5, 1), linewidth = .2) +
+  theme(legend.position = "bottom", 
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank(),
+        panel.background = element_blank(),
+        legend.text = element_text(size = 8),
+        legend.title = element_text(size = 8),
+        axis.ticks.y = element_blank())
+
+ggsave(null_model_beats, filename = 'Plots/FigureS5.png', height = 10, width = 18, units = 'cm')
+
 ```
 
-## Figure S4
+## Figure S6 - performance of DOY models over horizon/time
 
-Plotting only dates where all sites are evaluated shows the same pattern as when all dates are considered. 
+```{r}
+lake_order <- temperature_skill |> 
+  filter(model_id %in% top_mods$model_id) |> 
+  group_by(site_id, horizon) |> 
+  summarise(mean_crps = mean(skill_crps, na.rm = T)) %>% 
+  filter(horizon == 30) %>% 
+  arrange(desc(mean_crps)) %>% 
+  pull(site_id)
+
+baselines<- temperature_climatology |>
+  mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) %>% 
+  filter(horizon > 0, 
+         horizon != 15,
+         horizon <=30) |> 
+  group_by(horizon, site_id) |> 
+  summarise(median_crps = median(crps_climatology, na.rm = T)) %>% 
+  mutate(site_id = factor(site_id, levels = lake_order)) |>
+  ggplot(aes(x=horizon,y = median_crps, colour = site_id)) + 
+  geom_line() + 
+  scale_x_continuous(breaks = seq(2,30, 2)) + 
+  theme_bw() + 
+  labs(x= 'Horizon (days)', y = 'Median CRPS (°C)') + 
+  scale_colour_brewer(palette = 'Dark2', name = 'Site') +
+  theme(panel.grid.minor = element_blank())
+
+ggsave(baselines, filename = './Plots/FigureS6.png', width = 10, height = 10, units = 'cm')
+```
+
+## Figure S7 - Plotting only dates where all sites are evaluated shows the same pattern as when all dates are considered. 
 ```{r overlapping-periods}
@@
 cowplot::plot_grid(A, B, nrow = 2, rel_heights = c(1.5, 1), align = 'v', axis = 'lr', labels = 'auto') |> 
-  ggsave(filename = './Plots/FigureS4.png', width = 19, height = 16, units = 'cm')
+  ggsave(filename = './Plots/FigureS7.png', width = 19, height = 16, units = 'cm')
 ```
-```{r}
-skill_summary_RW <- temperature_skill_RW |>
-  group_by(model_id) |>
-  summarise(mean_crps = mean(skill_crps, na.rm = T)) |>
-  ungroup() |>
-  filter(mean_crps < 0) |>
-  left_join(model_meta) |>
-  mutate(manuscript_name = fct_reorder(manuscript_name, mean_crps),
-         uses_NOAA = ifelse(uses_NOAA == 1, 'yes', 'no'),
-         uses_AT = str_detect(NOAA_var, 'air_temperature')) |>
-  ggplot(aes(x=mean_crps, y = manuscript_name, fill = model_type)) +    # Different pattern for each group
-  geom_bar_pattern(aes(pattern = as.factor(uses_AT)),
-                   stat = "identity",
-                   alpha = 0.8,
-                   pattern_fill = "black",
-                   pattern_colour = "black",
-                   pattern_spacing = 0.04,
-                   pattern_alpha = 1,
-                   colour = 'black') +
-  scale_y_reordered() +
-  labs(y='Model', x = 'Mean relative skill (°C)') +
-  geom_vline(xintercept = 0) +
-  scale_fill_manual(values = cols_modeltype) +
-  scale_pattern_manual(values=c('none', 'stripe')) +
-  scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw(base_size = 10) +
-  guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02),
-                                title = 'Uses air temperature\nas a covariate?'),
-         fill = guide_legend(override.aes = list(pattern = "none"), title = 'Model class')) +
-    theme(legend.box = 'horizontal')
-
-skill_summary_RW_nolegend <- skill_summary_RW +
-  theme(legend.position = 'none')
-
-skill_horizon_RW <- temperature_skill_RW |> 
-  filter(model_id %in% c(top_mods$model_id) & model_id != 'climatology', 
-         horizon > 0, 
-         horizon != 15,
-         horizon <=30) |> 
-  group_by(model_id, horizon) |> 
-  summarise(mean_crps = mean(skill_crps, na.rm = T)) |> 
-  left_join(select(model_meta, model_id, manuscript_name)) |> 
-  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name))) |> 
-  ggplot(aes(x=horizon, 
-             y = mean_crps, colour = manuscript_name)) +
-  geom_hline(yintercept = 0, colour = 'grey40' ) + 
-  geom_line(linewidth = 0.8)  +  
-  labs(y='Mean relative skill (°C)', x = 'Horizon (days)') +
-  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
-  theme_bw()
-
-ggpubr::ggarrange(ggarrange(NULL, as_ggplot(get_legend(skill_summary_RW)), NULL, NULL, widths =  c(0.2, 0.9,0.2, 1.1), nrow = 1),
-                  ggarrange(skill_summary_RW_nolegend,  NULL, skill_horizon_RW, align = 'h', widths =  c(0.9,0.05, 1.1), nrow = 1), 
-                  heights = c(0.5, 1),nrow = 2) |> 
-  ggsave(filename = 'Plots/FigureSnew.png', height = 17, width = 18, units = 'cm')
-```
-```{r}
-null_scores <- temperature_scores |> mutate(horizon = as.numeric(as_date(datetime) - as_date(reference_datetime))) %>% 
-  filter(model_id %in% baselines$model_id, 
-         horizon > 0, 
-         horizon != 15,
-         horizon <=30) |> 
-  group_by(model_id, horizon) |> 
-  summarise(mean_crps = mean(crps, na.rm = T)) %>% 
-  mutate(model_id = factor(model_id, levels = c('climatology', 'persistenceRW'), labels = c("day-of-year", "persistence"))) %>% 
-  ggplot(aes(x=horizon,y = mean_crps, colour = model_id)) + 
-  geom_line() + 
-  scale_x_continuous(breaks = seq(2,30, 2)) + 
-  theme_bw() + 
-  labs(x= 'Horizon (days)', y = 'Mean CRPS (°C)') + 
-  scale_colour_viridis_d(name = 'Null Model', begin = 0.2, end=0.9) +
-  theme(panel.grid.minor = element_blank())
-
-ggsave(null_scores, filename = 'Plots/Figure_Snew2.png', height = 10, width = 15, units = 'cm')
-```
-
-```{r}
-
-temp_skill_RW_summary <- temperature_skill_RW |> 
-  filter(model_id %in% c(top_mods$model_id) & model_id != 'climatology', 
-         horizon > 0, 
-         horizon != 15,
-         horizon <=30) |> 
-  group_by(model_id, horizon) |> 
-  summarise(mean_crpsRW = mean(skill_crps, na.rm = T)) |> 
-  left_join(select(model_meta, model_id, manuscript_name)) |> 
-  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name)))
-
-temp_skill_clim_summary <- temperature_skill |> 
-  filter(model_id %in% c(top_mods$model_id) & model_id != 'persistence', 
-         horizon > 0, 
-         horizon != 15,
-         horizon <=30) |> 
-  group_by(model_id, horizon) |> 
-  summarise(mean_crpsDOY = mean(skill_crps, na.rm = T)) |> 
-  left_join(select(model_meta, model_id, manuscript_name)) |> 
-  mutate(manuscript_name = factor(manuscript_name, levels = levels(model_id_colours$manuscript_name)))
-
-null_model_beats <- full_join(temp_skill_clim_summary, temp_skill_RW_summary, 
-          by = join_by(model_id, horizon, manuscript_name)) %>% 
-  mutate(beats = ifelse(mean_crpsDOY < 0 & mean_crpsRW < 0, 'both',
-                        ifelse(mean_crpsDOY > 0 & mean_crpsRW < 0, 'RW',
-                               ifelse(mean_crpsDOY < 0 & mean_crpsRW > 0, 'DOY', 'none')))) %>% 
-  mutate(beats = factor(beats, levels = c("none", "RW", "DOY", "both"), labels = c("Neither", "Persistence", "Day-of-year", "Both"))) %>% 
-  na.omit() %>% 
-  ggplot(aes(x=horizon, y=manuscript_name, shape = beats, colour = beats)) + 
-  geom_point(size = 5) +
-  scale_shape_manual(values = c(0, 1, 4, 13), name = 'Null model that is out-performed') +
-  scale_colour_manual(values = c("grey90", "#455BCDFF", "#E1DD37FF", "#46F884FF"), name = 'Null model that is out-performed') +
-  scale_y_discrete(name = 'Model') +
-  scale_x_continuous(name = 'Horizon (days)', breaks = c(1,10, 20, 30), expand = c(0.02,0)) +
-  geom_hline(yintercept = seq(.5, 9.5, 1), linewidth = .2) +
-  theme(legend.position = "bottom", 
-        panel.grid.major = element_blank(),
-        panel.grid.minor = element_blank(),
-        panel.background = element_blank(),
-        legend.text = element_text(size = 8),
-        legend.title = element_text(size = 8),
-        axis.ticks.y = element_blank())
-
-ggsave(null_model_beats, filename = 'Plots/Figure_Snew3.png', height = 10, width = 18, units = 'cm')
-
-```

### 2024-10-08 manuscript_plots.Rmd
tidy plot sizes

@@
   scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw(base_size = 10) +
+  theme_bw(base_size = 12) +
   guides(pattern = guide_legend(override.aes = list(fill = "white", pattern_spacing = 0.02),
@@
   scale_pattern_type_manual(values=c(NA, NA), name = 'Uses air temperature as a covariate?') +
-  theme_bw(base_size = 10) + 
+  theme_bw(base_size = 12) + 
   guides(pattern = guide_legend(override.aes = list(fill = "white"), title = 'Uses air temperature as a covariate?'),
@@
   scale_pattern_type_manual(values=c(NA, NA)) +
-  theme_bw(base_size = 10) + 
+  theme_bw(base_size = 12) + 
   theme(legend.position = 'none')
@@
                   legend = 'none', labels = c('a', '', 'b', 'c')) |> 
-  ggsave(filename = 'Plots/Figure2.png', height = 20, width = 17, units = 'cm')
+  ggsave(filename = 'Plots/Figure2.png', height = 22, width = 18, units = 'cm')
 
@@
   scale_colour_manual(name = 'Model',values = model_id_colours$value) +
-  theme_bw()
+  theme_bw(base_size = 12)
 
@@
   scale_colour_manual(name = 'Model',values = model_id_colours$value) +
-  theme_bw() +
+  theme_bw(base_size = 12) +
   coord_cartesian(ylim = c(0, 2))
@@
                    order = levels(model_id_colours$manuscript_name)) +
-  scale_colour_manual(name = 'Top model',values = model_id_colours$value) +
-  scale_linetype_manual(name = 'Top model',values = model_id_colours$linetype) +
-  theme_bw() +
+  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
+  scale_linetype_manual(name = 'Model',values = model_id_colours$linetype) +
+  theme_bw(base_size = 10) +
   scale_y_continuous(limits = c(54, 100)) +
   labs(title = 'Top performing models') +
-  theme(title = element_text(size = 14))
+  theme(title = element_text(size = 12))
 
@@
                             order = levels(model_id_colours$manuscript_name)) +
-  scale_colour_manual(name = 'Top model',values = model_id_colours$value) +
-  scale_linetype_manual(name = 'Top model',values = model_id_colours$linetype) +
-  theme_bw() +
+  scale_colour_manual(name = 'Model',values = model_id_colours$value) +
+  scale_linetype_manual(name = 'Model',values = model_id_colours$linetype) +
+  theme_bw(base_size = 10) +
   scale_y_continuous(limits = c(54, 100))
@@
                    alpha  = 0.6) +
-  theme_bw()  +
+  theme_bw(base_size = 10) +
   scale_colour_manual(name = '', values = unname(pals::cols25(n= 20)))+ 
@@
   theme(legend.position = 'none') +
-  labs(title = 'All other submitted models') +
-  theme(title = element_text(size = 14))
+  labs(title = 'Other submitted models') +
+  theme(title = element_text(size = 11.5))
 
@@
                            alpha  =0.6) +
-  theme_bw() +
+  theme_bw(base_size = 10) +
   scale_y_continuous(limits = c(0, 100))   +
@@
 
-ggarrange(NULL, ggarrange(get_legend(good_80), NULL, nrow = 2),
-          ggarrange(good_80, good_95, bad_80, bad_95, align = 'hv', legend= 'none', labels = 'auto', hjust = -4, vjust = 3), 
+ggarrange(NULL, ggarrange(get_legend(good_80), NULL, nrow = 2, heights = c(1,0.8)),
+          ggarrange(good_80, good_95, bad_80, bad_95, align = 'hv', legend= 'none', labels = 'auto', hjust = -4.5, vjust = 2), 
           ncol = 3, 
-          widths = c(0.1,1,5)) |> 
-  ggsave(filename = 'Plots/Figure4.png', height = 17.6, width = 23, units = 'cm')
+          widths = c(0.2,1.2,5)) |> 
+  ggsave(filename = 'Plots/Figure4.png', height = 15, width = 18, units = 'cm')
 ```
@@
   scale_colour_manual(values = cols_modeltype, name = 'Model class') +
-  theme_bw()  +
+  theme_bw(base_size =  12)  +
   theme(legend.position = 'bottom')
@@
   scale_colour_brewer(palette = 'Dark2', name = 'Site')+
-  theme_bw()
+  theme_bw(base_size =  12)
 
 cowplot::plot_grid(A, B, nrow = 2, rel_heights = c(1.5, 1), align = 'v', axis = 'tlbr', labels = 'auto') |> 
-  ggsave(filename = './Plots/Figure5.png', width = 19, height = 16, units = 'cm')
+  ggsave(filename = './Plots/Figure5.png', width = 18, height = 15, units = 'cm')
 ```
@@
   geom_hline(yintercept = 0) +
-  theme_bw() +
+  theme_bw(base_size =  12)  +
   annotate("rect", 
@@
   scale_x_date(date_labels = "%b", breaks = '1 month') +
-  labs(y = 'Difference in median monthly water temperature (°C)', 
+  labs(y = 'Difference in median monthly\nwater temperature (°C)', 
        x = 'Month') +

### 2024-10-08 R/reliability_plot.R

@@ -18,8 +18,8 @@ plot_reliability <- function(scores, models, CI, order = levels(models), alpha =
       geom_hline(yintercept = 95, colour = 'grey3', linetype = 'dashed', linewidth = 0.8)+
       geom_line(linewidth = 0.8, alpha = alpha) +
       labs(y = 'Percentage of observations within\n95% confidence intervals', x='Horizon (days)')  +
-      annotate('text', x = 30, y = 100, label = 'underconfident', size = 5, hjust = 1) +
-      annotate('text', x = 30, y = 75, label = 'overconfident', size = 5, hjust = 1)
+      annotate('text', x = 30, y = 100, label = 'underconfident', size = 4, hjust = 1) +
+      annotate('text', x = 30, y = 75, label = 'overconfident', size = 4, hjust = 1)
   }
   if (CI == 80) {
     plot <- scores |> 
@@ -39,8 +39,8 @@ plot_reliability <- function(scores, models, CI, order = levels(models), alpha =
       geom_hline(yintercept = 80, colour = 'grey3', linetype = 'dashed', linewidth = 0.8)+
       geom_line(linewidth = 0.8, alpha = alpha) +
       labs(y = 'Percentage of observations within\n80% confidence intervals', x='Horizon (days)') +
-      annotate('text', x = 30, y = 100, label = 'underconfident', size = 5, hjust = 1) +
-      annotate('text', x = 30, y = 75, label = 'overconfident', size = 5, hjust = 1)
+      annotate('text', x = 30, y = 100, label = 'underconfident', size = 4, hjust = 1) +
+      annotate('text', x = 30, y = 75, label = 'overconfident', size = 4, hjust = 1)
   }
   if (CI != 80 & CI != 95) {
     message('CI must be 80 or 95')
